### 2023-06-15 paperCharts.Rmd
additional paper data and charts improvements

---
title: "2040 greenhouse gas reduction targets"
output:
  html_document:
    theme: paper
    toc: true
    toc_float:
      collapsed: false
---

<style>
  .main-container {
    max-width: 95% !important;
  }
  .toc-content {
    padding-left: 0px !important;
  }
  .svg-container {
    margin: auto !important;
  }
</style>

Description:
Charts created for the paper:

Rodrigues, R., Pietzcker, R., Luderer, G., Sitarz, J., Merfort, A., Hasse, R., Hoppe, J., Pehl, M., Ershad, A. M., Baumstark, L. 2040 greenhouse gas reduction targets and energy transitions in line with the EU Green Deal. in preparation (2023).

<!-- Options -->
```{r options, echo=FALSE}

forceUpdateData = FALSE

mainScen <- "Nzero_55_bio7p5" # "Nzero_55_ff55Eff_bio7p5" #Nzero_55_ff55Eff_bio12" # #Nzero_55_bio12" #"Nzero_55_ff55Eff_bio12"

tgt2030Scen <- c(55,57,59)

```

<!-- Loading required packages -->
```{r Loading_required_packages, echo=FALSE}

# install.packages("remotes")
#remotes::install_github("coolbutuseless/ggpattern")

  packagesList <- c("quitte","dplyr","ggplot2","geomtextpath","tidyr","grid","stringr","gridExtra","ggrepel","kableExtra")
  packages <- suppressWarnings(suppressMessages(lapply(packagesList, function(x){ if(!require(x, character.only = T, quietly = T)){ install.packages(x); return(paste0("installed missing package: ", x)) } else paste0("required package is installed: ", x) } )))
```

<!-- Setting default chunk options -->
```{r chunkOptions, echo=FALSE}
#knitr::opts_chunk$set(echo = TRUE)
# setting global R chunk options (https://yihui.name/knitr/options/#chunk_options)
  knitr::opts_chunk$set(dev='svglite', # svg, png,...
                        fig.ext = ".svg",
                        #fig.asp = .8 # default aspect ratio
                        fig.width = 12,
                        fig.height = 6,
                        dpi=100
                        )
```

<!-- Aesthetics -->
```{r loading_aesthetics, echo=FALSE, include=FALSE}

#loading theme and colors
source("./R/aesthetics.R")

```

<!-- load data -->
```{r loading_data, echo=FALSE, include=FALSE}

#loading data
source("./R/loadData.R")

```

<!-- loading reference emissions -->
```{r loading_referenceEmissions, echo=FALSE}

#loading reference emissions
source("./R/loadReferenceEmissions.R")

```

<!-- harmonizing data -->
```{r harmonizingData, echo=FALSE}

#harmonizing data (scale emissions)
source("./R/harmonizeData.R")

```

<!-- efficiency target -->
```{r loading_efficiencyTarget, echo=FALSE}

#loading reference emissions
source("./R/efficiencyTarget.R")

```

---

# Data {.tabset}

<!-- Outputs -->

## Reference emissions

```{r, echo=FALSE}

selectedTgt %>%
  replace(is.na(.),"") %>% 
  select(target,region,period,tgt55,tgt57,tgt59,tgt61,tgt63,tgt65) %>% 
  kbl() %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"))

```

---

## Efficiency targets

```{r, echo=FALSE}

effTgt %>%
  kbl() %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"))

```

---

## 2040 Emission Reductions

```{r, echo=FALSE}

reductions2040 <- rbind(
  remindEmi %>% 
  filter(variable == "Emissions with LULUCF and with international transport", region == "EU27", period == 2040, tgt2050 != "Npi") %>% 
  mutate(value = 1-(value / tgt %>% filter(target == "Emissions with LULUCF and with international transport") %>% pull(reference)))
  ,
  remindEmi %>% 
  filter(variable == "Emissions without LULUCF and with international transport", region == "EU27", period == 2040, tgt2050 != "Npi") %>% 
  mutate(value = 1-(value / tgt %>% filter(target == "Emissions without LULUCF and with international transport") %>% pull(reference)))
  ,
  remindEmi %>% 
  filter(variable == "Emissions with LULUCF and with intra-EU aviation", region == "EU27", period == 2040, tgt2050 != "Npi") %>% 
  mutate(value = 1-(value / tgt %>% filter(target == "Emissions with LULUCF and with intra-EU aviation") %>% pull(reference)))
  ,
  remindEmi %>% 
  filter(variable == "Emissions without LULUCF and with intra-EU aviation", region == "EU27", period == 2040, tgt2050 != "Npi") %>% 
  mutate(value = 1-(value / tgt %>% filter(target == "Emissions with LULUCF and with intra-EU aviation") %>% pull(reference)))
)

reductions2040 %>% 
  filter(tgt2030 %in% tgt2030Scen) %>% 
  group_by(variable) %>%
  summarize(minTarget = min(value), maxTarget=max(value),.groups = "keep") %>%
  mutate(minTarget = paste0(round(minTarget*100), "%"),
         maxTarget = paste0(round(maxTarget*100), "%"),
         period = 2040) %>% 
  rename(TargetReference = variable) %>%
  relocate(period, .after=TargetReference) %>%
  kbl() %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"))

bdgt <- remindEmi %>% 
  filter(region == "EU27", period %in% c(2020:2050), tgt2030 %in% tgt2030Scen) %>%
    mutate(weight = ifelse(period %in% c(2020,2050),2.5,5),
           budget = value*weight) %>%
    group_by(model,scenario,region,variable,tgt2050,tgt2030,efficiency,bioLim,limCCS) %>%
    summarize(budget=round(sum(budget)/1000,2),.groups = "keep")

budget <- data.frame(
  Budget=c("Emissions with LULUCF and with international transport","Emissions without LULUCF and with international transport","Emissions with LULUCF and with intra-EU aviation","Emissions without LULUCF and with intra-EU aviation"),
  reference_value=c(round(bdgt %>% filter(variable == "Emissions with LULUCF and with international transport", scenario == mainScen) %>% pull(budget)),
                    round(bdgt %>% filter(variable == "Emissions without LULUCF and with international transport", scenario == mainScen) %>% pull(budget)),
                    round(bdgt %>% filter(variable == "Emissions with LULUCF and with intra-EU aviation", scenario == mainScen) %>% pull(budget)),
                    round(bdgt %>% filter(variable == "Emissions without LULUCF and with intra-EU aviation", scenario == mainScen) %>% pull(budget))),
  min = c(round(min(bdgt %>% filter(variable == "Emissions with LULUCF and with international transport") %>% pull(budget))),
                    round(min(bdgt %>% filter(variable == "Emissions without LULUCF and with international transport") %>% pull(budget))),
                    round(min(bdgt %>% filter(variable == "Emissions with LULUCF and with intra-EU aviation") %>% pull(budget))),
                    round(min(bdgt %>% filter(variable == "Emissions without LULUCF and with intra-EU aviation") %>% pull(budget)))),
  max = c(round(max(bdgt %>% filter(variable == "Emissions with LULUCF and with international transport") %>% pull(budget))),
                    round(max(bdgt %>% filter(variable == "Emissions without LULUCF and with international transport") %>% pull(budget))),
                    round(max(bdgt %>% filter(variable == "Emissions with LULUCF and with intra-EU aviation") %>% pull(budget))),
                    round(max(bdgt %>% filter(variable == "Emissions without LULUCF and with intra-EU aviation") %>% pull(budget))))
)

budget %>% 
  kbl() %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"))

```

---

## Results

```{r, echo=FALSE}

coal_in_generation <- left_join(
  df %>%
  filter(region == "EU27", period == 2040, variable %in% c("SE|Electricity|Coal")) %>%  
  select(scenario,tgt2030,value)
  ,
  df %>% 
  filter(region == "EU27", period == 2040, variable %in% c("SE|Electricity")) %>%
  mutate(total = value) %>%  
  select(scenario,tgt2030,total)
  , by = join_by(scenario == scenario, tgt2030 == tgt2030)
  ) %>%
  mutate(percentage = value/total)

gas_in_generation <- left_join(
  df %>%
  filter(region == "EU27", period == 2040, variable %in% c("SE|Electricity|Gas|w/o CC","SE|Electricity|Gas|w/ CC")) %>%
  group_by(scenario,tgt2030) %>%
  summarize(value = sum(value),.groups = "keep")
  ,
  df %>% 
  filter(region == "EU27", period == 2040, variable %in% c("SE|Electricity")) %>%
  mutate(total = value) %>%  
  select(scenario,tgt2030,total)
  , by = join_by(scenario == scenario, tgt2030 == tgt2030)
  ) %>%
  mutate(percentage = value/total)

wind_and_Solar_gen <- left_join(
  df %>%
  filter(region == "EU27", period == 2040, variable %in% c("SE|Electricity|Wind|Offshore", "SE|Electricity|Wind|Onshore", "SE|Electricity|Solar|PV", "SE|Electricity|Solar|CSP")) %>%
  group_by(scenario,tgt2030) %>%
  summarize(value = sum(value),.groups = "keep")
  ,
  df %>% 
  filter(region == "EU27", period == 2040, variable %in% c("SE|Electricity")) %>%
  mutate(total = value) %>%  
  select(scenario,tgt2030,total)
  , by = join_by(scenario == scenario, tgt2030 == tgt2030)
  ) %>%
  mutate(percentage = value/total)

wind_and_Solar_cap <- df %>%
  filter(region == "EU27", period == 2040, variable %in% c("Cap|Electricity|Wind|Offshore", "Cap|Electricity|Wind|Onshore", "Cap|Electricity|Solar|PV", "Cap|Electricity|Solar|CSP")) %>%
  group_by(scenario,tgt2030) %>%
  summarize(value = sum(value),.groups = "keep")

nuclear_in_generation <- left_join(
  df %>%
  filter(region == "EU27", period == 2040, variable %in% c("SE|Electricity|Nuclear")) %>%  
  select(scenario,tgt2030,value)
  ,
  df %>% 
  filter(region == "EU27", period == 2040, variable %in% c("SE|Electricity")) %>%
  mutate(total = value) %>%  
  select(scenario,tgt2030,total)
  , by = join_by(scenario == scenario, tgt2030 == tgt2030)
  ) %>%
  mutate(percentage = value/total)

Generation <- data.frame(
  Electricity_Generation=c("coal_in_generation_(%)","gas_in_generation_(%)","nuclear_in_generation_(%)","wind_and_Solar_gen_(%)","wind_and_Solar_cap_(TW)"),
  reference_value=c(
    round(coal_in_generation %>% filter(scenario == mainScen) %>% pull(percentage)*100,1),
    round(gas_in_generation %>% filter(scenario == mainScen) %>% pull(percentage)*100,1),
    round(nuclear_in_generation %>% filter(scenario == mainScen) %>% pull(percentage)*100,1),
    round(wind_and_Solar_gen %>% filter(scenario == mainScen) %>% pull(percentage)*100,1),
    round(wind_and_Solar_cap %>% filter(scenario == mainScen) %>% pull(value)/ 1000,1)

    ),
  min = c(
   round(min(coal_in_generation %>% filter(tgt2030 %in% tgt2030Scen) %>% pull(percentage)*100),1),
   round(min(gas_in_generation %>% filter(tgt2030 %in% tgt2030Scen) %>% pull(percentage)*100),1),
   round(min(nuclear_in_generation %>% filter(tgt2030 %in% tgt2030Scen) %>% pull(percentage)*100),1),
   round(min(wind_and_Solar_gen %>% filter(tgt2030 %in% tgt2030Scen) %>% pull(percentage)*100),1),
   round(min(wind_and_Solar_cap %>% filter(tgt2030 %in% tgt2030Scen) %>% pull(value)/ 1000),1)
   ),
  max = c(
   round(max(coal_in_generation %>% filter(tgt2030 %in% tgt2030Scen) %>% pull(percentage)*100),1),
   round(max(gas_in_generation %>% filter(tgt2030 %in% tgt2030Scen) %>% pull(percentage)*100),1),
   round(max(nuclear_in_generation %>% filter(tgt2030 %in% tgt2030Scen) %>% pull(percentage)*100),1),
   round(max(wind_and_Solar_gen %>% filter(tgt2030 %in% tgt2030Scen) %>% pull(percentage)*100),1),
   round(max(wind_and_Solar_cap %>% filter(tgt2030 %in% tgt2030Scen) %>% pull(value)/ 1000),1)
   )
)

share_of_fossil_in_electricity <- data.frame(
  FE=c("share_of_fossil_without_CCS_in_electricity_(%)"),
  reference_value_2020=c(
    round((1-df %>% filter(region == "EU27", period == 2020, variable == "SE|Electricity|carbon-free", tgt2030 %in% tgt2030Scen, scenario == mainScen) %>% pull(value))*100 )),
  reference_value_2040=c(
    round((1-df %>% filter(region == "EU27", period == 2040, variable == "SE|Electricity|carbon-free", tgt2030 %in% tgt2030Scen, scenario == mainScen) %>% pull(value))*100 )),
  min = c(round(min((1-df %>% filter(region == "EU27", period == 2040, variable == "SE|Electricity|carbon-free", tgt2030 %in% tgt2030Scen) %>% pull(value))*100))),
  max = c(round(max((1-df %>% filter(region == "EU27", period == 2040, variable == "SE|Electricity|carbon-free", tgt2030 %in% tgt2030Scen) %>% pull(value))*100)))
)

pe <- df %>% 
  filter(region == "EU27", period %in% c(2020:2060), tgt2030 %in% tgt2030Scen, variable %in% c("PE|Biomass","PE|Coal","PE|Gas","PE|Oil"))

primaryEnergy <- data.frame(
  Primary_Energy=c("Biomass","Coal","Gas","Oil"),
  relative_to_2020=c(
    round((pe %>% filter(period == 2040, scenario == mainScen, variable == "PE|Biomass") %>% mutate(value = value / (pe %>% filter(period == 2020, scenario == mainScen, variable == "PE|Biomass") %>% pull(value)))%>% pull(value))*100, 1),
    round((pe %>% filter(period == 2040, scenario == mainScen, variable == "PE|Coal") %>% mutate(value = value / (pe %>% filter(period == 2020, scenario == mainScen, variable == "PE|Coal") %>% pull(value)))%>% pull(value))*100, 1),
    round((pe %>% filter(period == 2040, scenario == mainScen, variable == "PE|Gas") %>% mutate(value = value / (pe %>% filter(period == 2020, scenario == mainScen, variable == "PE|Gas") %>% pull(value)))%>% pull(value))*100, 1),
    round((pe %>% filter(period == 2040, scenario == mainScen, variable == "PE|Oil") %>% mutate(value = value / (pe %>% filter(period == 2020, scenario == mainScen, variable == "PE|Oil") %>% pull(value)))%>% pull(value))*100, 1))
)

fe_transport_electricity <- left_join(
  df %>%
  filter(region == "EU27", period == 2040, variable %in% c("FE|Transport|Electricity")) %>%  
  select(scenario,tgt2030,value)
  ,
  df %>% 
  filter(region == "EU27", period == 2030, variable %in% c("FE|Transport|Electricity")) %>%
  mutate(base = value) %>%  
  select(scenario,tgt2030,base)
  , by = join_by(scenario == scenario, tgt2030 == tgt2030)
  ) %>%
  mutate(percentage = value/base)

FE <- data.frame(
  FE=c("transport_electricity_(%)"),
  reference_value=c(round(fe_transport_electricity %>% filter(scenario == mainScen) %>% pull(percentage)*100,1)),
  min = c(round(min(fe_transport_electricity %>% filter(tgt2030 %in% tgt2030Scen) %>% pull(percentage)*100),1)),
  max = c(round(max(fe_transport_electricity %>% filter(tgt2030 %in% tgt2030Scen) %>% pull(percentage)*100),1))
)

Bunkers <- list(
  "Aviation" = c(
      "FE|Transport|Pass|Aviation|International|Electricity" = "Electricity",
      "FE|Transport|Pass|Aviation|International|Gases" = "Gases",
      "FE|Transport|Pass|Aviation|International|Hydrogen" = "Hydrogen",
#     "FE|Transport|Pass|Aviation|International|Liquids" = "Liquids"
      "FE|Transport|Pass|Aviation|International|Liquids|Fossil" = "Liquids - Fossil",
      "FE|Transport|Pass|Aviation|International|Liquids|Biomass" = "Liquids - Biomass",
      "FE|Transport|Pass|Aviation|International|Liquids|Hydrogen" = "Liquids - Synthetic"
    ),
    "Shipping" = c(
      "FE|Transport|Freight|Navigation|Electricity" = "Electricity",
      "FE|Transport|Freight|Navigation|Gases" = "Gases",
      "FE|Transport|Freight|Navigation|Hydrogen" = "Hydrogen",
#     "FE|Transport|Freight|Navigation|Liquids" = "Liquids",
      "FE|Transport|Freight|Navigation|Liquids|Fossil" = "Liquids - Fossil",
      "FE|Transport|Freight|Navigation|Liquids|Biomass" = "Liquids - Biomass",
      "FE|Transport|Freight|Navigation|Liquids|Hydrogen" = "Liquids - Synthetic",
      "FE|Transport|Freight|International Shipping|Electricity" = "Electricity",
      "FE|Transport|Freight|International Shipping|Gases" = "Gases",
      "FE|Transport|Freight|International Shipping|Hydrogen" = "Hydrogen",
#     "FE|Transport|Freight|International Shipping|Liquids" = "Liquids"
      "FE|Transport|Freight|International Shipping|Liquids|Fossil" = "Liquids - Fossil",
      "FE|Transport|Freight|International Shipping|Liquids|Biomass" = "Liquids - Biomass",
      "FE|Transport|Freight|International Shipping|Liquids|Hydrogen" = "Liquids - Synthetic"
    )
)

cleanAviation <- c("FE|Transport|Pass|Aviation|International|Electricity", "FE|Transport|Pass|Aviation|International|Hydrogen", "FE|Transport|Pass|Aviation|International|Liquids|Biomass", "FE|Transport|Pass|Aviation|International|Liquids|Hydrogen")

AviationData <- left_join(
  df %>% 
  filter(region == "EU27", period == 2040, variable %in% cleanAviation) %>%
  group_by(scenario,tgt2030) %>%
  summarize(value = sum(value),.groups = "keep")
  ,
  df %>% 
  filter(region == "EU27", period == 2040, variable %in% c(names(Bunkers$Aviation))) %>%
  group_by(scenario,tgt2030) %>%
  summarize(total = sum(value),.groups = "keep")
  , by = join_by(scenario == scenario, tgt2030 == tgt2030)
) %>%
  mutate(percentage = value/total)

cleanShipping <- c("FE|Transport|Freight|Navigation|Electricity", "FE|Transport|Freight|Navigation|Hydrogen", "FE|Transport|Freight|Navigation|Liquids|Biomass",  "FE|Transport|Freight|Navigation|Liquids|Hydrogen", "FE|Transport|Freight|International Shipping|Electricity", "FE|Transport|Freight|International Shipping|Hydrogen", "FE|Transport|Freight|International Shipping|Liquids|Biomass", "FE|Transport|Freight|International Shipping|Liquids|Hydrogen")

ShippingData <- left_join(
  df %>% 
  filter(region == "EU27", period == 2040, variable %in% cleanShipping) %>%
  group_by(scenario,tgt2030) %>%
  summarize(value = sum(value),.groups = "keep")
  ,
  df %>% 
  filter(region == "EU27", period == 2040, variable %in% c(names(Bunkers$Shipping))) %>%
  group_by(scenario,tgt2030) %>%
  summarize(total = sum(value),.groups = "keep")
  , by = join_by(scenario == scenario, tgt2030 == tgt2030)
) %>%
  mutate(percentage = value/total)

share_of_sustainable_fuel <- data.frame(
  FE=c("share_of_sustainable_fuel_in_aviation_(%)","share_of_sustainable_fuel_in_shipping_(%)"),
  reference_value=c(round(AviationData %>% filter(scenario == mainScen) %>% pull(percentage)*100,1),
                    round(ShippingData %>% filter(scenario == mainScen) %>% pull(percentage)*100,1)),
  min = c(round(min(AviationData %>% filter(tgt2030 %in% tgt2030Scen) %>% pull(percentage)*100),1),
          round(min(ShippingData %>% filter(tgt2030 %in% tgt2030Scen) %>% pull(percentage)*100),1)),
  max = c(round(max(AviationData %>% filter(tgt2030 %in% tgt2030Scen) %>% pull(percentage)*100),1),
          round(max(ShippingData %>% filter(tgt2030 %in% tgt2030Scen) %>% pull(percentage)*100),1))
)

build_UE_heat <- c(
  "Heat pump" = "UE|Buildings|Heating|Electricity|Heat pump",
  "District Heating" = "UE|Buildings|Heating|District Heating",
  "Resistance" = "UE|Buildings|Heating|Electricity|Resistance",
  "Hydrogen" = "UE|Buildings|Heating|Hydrogen",
  "Gases" = "UE|Buildings|Heating|Gases",
  "Liquids" = "UE|Buildings|Heating|Liquids",
  "Solids" = "UE|Buildings|Heating|Solids"#,
#  "non-Heating Electricity" = "UE|Buildings|non-Heating|Electricity|Conventional"
)

build_UE_heating_data <- left_join(
  df %>% 
  filter(region == "EU27", period == 2040, variable %in% build_UE_heat[c("Heat pump","District Heating","Resistance")], tgt2030 %in% tgt2030Scen) %>%
  group_by(scenario) %>%    
  summarize(value = sum(value),.groups = "keep")
  ,
  df %>% 
  filter(region == "EU27", period == 2040, variable %in% build_UE_heat, tgt2030 %in% tgt2030Scen) %>%
  group_by(scenario) %>%  
  summarize(total = sum(value)) %>% 
  select(scenario,total)  
  ,
  by = join_by(scenario == scenario)
) %>% 
  mutate(pctg = value/total)

build_UE_liq_gas_data <- left_join(
  df %>% 
  filter(region == "EU27", period == 2040, variable %in% build_UE_heat[c("Gases","Hydrogen","Liquids")], tgt2030 %in% tgt2030Scen) %>%
  group_by(scenario) %>%    
  summarize(value = sum(value),.groups = "keep")
  ,
  df %>% 
  filter(region == "EU27", period == 2040, variable %in% build_UE_heat, tgt2030 %in% tgt2030Scen) %>%
  group_by(scenario) %>%  
  summarize(total = sum(value)) %>% 
  select(scenario,total)  
  ,
  by = join_by(scenario == scenario)
) %>% 
  mutate(pctg = value/total)

build_UE_solids_data <- left_join(
  df %>% 
  filter(region == "EU27", period == 2040, variable %in% build_UE_heat[c("Solids")], tgt2030 %in% tgt2030Scen) %>%
  group_by(scenario) %>%    
  summarize(value = sum(value),.groups = "keep")
  ,
  df %>% 
  filter(region == "EU27", period == 2040, variable %in% build_UE_heat, tgt2030 %in% tgt2030Scen) %>%
  group_by(scenario) %>%  
  summarize(total = sum(value)) %>% 
  select(scenario,total)  
  ,
  by = join_by(scenario == scenario)
) %>% 
  mutate(pctg = value/total)

share_of_build_UE_heating <- data.frame(
  FE=c("share_of_electric_and_district_buildings_heating_useful energy_(%)","share_of_liquids_and_gases_buildings_heating_useful_energy_(%)","share_of_solids_buildings_heating_useful energy_(%)"),
  reference_value=c(round(build_UE_heating_data %>% filter(scenario == mainScen) %>% summarize(total = sum(pctg),.groups = "keep") %>% pull(total)*100 ),
                    round(build_UE_liq_gas_data %>% filter(scenario == mainScen) %>% summarize(total = sum(pctg),.groups = "keep") %>% pull(total)*100 ),
                    round(build_UE_solids_data %>% filter(scenario == mainScen) %>% summarize(total = sum(pctg),.groups = "keep") %>% pull(total)*100 )),
  min = c(round(min(build_UE_heating_data %>% pull(pctg) *100)),
          round(min(build_UE_liq_gas_data %>% pull(pctg) *100)),
          round(min(build_UE_solids_data %>% pull(pctg) *100))),
  max = c(round(max(build_UE_heating_data %>% pull(pctg) *100)),
          round(max(build_UE_liq_gas_data %>% pull(pctg) *100)),
          round(max(build_UE_solids_data %>% pull(pctg) *100)))
)

build_FE_data <- left_join(
  df %>% 
  filter(region == "EU27", period == 2040, variable == "FE|Buildings|Heating", tgt2030 %in% tgt2030Scen) %>% 
  select(scenario,value)  
  ,
  df %>% 
  filter(region == "EU27", period == 2020, variable == "FE|Buildings|Heating", tgt2030 %in% tgt2030Scen) %>%
  mutate(base = value) %>% 
  select(scenario,base)  
  ,
  by = join_by(scenario == scenario)
) %>% 
  mutate(pctg = value/base)

share_of_build_FE_in_relation_to_2020 <- data.frame(
  FE=c("share_of_build_FE_in_relation_to_2020_(%)"),
  reference_value=c(round((1-build_FE_data %>% filter(scenario == mainScen) %>% pull(pctg))*100 )),
  min = c(round(min((1-build_FE_data %>% pull(pctg)) *100))),
  max = c(round(max((1-build_FE_data %>% pull(pctg)) *100)))
)

```

```{r, echo=FALSE, fig.width=12, fig.height=5}

Generation %>% 
  kbl() %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"))

share_of_fossil_in_electricity %>% 
  kbl() %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"))

primaryEnergy %>% 
  kbl() %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"))

FE %>% 
  kbl() %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"))

share_of_sustainable_fuel %>% 
  kbl() %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"))

share_of_build_UE_heating %>% 
  kbl() %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"))

share_of_build_FE_in_relation_to_2020 %>% 
  kbl() %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"))

```

---

<!-- Paper Figures -->
```{r, echo=FALSE}
### Initialize plot objects
g <- NULL

```

# Figure 1. Emission results panel

(a) Total GHG emissions (Mt CO2eq/yr)                    (b) Emission reductions (%)

```{r, echo=FALSE}

# all emission reductions were calculated relative to 1990 emissions with LULUCF and international transport (4814 Mt CO2e), except for 2030 emission reduction calculated relative to 1990 emissions with LULUCF and intra-EU aviation only (4713 Mt CO2e).

plotData <- df %>% 
  filter(region == "EU27", period %in% c(2020:2060), tgt2030 %in% tgt2030Scen) %>% 
  calc_addVariable( "`Emissions with LULUCF and with international transport`" = "`Emi|GHG|LULUCF national accounting` + `Emi|CO2|Energy|Demand|Transport|International Bunkers`" , units = "Mt CO2/yr", only.new=T)  %>%
  group_by(model,region,variable,period) %>% 
  filter(tgt2050 != "Npi") %>% 
  mutate(ribbon.max=max(value),
         ribbon.min=min(value))

minValue <- min(plotData$value)

h <- refEmi %>% filter(variable == "Emissions with LULUCF and with international transport", region == "EU27", pollutant == "All greenhouse gases - (CO2 equivalent)") %>% ungroup() %>% select(region,variable,period,value)

emi1990 <- tgt %>% filter(target == "Emissions with LULUCF and with international transport") %>% pull(reference)

g[["GHG emi"]] <- ggplot(plotData, aes(x=period,y=value)) +
    geom_hline(yintercept=0, color = "black", linewidth=2, alpha = 0.5) +
    geom_vline(xintercept=2020, color = "black", linewidth=1.5, alpha = 0.5,linetype="dashed") +
     geom_ribbon(aes(ymax = ribbon.max, ymin = ribbon.min), alpha = 0.5, fill = "#545454") + 
     geom_segment(data=plotData %>% filter(scenario == mainScen, period == 2040), aes(x=2040, xend=2040, y=0, yend=value), linetype="dashed", color="black") +
     geom_line(data=plotData %>% filter(scenario == mainScen), linewidth=1.5) +
     #geom_line(data=plotData, color= "#545454", linewidth=0.2, alpha=0.2, aes(group=scenario)) +
     geom_line(data=h,linewidth=0.7,linetype="dashed") +
     geom_point(data=plotData %>% filter(scenario == mainScen, period == 2040),size=3,shape=21,fill="white",stroke=1.5) +
     theme_doubleColumn() +
#    ggtitle("GHG Emissions") +
    theme(axis.title.y=element_blank()) +
     expand_limits(y=emi1990) +
#     ylab(expression(paste("Mt ", CO[2],"eq/yr"))) +
     scale_y_continuous(breaks = c(0,1000,2000,3000,4000,5000))

plotData <- rbind(
  df %>% 
    filter(region == "EU27", period %in% c(2030:2040), tgt2030 %in% tgt2030Scen) %>%
    calc_addVariable( "`Emissions with LULUCF and with intra-EU aviation`" = "`Emi|GHG|LULUCF national accounting` + `Emi|CO2|Energy|Demand|Transport|International Bunkers|Intra-region`" , units = "Mt CO2/yr", only.new=T) %>% 
    mutate(reduction = round(value / tgt %>% filter(target == "Emissions with LULUCF and with intra-EU aviation") %>% pull(reference) ,2)) %>% 
    select(-variable)
  ,
  df %>% 
    filter(region == "EU27", period %in% c(2045:2050), tgt2030 %in% tgt2030Scen) %>%
    calc_addVariable( "`Emissions with LULUCF and with international transport`" = "`Emi|GHG|LULUCF national accounting` + `Emi|CO2|Energy|Demand|Transport|International Bunkers`" , units = "Mt CO2/yr", only.new=T) %>% 
    mutate(reduction = round(value / tgt %>% filter(target == "Emissions with LULUCF and with international transport") %>% pull(reference) ,2)) %>% 
    select(-variable)
  ) %>%
  filter(tgt2050 != "Npi") %>%
  mutate(x=ifelse(reduction>0,period+1,NA),
         xend=ifelse(reduction>0,period+1,NA),
         y=ifelse(reduction>0,1,NA),
         yend=ifelse(reduction>0,ifelse(reduction<=0,0,reduction),NA),
         bar = ifelse(period==2040,"2040","emi"),
         #label = ifelse(value>0,paste0( "- ", (1-value) * 100, "% (-", (1-intraEURed) * 100 ,"%)"),NA),
         labelY = reduction + (1-reduction)/2,
         labelX=period+1) %>%
  group_by(model,region,period) %>% 
  mutate(max=max(reduction),
         min=min(reduction),
         range=max-min,
         label= paste0(round((1-max) * 100), "% to ", round((1-min) * 100) ,"%"))

g[["target"]] <- ggplot(data=plotData,aes(x=period,y=reduction,fill=bar)) +
  geom_bar(data=plotData %>% filter(scenario == mainScen), stat='identity', show.legend = FALSE, alpha = 0.7, color="black") +
  geom_errorbar(aes(ymin=min, ymax=max), width=1, colour="black", alpha=0.9, linewidth=1.3) +
  geom_hline(yintercept=0, color = "black", linewidth=2, alpha = 0.5) +
  theme_doubleColumn() +
  scale_fill_manual(values = c("emi" = "black", "2040" = "#5a788b")) + 
  guides(y = "none") +
  theme(axis.title.y=element_blank()) +
  geom_segment(data=plotData %>% filter(scenario == mainScen), aes(x=x, y=y, xend=xend, yend=yend), arrow = arrow(length=unit(.5, 'cm')), na.rm=TRUE) +
  geom_text(data=plotData %>% filter(scenario == mainScen), aes(label = label, y=labelY, x=labelX, color=factor(period)), vjust=-1, hjust=0.5, na.rm=TRUE, size = geom.text.single.size, angle = 90) +
  scale_color_manual(values = c("2030" = "black", "2035" = "#777777","2040" = "black", "2045" = "#777777","2050" = "black"), breaks=c("2030", "2035","2040", "2045","2050")) +
  expand_limits(y=minValue/emi1990) +
  annotate("text", x=2052, y=1, label= "1990 emissions", vjust=-1, hjust=1, size = geom.text.single.size) +
  geom_hline(yintercept=1, color = "black", linetype="dashed", linewidth=2, alpha = 0.5) +
  coord_cartesian(clip = 'off') +
  scale_y_continuous(breaks = c(0,1000/emi1990,2000/emi1990,3000/emi1990,4000/emi1990,5000/emi1990)) + 
  theme(legend.position = "none")

g[["GHG Emission reductions"]] <- gridExtra::arrangeGrob(
  g[["GHG emi"]], #+ theme(plot.margin=margin(10,10,30,10)),
  g[["target"]], # + theme(plot.margin=margin(10,10,10,10)),
  layout_matrix = rbind(c(1, 1, 1, 1, 2, 2),
                          c(1, 1, 1, 1, 2, 2))
   )

 ggsave(paste0(outFolder,"/GHG Emission reductions.svg"), g[["GHG Emission reductions"]] , device="svg", width = 12, height = 5, dpi=100, units = "in")
 ggsave(paste0(outFolder,"/GHG Emission reductions.png"), g[["GHG Emission reductions"]] , device="png", width = 12, height = 5, dpi=300, units = "in")

```

```{r, echo=FALSE, fig.width=12, fig.height=5}

grid.newpage()
grid.draw(g[["GHG Emission reductions"]])

```

(c) Breakdown of residual CO2 emissions from energy and industrial processes (% in relation to 2020)

```{r, echo=FALSE}

vars <- list(
  "Industrial\nProcesses" = "Emi|GHG|Industrial Processes",
  "Industry" = "Emi|GHG|Gross|Energy|Demand|Industry",
  "Buildings" ="Emi|GHG|Energy|Demand|Buildings",
  "Transportation" ="Emi|GHG|Energy|Demand|Transport",
  "Bunkers" = "Emi|CO2|Energy|Demand|Transport|International Bunkers",
#  "CDR" = "Emi|GHG|Gross|Energy|Demand|CDR",
#  "AFOFI" = "Emissions|CO2|Energy|Demand|AFOFI",
#  "Other Sector" ="Emissions|CO2|Energy|Demand|Other Sector",
  "Hydrogen" ="Emi|CO2|Gross|Energy|Supply|Hydrogen",
  "Liquids"="Emi|CO2|Gross|Energy|Supply|Liquids",
  "Gases"="Emi|CO2|Gross|Energy|Supply|Gases",
  "Solids"="Emi|CO2|Gross|Energy|Supply|Solids",
#  "Other"="Emissions|CO2|Energy|Supply|Other",
  "Heat"="Emi|CO2|Gross|Energy|Supply|Heat",
  "Electricity"="Emi|GHG|Gross|Energy|Supply|Electricity"
)

# "Emi|GHG|Agriculture",
#   "Emi|GHG|Land-Use Change",
#   "Emi|GHG|Waste",
#   "Emi|CO2|CDR|BECCS",
#   "Emi|CO2|CDR|Industry CCS|Synthetic Fuels",
#   "Emi|CO2|CDR|DACCS",
#   "Emi|CO2|CDR|EW"

plotColor <- c(
  "Electricity"= "#ffb200",
  "Heat"= "#cc0000",
  "Gases"= "#999959",
  "Liquids"= "#0000cc",
  "Solids"= "#0c0c0c",
  "Other" = "#79a6d2",
  "Hydrogen" = "#5ed5b0",
  "Transportation" = "#a1dd70",
  "Bunkers" = "#ffc0cb",
  "Buildings" = "#5edfff",
  "Other Sector" = "#999959",
  "AFOFI" = "#005900",
  "Industry" = "#5f6769",
  "Industrial\nProcesses" = "#5c5c5c",
  "Other" = "#5b0d8f",
  "Waste" = "#A52A2A",
  "AFOLU" = "#4DAF4A",
  "Agriculture" = "#357933",
  "LULUCF" = "#6ec16b"
)

plotData <- df %>%
  filter(region == "EU27",
         period %in% c(2020,2030,2040),
         tgt2030 %in% tgt2030Scen,
         variable %in% unlist(vars),
         scenario == mainScen)

labels <- setNames(names(vars),vars[names(vars)])

plotData$labels <- labels[as.character(plotData$variable)]

plotData[plotData$value <= 0,]$value <- 0

data <-
  plotData %>%
  mutate(complement = FALSE) %>%
  group_by(variable) %>%
  mutate(pct = paste0(round(value / value[period==2020],2)*100, "%")) %>%
  bind_rows(
    plotData %>%
      mutate(complement = TRUE) %>%
      group_by(variable) %>%
      mutate(value = value[period==2020] - value) %>%
      mutate(labels=NA)
  ) %>%
  mutate(ext_Var = ifelse(complement, paste0(as.character(variable),"_"), as.character(variable))) %>%
  arrange(period,ext_Var) %>%
  arrange(match(variable, c("25pct",vars))) %>%
  group_by(period) %>%
  mutate(ymid = as.numeric(sub("\\D+", "", period)),
         ymax = ymid + 4, ymin = ymid - 4,
         xmin = c(0, head(cumsum(value), -1)),
         xmax = cumsum(value),
         xmid = (xmax + xmin) / 2,
         )

data[data$period!=2020,]$labels <- NA
data[data$period != 2040,]$pct <- NA
data[data$variable == "25pct",]$pct <- NA

max <- data %>%
  filter(variable == tail(vars,1), complement == TRUE) %>% 
  group_by(period)

data$labelsPosY <- data$ymid
data[data$labels %in% c("Solids"),]$labelsPosY <- 2015
data[data$labels %in% c("Bunkers","Industrial Processes","Liquids"),]$labelsPosY <- 2016
data[data$labels %in% c("Hydrogen"),]$labelsPosY <- 2010

data$pctPosY <- data$ymid
#data[data$Variable %in% c("Emi|GHG|Industrial Processes","Emi|CO2|Gross|Energy|Supply|Gases","Emi|GHG|Gross|Energy|Supply|Electricity"),]$pctPosY <- 2047
data[data$variable %in% c("Emi|CO2|Gross|Energy|Supply|Solids","Emi|CO2|Gross|Energy|Supply|Heat","Emi|CO2|Gross|Energy|Supply|Hydrogen"),]$pctPosY <- 2100
data[data$variable %in% c("Emi|CO2|Gross|Energy|Supply|Liquids"),]$pctPosY <- 2043
data$pctPosX <- data$xmid
#data[data$variable %in% c("Emi|GHG|Industrial Processes"),]$pctPosX <- data[data$variable %in% c("Emi|GHG|Industrial Processes"),]$pctPosX - 100
#data[data$variable %in% c("Emi|GHG|Gross|Energy|Demand|Industry"),]$pctPosX <- data[data$variable %in% c("Emi|GHG|Gross|Energy|Demand|Industry"),]$pctPosX + 100

#data$ext_Var <- as.character(data$ext_Var)

#remove labels for small values
data <- data %>% mutate(pct = ifelse(variable %in% vars[c("Hydrogen","Solids","Heat")], NA, pct))

g[["2040 Emissions - Energy and Industrial Processes - horizontal bar"]] <- data %>%
  ggplot(aes(xmid, ymid, fill = ext_Var)) +
  geom_rect(aes(xmin = xmin, xmax = xmax, ymin = ymin, ymax = ymax,
                alpha = factor(period), color = variable)) +
  #geom_text(data=max, aes(label = period, y=period, x=xmax), vjust=0.5, hjust=-0.5) + 
  geom_text(aes(y = 2015, label = labels, group = labels, hjust=1, angle = 90), vjust=0.5, hjust=1.1, na.rm=TRUE, size = geom.text.single.size, angle = 90) +
  #geom_label_repel(aes(y = 2014, label = labels, group = labels), vjust=0.5, hjust=1, na.rm=TRUE, size = geom.text.single.size, angle = 90) +
  #geomtextpath::geom_textpath(aes(y = 2014, label = labels, group = labels, vjust=0.5, hjust=1, angle = 90),text_only = TRUE, na.rm=TRUE, size = geom.text.double.size) +
  geom_text(aes(x=pctPosX , y = 2045, label = pct), vjust=0.5, hjust=-0.1, na.rm=TRUE, size = geom.text.single.size, angle = 90) +
  #geomtextpath::geom_textpath(aes(x=pctPosX , y = 2046, label = pct),text_only = TRUE, na.rm=TRUE, size = geom.text.double.size) +
  #geom_text(aes(y = ymid + 3, label = labels, group = labels)) +
  scale_fill_manual(values = c(setNames(plotColor[names(vars)],vars),setNames(rep("transparent",length(vars)),paste0(vars,"_"))), labels = setNames(names(vars), vars)) +
  scale_alpha_manual(values = c(0.2, 0.3, 1)) +
  scale_color_manual(values = c(setNames(plotColor[names(vars)],vars)), labels = setNames(names(vars), vars)) +
  #scale_fill_manual(values = c("#cd9900", "#00817e")) +
  #scale_colour_manual(values = c("#cd9900", "#00817e")) +
  #scale_y_continuous(limits = c(1990, 2045)) +
  #scale_y_reverse(limits = c(2065, 2013)) +
  theme_doubleColumn() +
  theme(legend.position = "none") +
  scale_y_continuous(breaks=c(2020,2030,2040), minor_breaks = NULL, limits=c(2008,2047)) +
  theme(axis.title.x=element_blank(),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank(),
        axis.title.y=element_blank()) +
  scale_x_continuous(breaks = NULL) +
  coord_cartesian(clip="off") +
  theme(plot.margin = unit(c(1,0,2,0), "cm"))

ggsave(paste0(outFolder,"/2040 Emissions - Energy and Industrial Processes - horizontal bar.svg"), g[["2040 Emissions - Energy and Industrial Processes - horizontal bar"]] , device="svg", width = 12, height = 3, dpi=100, units = "in")
ggsave(paste0(outFolder,"/2040 Emissions - Energy and Industrial Processes - horizontal bar.png"), g[["2040 Emissions - Energy and Industrial Processes - horizontal bar"]] , device="png", width = 12, height = 3, dpi=300, units = "in")

```

```{r, echo=FALSE, fig.width=12, fig.height=3}

  g[["2040 Emissions - Energy and Industrial Processes - horizontal bar"]]

```

Figure 1. Emission results panel. Reference scenario: Net-zero by 2050, 55% reduction by 2030, REMIND future final energy projections, biomass availability limited to 7.5 EJ/yr and default CCS assumption.

---

# Figure 2. Electricity panel

:::: {style="display: flex;"}

::: {}

(a) Electricity Generation (%)

```{r, echo=FALSE}

vars <- c(
  "Net Imports" = "SE|Electricity|Net Imports",
  "Coal w/ CC" = "SE|Electricity|Coal|w/ CC",
  "Coal w/o CC" = "SE|Electricity|Coal|w/o CC",
  "Oil" = "SE|Electricity|Oil",
  "Gas w/ CC" = "SE|Electricity|Gas|w/ CC",
  "Gas w/o CC" = "SE|Electricity|Gas|w/o CC",
  "Hydrogen" = "SE|Electricity|Hydrogen",
  "Nuclear" = "SE|Electricity|Nuclear",
  "Geothermal" = "SE|Electricity|Geothermal",
  "Hydro" = "SE|Electricity|Hydro",
  "Biomass w/ CC" = "SE|Electricity|Biomass|w/ CC",
  "Biomass w/o CC" = "SE|Electricity|Biomass|w/o CC",
  "Solar CSP" = "SE|Electricity|Solar|CSP",
  "Solar PV" = "SE|Electricity|Solar|PV",
  "Wind Onshore" = "SE|Electricity|Wind|Onshore",
  "Wind Offshore" = "SE|Electricity|Wind|Offshore")

inv_vars <- setNames(names(vars),vars)

color <- c(
  "Net Imports"   = "#472b62",
  "Coal w/ CC"    = "#b2b2b2",
  "Coal w/o CC"   = "#0c0c0c",
  "Oil"           = "#b30000",
  "Gas w/ CC"     = "#b79e38",
  "Gas w/o CC"    = "#999959",
  "Hydrogen"      = "#5ed5b0",
  "Nuclear"       = "#ff33ff",
  "Geothermal"    = "#e51900",
  "Hydro"         = "#191999",
  "Biomass w/ CC" = "#33ff00",
  "Biomass w/o CC"= "#005900",
  "Solar CSP"     = "#ffcc00",
  "Solar PV"      = "#ffe600",
  "Wind Onshore"  = "#193F7F",
  "Wind Offshore" = "#337fff")

plotData <- df %>% 
  filter(region == "EU27", period >= 2020, period <= 2060, scenario  == mainScen, variable %in% unlist(vars)) %>% 
  mutate(variable = factor(variable, levels=vars)) %>% 
  arrange(factor(variable, levels = rev(vars))) %>%
  group_by(model,scenario,region,period) %>%
  mutate(cumValue = cumsum(value),
         n = sum(value),
         percentage = value / n,
         cumPercentage = cumValue / n)

plotData2040 <- plotData[plotData$period == 2040,]
#plotData$value <- plotData$value / plotData[plotData$period == "2020",]$value
#plotData$period <- as.character(plotData$period)
#plotData$period <- factor(plotData$period , levels = rev(c("2020","2030","2040","2050")))

plotData2040$xPos <- 2042.5
#plotData2040[plotData2040$variable %in% c("SE|Electricity|Hydro"),]$xPos <- 2028.5
#plotData2040[plotData2040$variable %in% c("SE|Electricity|Biomass|w/o CC"),]$xPos <- 2020
plotData2040[plotData2040$variable %in% c("SE|Electricity|Nuclear","SE|Electricity|Biomass|w/o CC"),]$xPos <- 2035

plotDataRenewables <- plotData %>%
  filter(variable %in% c("SE|Electricity|Geothermal","SE|Electricity|Hydro","SE|Electricity|Biomass|w/ CC","SE|Electricity|Biomass|w/o CC","SE|Electricity|Solar|CSP","SE|Electricity|Solar|PV","SE|Electricity|Wind|Onshore","SE|Electricity|Wind|Offshore","SE|Electricity|Nuclear")) %>%
  group_by(model,scenario,region,period) %>%
  summarize(percentage = sum(percentage),.groups = "keep")

g[["Electricity Generation"]] <- 
  ggplot(plotData,aes(x=period,y=percentage)) +
  geom_area(aes(fill=variable),alpha=0.4) +
  geom_line(data=plotDataRenewables, linetype = "dashed", linewidth = 1.5, color="#666666") +
  geom_bar(data=plotData2040,aes(fill=variable),stat='identity', show.legend = FALSE, alpha = 0.7, color="white", width = 3, linewidth = 1.5) +
  #ggrepel::geom_text_repel(data=plotData2040[plotData2040$percentage >= 0.01,],aes(label = paste0(round(percentage,2)*100,"%"), y=cumPercentage-percentage/2, x = 2043.5), vjust=0.5, hjust=0.5, na.rm=TRUE, color = "black", size = 12) +
  #geom_text(data=plotData2040[plotData2040$percentage >= 0.01,],aes(label = paste0(inv_vars[variable], " (",round(percentage,2)*100,"%)"), y=cumPercentage-percentage/2, x = xPos), vjust=0.5, hjust=0, na.rm=TRUE, color = "black", size = 12) +
  geom_text(data=plotData2040[plotData2040$percentage >= 0.02,],aes(label = paste0(round(percentage,2)*100,"%", ifelse(inv_vars[variable] %in% c("Solar PV","Wind Onshore","Wind Offshore"),paste0(" - ", inv_vars[variable]), "")), y=cumPercentage-percentage/2, x = xPos), vjust=0.5, hjust=0, na.rm=TRUE, color = "black", size=geom.text.double.size) +
  theme_doubleColumn()  +
  scale_y_continuous(minor_breaks = seq(0.2, 0.8, 0.2), breaks=c(0,1),labels = scales::percent) +
  scale_x_continuous(breaks=c(2020,2040,2060)) +
  scale_fill_manual(values = setNames(color[names(vars)],vars), labels = setNames(names(vars), vars)) +
  theme(axis.title.y=element_blank()) +
  guides(fill=guide_legend(nrow=8,override.aes = list(alpha=1), byrow = TRUE)) +
  #annotate(geom = "text", x = 2022, y = plotDataRenewables[plotDataRenewables$period == 2020,]$percentage + 0.06, label = "renewables", color = "#888888",angle = 28, hjust=0, vjust=-0.5, size = 12) +
  #annotate(geom = "text", x = 2023, y = plotDataRenewables[plotDataRenewables$period == 2020,]$percentage - 0.06, label = "and nuclear", color = "#888888",angle = 28, hjust=0, vjust=-0.5, size = 12) +
  annotate(geom = "text", x = 2020, y = plotDataRenewables[plotDataRenewables$period == 2020,]$percentage, label = "renewables\nand nuclear", color = "#666666",angle = 26, hjust=0, vjust=1, size=geom.text.double.size) +
  #theme(axis.text.y=element_blank()) +
  theme(legend.key.height = unit(0.8, "cm"),
    legend.key.width = unit(0.8,"cm")) +
  theme(legend.spacing.y = unit(0.1, 'cm')) +
  theme(legend.margin=margin(l = -2, unit='cm'))

ggsave(paste0(outFolder,"/Electricity Generation.svg"), g[["Electricity Generation"]] , device="svg", width = 6, height = 10, dpi=100, units = "in")
ggsave(paste0(outFolder,"/Electricity Generation.png"), g[["Electricity Generation"]] , device="png", width = 6, height = 10, dpi=300, units = "in")

```

```{r, echo=FALSE, fig.width=6, fig.height=10}

g[["Electricity Generation"]]

```

:::

::: {}

(b) Solar Photovoltaics capacity (GW)

```{r, echo=FALSE}

# Solar photovoltaics roll out
plotData <- df %>% 
  filter(region == "EU27", period %in% c(2020:2060), tgt2030 %in% tgt2030Scen, variable == "Cap|Electricity|Solar|PV") %>% 
  filter(tgt2050 != "Npi") %>% 
  group_by(model,region,variable,period) %>% 
  mutate(ribbon.max=max(value),
         ribbon.min=min(value))

target <- data.frame(target=c("REPowerEU","REPowerEU"),period=c(2025,2030),value=c(320,600))

#value <- data.frame(period=c(2025,2030,2040),value=c(plotData[plotData$period==2025,]$value,plotData[plotData$period==2030,]$value,plotData[plotData$period==2040,]$value))

# historical data
h <- hist %>% filter(model == "IRENA", region == "EU27", variable == "Cap|Electricity|Solar|PV", period >= 2005) %>% drop_na(value)

g[["Solar Photovoltaics capacity"]] <- ggplot(plotData,aes(x=period,y=value)) +
    geom_hline(yintercept=0, color = "black", linewidth=2, alpha = 0.5) +
    geom_vline(xintercept=2020, color = "black", linewidth=1.5, alpha = 0.5,linetype="dashed") +
    geom_segment(aes(x=2040, y=0, xend=2040, yend=plotData %>% filter(scenario == mainScen, period == 2040) %>% pull(value)), linetype="dashed", color="black") +
    geom_ribbon(aes(ymax = ribbon.max, ymin = ribbon.min), alpha = 0.5, fill = "#545454") + 
    geom_line(data=plotData %>% filter(scenario == mainScen),linewidth=1.5) +
    geom_line(data=h,linewidth=1,linetype="dashed") +
    geom_point(data=plotData %>% filter(scenario == mainScen, period %in% c(2030,2040)),size=3,shape=21,fill="white",stroke=1.5) +
    geom_point(data=target, size=3, shape=4, stroke = 1.5) +
    geom_label_repel(data=target, aes(label = paste0(round(value),"GW - ", target)),box.padding = 0.1,
                  #segment.curvature = -0.1,
                  segment.square = TRUE,
                  point.padding = 0.6,
                  nudge_x = 5,
                  #nudge_y = -0.02,
                  #force = 0.5,
                  hjust = 0,
                  direction="y", size = geom.text.double.size*0.9, xlim = c(2020, 2070)) +
  geom_label_repel(data=plotData %>% filter(scenario == mainScen, period == 2030), aes(label = paste0(round(value),"GW")),box.padding = 0.1,
                  point.padding = 1,
                  nudge_x = -2,
                  nudge_y = 500,
                  hjust = 1, direction="y",
                  size = geom.text.double.size) +
  geom_label_repel(data=plotData %>% filter(scenario == mainScen, period == 2040), aes(label = paste0(round(value),"GW")),box.padding = 0.1,
                  point.padding = 1,
                  nudge_x = -2,
                  nudge_y =500,
                  #hjust = 0, 
                  direction="y",
                  size = geom.text.double.size) +
    theme_doubleColumn() + 
    theme(axis.title.y = element_blank())+
    coord_cartesian(clip = 'off') +
    expand_limits(y = 3000)

ggsave(paste0(outFolder,"/Solar Photovoltaics capacity.svg"), g[["Solar Photovoltaics capacity"]] , device="svg", width = 6, height = 4, dpi=100, units = "in")
ggsave(paste0(outFolder,"/Solar Photovoltaics capacity.png"), g[["Solar Photovoltaics capacity"]] , device="png", width = 6, height = 4, dpi=300, units = "in")

```

```{r, echo=FALSE, fig.width=6, fig.height=4}

g[["Solar Photovoltaics capacity"]]

```

(c) Electricity share in final energy (%)

```{r, echo=FALSE}

# Electricity share
vars <- c("Total" = "Final Energy|Electricity Share",
          "Transport"="Final Energy|Transport|Electricity Share",
          "Industry"="Final Energy|Industry|Electricity Share",
          "Buildings"="Final Energy|Buildings|Electricity Share")

plotData <- df %>% 
  filter(region == "EU27", period %in% c(2020:2060), tgt2030 %in% tgt2030Scen, variable %in% unlist(vars)) %>% 
  group_by(model,region,variable,period) %>% 
  filter(tgt2050 != "Npi") %>% 
  mutate(ribbon.max=max(value),
         ribbon.min=min(value))

h_Orig <- hist[hist$model == "IEA" & hist$region == "EU27" & hist$variable %in% c("FE","FE|Electricity","FE|Transport","FE|Transport|Electricity","FE|Industry", "FE|Industry|Electricity","FE|Buildings","FE|Buildings|Electricity"),] %>% drop_na(value)

h <- rbind(
  h_Orig %>% calc_addVariable("`Final Energy|Electricity Share`" = "`FE|Electricity` / `FE`", units = "%", only.new=T),
  h_Orig %>% calc_addVariable("`Final Energy|Transport|Electricity Share`" = "`FE|Transport|Electricity` / `FE|Transport`", units = "%", only.new=T),
  h_Orig %>% calc_addVariable("`Final Energy|Industry|Electricity Share`"  = "`FE|Industry|Electricity`  / `FE|Industry`", units = "%", only.new=T),
  h_Orig %>% calc_addVariable("`Final Energy|Buildings|Electricity Share`" = "`FE|Buildings|Electricity` / `FE|Buildings`", units = "%", only.new=T))

g[["Electricity Share"]] <- ggplot(plotData,aes(x=period,y=value)) +
    geom_hline(yintercept=0, color = "black", linewidth=2, alpha = 0.5) +
    geom_vline(xintercept=2020, color = "black", linewidth=1.5, alpha = 0.5,linetype="dashed") +
    geom_segment(aes(x=2040, y=0, xend=2040, yend=max(plotData %>% filter(scenario == mainScen, period == 2040) %>% pull(value))), linetype="dashed", color="black") +
    geom_ribbon(aes(ymax = ribbon.max, ymin = ribbon.min, fill=variable), alpha = 0.3) + 
    geom_line(data=plotData %>% filter(scenario == mainScen), linewidth=1.5,aes(color=variable)) +
    #geom_line(data=h,linewidth=1,linetype="dashed") +
    geom_point(data=plotData %>% filter(scenario == mainScen, period == 2040),size=3,shape=21,fill="white",stroke=1.5) +
    geom_line(data=h %>% filter(period >= 2005),linetype="dashed", linewidth = 1, aes(color=variable)) +
    geom_label_repel(data= plotData %>% filter(scenario == mainScen, period == 2040),
                         aes(
                           label = paste0(round(value*100),"%")),
                      box.padding = 3,
                      #segment.curvature = -0.1,
                      segment.square = TRUE,
                      point.padding = 0.6,
                      nudge_x = 5,
                      nudge_y = c(0.55,0.27,-0.05,-0.6),
                      #force = 0.5,
                      #hjust = 0,
                      direction="y"
                      , size = geom.text.double.size*0.9
                      #, xlim = c(2020, 2070)
                      ) +
    scale_colour_manual(values = c("Final Energy|Electricity Share"="black","Final Energy|Transport|Electricity Share"="#a1dd70","Final Energy|Industry|Electricity Share"="#5f6769","Final Energy|Buildings|Electricity Share"="#5edfff"), labels = setNames(names(vars), vars)) +
    scale_fill_manual(values = c("Final Energy|Electricity Share"="black","Final Energy|Transport|Electricity Share"="#a1dd70","Final Energy|Industry|Electricity Share"="#5f6769","Final Energy|Buildings|Electricity Share"="#5edfff"), labels = setNames(names(vars), vars)) +
    theme_doubleColumn() + 
    scale_y_continuous(breaks = seq(0, 1, 0.2),labels = scales::percent) +
    expand_limits(y=c(0,1)) +
    guides(color=guide_legend(nrow=2)) +
  theme(axis.title.y = element_blank())

ggsave(paste0(outFolder,"/Electricity Share.svg"), g[["Electricity Share"]] , device="svg", width = 6, height = 6, dpi=100, units = "in")
ggsave(paste0(outFolder,"/Electricity Share.png"), g[["Electricity Share"]] , device="png", width = 6, height = 6, dpi=300, units = "in")

```

```{r, echo=FALSE, fig.width=6, fig.height=6}

g[["Electricity Share"]]

```

:::

::::

---

# Figure 3. Energy supply panel

:::: {style="display: flex;"}

::: {}

(a) Primary energy (EJ/yr)

```{r, echo=FALSE}

# Primary Energy
vars <- c("Biomass" = "PE|Biomass",
          "Coal"="PE|Coal",
          "Gas"="PE|Gas",
          "Oil"="PE|Oil")

plotData <- df %>% 
  filter(region == "EU27", period %in% c(2020:2060), tgt2030 %in% tgt2030Scen, variable %in% unlist(vars)) %>% 
  group_by(model,region,variable,period) %>% 
  filter(tgt2050 != "Npi") %>% 
  mutate(ribbon.max=max(value),
         ribbon.min=min(value))

# historical data
h <- hist %>% filter(model == "IEA", region == "EU27", variable %in% vars, period >= 2005, period <= 2020) %>% drop_na(value) %>% filter(value != 0)

g[["Primary Energy"]] <- ggplot(plotData,aes(x=period,y=value)) +
    geom_hline(yintercept=0, color = "black", linewidth=2, alpha = 0.5) +
    geom_vline(xintercept=2020, color = "black", linewidth=1.5, alpha = 0.5,linetype="dashed") +
    geom_segment(aes(x=2040, y=0, xend=2040, yend=max(plotData %>% filter(scenario == mainScen, period == 2040) %>% pull(value))), linetype="dashed", color="black") +
    geom_line(data=plotData %>% filter(scenario == mainScen), aes(color=variable), linewidth=1.5) +
    geom_line(data=h,aes(color=variable),linewidth=1,linetype="dashed") +
    geom_ribbon(aes(ymax = ribbon.max, ymin = ribbon.min, fill=variable), alpha = 0.3) + 
    #geom_point(data=plotData[plotData$period %in% c(2020,2030,2040),],size=4,shape=21,fill="white") +
    geom_point(data=plotData %>% filter(scenario == mainScen, period == 2040),aes(color=variable),size=3,shape=21,fill="white",stroke=1.5) +
    #annotate(geom = "text", x = 2020, y = plotDataRenewables[plotDataRenewables$period == 2020,]$percentage, label = "renewables\nand nuclear", color = "#666666",angle = 26, hjust=0, vjust=1, size=geom.text.double.size) +
    scale_color_manual(values = c(setNames(options$aesthetics$legend$color[names(vars)],vars)), labels = setNames(names(vars), vars)) +
    scale_fill_manual(values = c(setNames(options$aesthetics$legend$color[names(vars)],vars)), labels = setNames(names(vars), vars)) +
    theme_doubleColumn() + 
    expand_limits(y=c(0,1)) +
    guides(color=guide_legend(nrow=2)) +
    theme(axis.title.y = element_blank()) +
    expand_limits(y = 30)
    #guides(fill = FALSE)

ggsave(paste0(outFolder,"/Primary Energy.svg"), g[["Primary Energy"]] , device="svg", width = 6, height = 6, dpi=100, units = "in")
ggsave(paste0(outFolder,"/Primary Energy.png"), g[["Primary Energy"]] , device="png", width = 6, height = 6, dpi=300, units = "in")

```

```{r, echo=FALSE, fig.width=6, fig.height=6}

g[["Primary Energy"]]

```

(c) Natural gas and syngas final energy demand (EJ/yr)

```{r, echo=FALSE}

vars <- c(
  "Hydrogen"="FE|Hydrogen",
  "Gases"="FE|Gases"
)

plotData <- df %>% 
  filter(region == "EU27", period %in% c(2020,2030,2040,2050), tgt2030 %in% tgt2030Scen, variable %in% unlist(vars)) %>% 
  mutate(variable = factor(variable, levels=vars))  %>% 
  filter(tgt2050 != "Npi")  %>%
  group_by(model,scenario,region,period) %>% 
  mutate(cumsum = cumsum(value),
         sum = sum(value)) %>% 
  group_by(model,region,variable,period) %>% 
  mutate(ribbon.max=max(cumsum),
         ribbon.min=min(cumsum)) %>% 
  group_by(model,region,period) %>% 
  mutate(ribbon.max.total=max(sum),
         ribbon.min.total=min(sum))

g[["FE Gases"]] <- 
  ggplot(plotData,aes(x=period,y=value)) +
  geom_col(data=plotData %>% filter(scenario == mainScen), aes(fill = variable)) +
  #geom_errorbar(data=plotData %>% filter(period != 2020), aes(ymin=ribbon.min, ymax=ribbon.max, group=variable), width=8, colour="black", alpha=0.9, size=1,position = position_dodge(width=2)) +
  geom_errorbar(data=plotData %>% filter(period != 2020, variable == vars[1], scenario == mainScen), aes(ymin=ribbon.min.total, ymax=ribbon.max.total), width=4, colour="black", alpha=0.9, linewidth=1,position = position_dodge(width=2)) +
  # scale_y_continuous(labels = scales::percent, breaks = c(0,1,plotData[plotData$period == "2030",]$value,plotData[plotData$period == "2040",]$value,plotData[plotData$period == "2050",]$value)) +
  coord_flip() +
  theme_doubleColumn() +
  theme(panel.grid.minor.x = element_blank(),
        panel.grid.major.y = element_blank()) +
  #guides(y = "none") +
  theme(axis.title.y=element_blank()) +
  #theme(legend.position = "none") #+
  scale_fill_manual(values = setNames(options$aesthetics$legend$color[names(vars)],vars), labels = setNames(names(vars), vars), guide = guide_legend(reverse = TRUE, nrow = 1)) +
  theme(legend.text=element_text(size=theme.double.text.size)) +
  theme(legend.key.height = unit(0.7, "cm"),
    legend.key.width = unit(0.7,"cm")) +
  scale_y_continuous(breaks = c(0,round(sum(plotData %>% filter(scenario == mainScen, period == 2020) %>% pull(value)),1),round(sum(plotData %>% filter(scenario == mainScen, period == 2030) %>% pull(value)),1),round(sum(plotData %>% filter(scenario == mainScen, period == 2040) %>% pull(value)),1))) #+
  #scale_x_reverse()

ggsave(paste0(outFolder,"/FE Gases.svg"), g[["FE Gases"]] , device="svg", width = 6, height = 4, dpi=100, units = "in")
ggsave(paste0(outFolder,"/FE Gases.png"), g[["FE Gases"]] , device="png", width = 6, height = 4, dpi=300, units = "in")

```

```{r, echo=FALSE, fig.width=6, fig.height=4}

g[["FE Gases"]]

```

:::

::: {}

(b) Fossil trade (EJ/yr)

```{r, echo=FALSE}

vars <- list(
 "Coal" = c("Trade|Imports|Coal","Trade|Exports|Coal"),
 "Oil" = c("Trade|Imports|Oil","Trade|Exports|Oil"),
 "Gas" = c("Trade|Imports|Gas","Trade|Exports|Gas"),
 "Biomass" = c("Trade|Imports|Biomass","Trade|Exports|Biomass")
)

# plotData <- plotData %>% 
#   arrange(factor(variable, levels = rev(vars))) %>%
#   group_by(model,scenario,region,period) %>%
#   mutate(cumValue = cumsum(value),
#          n = sum(value),
#          percentage = value / n,
#          cumPercentage = cumValue / n)

plotData <- df %>% 
  filter(region == "EU27", period  %in% c(2020,2030,2040,2050), scenario == mainScen, variable %in% unlist(vars))

carrier <- c()
for(c in names(vars)){
  for(var in vars[[c]]){
    v <- c
    names(v) <- var
    carrier <- c(carrier,v)
  }
}

plotData <- plotData %>%
  mutate(type = ifelse(grepl("Imports",variable),"Imports","Exports"),
         carrier = carrier[as.character(variable)])

plotData[plotData$type == "Exports",]$value <- -plotData[plotData$type == "Exports",]$value

plotData$carrier <- factor(plotData$carrier, levels=c("Coal","Oil","Gas","Biomass"))

plotDataTotal <- plotData %>%
  group_by(model,scenario,region,unit,period,carrier) %>%
  summarize(value=sum(value),.groups = "keep") %>%
  mutate(type="EU-27 Net-trade")

plotColor <- c(
  "Imports"= "#e56b6f",
  "Exports"= "#355070"
)

# g[["security of supply"]] <- ggplot(plotData,aes(x=period,y=value)) +
#   geom_bar(aes(fill=type),stat='identity') +
#   geom_errorbar(data  = plotDataTotal, aes(y=value, ymax=value, ymin=value, color=type), size=1) +
#   facet_wrap(~carrier,nrow=1) +
#    theme_doubleColumn() +
#    theme(axis.title.y=element_blank()) +
#    scale_fill_manual(values = plotColor, guide = guide_legend(nrow = 1)) + #, guide = guide_legend(nrow = 2) setNames(options$aesthetics$legend$color[names(vars)],vars)
#   scale_color_manual(values = c("Total"="black")) +
#    scale_x_continuous(breaks = c(2020,2030,2040,2050)) + 
#   #theme(
#   #  strip.background = element_blank(),
#   #  strip.text.x = element_blank()) +
#   theme(legend.text=element_text(size=theme.single.text.size)) +
#   theme(legend.key.height = unit(0.7, "cm"),
#     legend.key.width = unit(0.7,"cm")) + 
#   theme(axis.text = element_text(size=theme.single.text.size)) #+                                                              
#   #theme(strip.text.x = element_text(size=0.8*theme.single.text.size)) +
#   #theme(strip.clip = "off")
# 
# 
#  ggsave(paste0(outFolder,"/security of supply.svg"), g[["security of supply"]] , device="svg", width = 6, height = 5, dpi=100, units = "in")
#  ggsave(paste0(outFolder,"/security of supply.png"), g[["security of supply"]] , device="png", width = 6, height = 5, dpi=300, units = "in")

g[["security of supply - fossil"]] <- ggplot(plotData %>% filter(period <= 2050, carrier != "Biomass"),aes(x=period,y=value)) +
  geom_bar(aes(fill=type),stat='identity', alpha = 0.7) +
  geom_errorbar(data  = plotDataTotal %>% filter(period <= 2050, carrier != "Biomass"), aes(y=value, ymax=value, ymin=value, color=type), linewidth=1.5) +
  facet_wrap(~carrier,nrow=1) +
  theme_doubleColumn()  +
  theme(axis.title.y=element_blank()) +
  scale_fill_manual(values = plotColor, guide = guide_legend(ncol = 1, order = 2)) + #, guide = guide_legend(nrow = 2) setNames(options$aesthetics$legend$color[names(vars)],vars)
  scale_color_manual(values = c("EU-27 Net-trade"="black"), guide = guide_legend(order = 1)) +
  scale_x_continuous(breaks = c(2020,2030,2040,2050)) + 
  #theme(
  #  strip.background = element_blank(),
  #  strip.text.x = element_blank()) +
  theme(legend.text=element_text(size=theme.single.text.size)) +
  theme(legend.key.height = unit(0.7, "cm"),
    legend.key.width = unit(1.5,"cm")) #+ 
  #theme(axis.text = element_text(size=theme.single.text.size)) #+                                                              
  #theme(strip.text.x = element_text(size=0.8*theme.single.text.size)) +
  #theme(strip.clip = "off")

ggsave(paste0(outFolder,"/security of supply - fossil.svg"), g[["security of supply - fossil"]] , device="svg", width = 6, height = 6, dpi=100, units = "in")
ggsave(paste0(outFolder,"/security of supply - fossil.png"), g[["security of supply - fossil"]] , device="png", width = 6, height = 6, dpi=300, units = "in")

```

```{r, echo=FALSE, fig.width=6, fig.height=5}

  g[["security of supply - fossil"]]

```

(d) Hydrogen production (EJ/yr)

```{r, echo=FALSE}

vars <- c(
  "Biomass w/ CC" = "SE|Hydrogen|Biomass|w/ CC",
  "Biomass w/o CC" = "SE|Hydrogen|Biomass|w/o CC",
  "Coal w/ CC" = "SE|Hydrogen|Coal|w/ CC",
  "Coal w/o CC" = "SE|Hydrogen|Coal|w/o CC",
  "Gas w/ CC" = "SE|Hydrogen|Gas|w/ CC",
  "Gas w/o CC" = "SE|Hydrogen|Gas|w/o CC",
  "Electrolysis" = "SE|Hydrogen|Electricity",
  #"VRE Storage Electrolysis" = "SE|Hydrogen|Electricity|VRE Storage Electrolysis",
  #"Standard Electrolysis" = "SE|Hydrogen|Electricity|Standard Electrolysis",
  "Net Imports" = "SE|Hydrogen|Net Imports")

#1EJ <- 7 million tons #https://hydrogencouncil.com/wp-content/uploads/2017/11/Hydrogen-Scaling-up_Hydrogen-Council_2017.compressed.pdf
#1EJ <- 8 million tons adrian

color <- setNames(options$aesthetics$legend$color[names(vars)],vars)
color["SE|Hydrogen|Biomass|w/ CC"] <- color["SE|Hydrogen|Biomass|w/o CC"]

target <- data.frame(target=c("REPowerEU","REPowerEU"),period=c(2030,2030),value=c(10,20)) %>%
  mutate(value = value/8)

plotData <- df %>% 
  filter(region == "EU27", period %in% c(2020,2030,2040,2050), tgt2030 %in% tgt2030Scen, variable %in% unlist(vars)) %>% 
  mutate(#value = value*8, #1EJ <- 8 million tons adrian
    variable = factor(variable, levels=vars))  %>% 
  filter(tgt2050 != "Npi")  %>%
  group_by(model,scenario,region,period) %>% 
  mutate(cumsum = cumsum(value),
         sum = sum(value)) %>% 
  group_by(model,region,variable,period) %>% 
  mutate(ribbon.max=max(cumsum),
         ribbon.min=min(cumsum)) %>% 
  group_by(model,region,period) %>% 
  mutate(ribbon.max.total=max(sum),
         ribbon.min.total=min(sum))

g[["Hydrogen"]] <- 
  ggplot(plotData,aes(x=period,y=value)) +
  geom_hline(yintercept= target[1,]$value, color = "black", linewidth=1, alpha = 0.5, linetype = "dashed") +
  annotate("text", x = 2015, y = target[1,]$value, label = "REPowerEU\ndomestic", angle=90, vjust = 0.4, hjust = 0, color = "black", alpha = 0.5, size = 0.3*theme.single.text.size) +
  geom_hline(yintercept= target[2,]$value, color = "black", linewidth=1, alpha = 0.5, linetype = "dashed") +
  annotate("text", x = 2015, y = target[2,]$value, label = "REPowerEU\ntotal", angle=90, vjust = 0.4, hjust = 0, color = "black", alpha = 0.5, size = 0.3*theme.single.text.size) +
  geom_col(data=plotData %>% filter(scenario == mainScen), aes(fill = variable)) +
  geom_errorbar(data=plotData %>% filter(period != 2020, variable == vars[1], scenario == mainScen), aes(ymin=ribbon.min.total, ymax=ribbon.max.total), width=4, colour="black", alpha=0.9, size=1,position = position_dodge(width=2)) +
  # scale_y_continuous(labels = scales::percent, breaks = c(0,1,plotData[plotData$period == "2030",]$value,plotData[plotData$period == "2040",]$value,plotData[plotData$period == "2050",]$value)) +
  coord_flip() +
  theme_doubleColumn() +
  theme(panel.grid.minor.x = element_blank(),
        panel.grid.major.y = element_blank()) +
  #guides(y = "none") +
  theme(axis.title.y=element_blank()) +
  #theme(legend.position = "none") #+
  scale_fill_manual(values = color,
                    labels = setNames(names(vars), vars)[c("SE|Hydrogen|Electricity", "SE|Hydrogen|Gas|w/o CC", "SE|Hydrogen|Biomass|w/ CC")],
                    breaks = c("SE|Hydrogen|Electricity", "SE|Hydrogen|Gas|w/o CC", "SE|Hydrogen|Biomass|w/ CC"), guide = guide_legend(nrow = 1)) +
  theme(legend.text=element_text(size=theme.double.text.size)) +
  theme(legend.key.height = unit(0.7, "cm"),
    legend.key.width = unit(0.7,"cm")) +
  scale_y_continuous(breaks = c(0,round(sum(plotData %>% filter(scenario == mainScen, period == 2030) %>% pull(value))),round(sum(plotData %>% filter(scenario == mainScen, period == 2040) %>% pull(value))),round(sum(plotData %>% filter(scenario == mainScen, period == 2050) %>% pull(value))))) +
  #scale_x_reverse() +
  theme(legend.margin=margin(l = -2, unit='cm'))

ggsave(paste0(outFolder,"/Hydrogen.svg"), g[["Hydrogen"]] , device="svg", width = 6, height = 4, dpi=100, units = "in")
ggsave(paste0(outFolder,"/Hydrogen.png"), g[["Hydrogen"]] , device="png", width = 6, height = 4, dpi=300, units = "in")

```

```{r, echo=FALSE, fig.width=6, fig.height=4}

g[["Hydrogen"]]

```

:::

::::

---

# Figure 4. Demand panel

(a) Transport final energy (EJ/yr)

```{r, echo=FALSE}

FEvars <- list(
  "Domestic" = list(
    "LDV" = c(
      "FE|Transport|Pass|Road|LDV|Electricity" = "Electricity",
      "FE|Transport|Pass|Road|LDV|Gases" = "Gases",
      "FE|Transport|Pass|Road|LDV|Hydrogen" = "Hydrogen",
#      "FE|Transport|Pass|Road|LDV|Liquids" = "Liquids"
       "FE|Transport|Pass|Road|LDV|Liquids|Fossil" = "Liquids - Fossil",
       "FE|Transport|Pass|Road|LDV|Liquids|Biomass" = "Liquids - Biomass",
        "FE|Transport|Pass|Road|LDV|Liquids|Hydrogen" = "Liquids - Synthetic"
    ),
    "Buses" = c(
      "FE|Transport|Pass|Road|Bus|Electricity" = "Electricity",
      "FE|Transport|Pass|Road|Bus|Gases" = "Gases",
      "FE|Transport|Pass|Road|Bus|Hydrogen" = "Hydrogen",
#      "FE|Transport|Pass|Road|Bus|Liquids" = "Liquids"
        "FE|Transport|Pass|Road|Bus|Liquids|Fossil" = "Liquids - Fossil",
        "FE|Transport|Pass|Road|Bus|Liquids|Biomass" = "Liquids - Biomass",
        "FE|Transport|Pass|Road|Bus|Liquids|Hydrogen" = "Liquids - Synthetic"

    ),
    "Trucks" = c(
      "FE|Transport|Freight|Road|Electricity" = "Electricity",
      "FE|Transport|Freight|Road|Gases" = "Gases",
      "FE|Transport|Freight|Road|Hydrogen" = "Hydrogen",
#      "FE|Transport|Freight|Road|Liquids" = "Liquids"
        "FE|Transport|Freight|Road|Liquids|Fossil" = "Liquids - Fossil",
        "FE|Transport|Freight|Road|Liquids|Biomass" = "Liquids - Biomass",
        "FE|Transport|Freight|Road|Liquids|Hydrogen" = "Liquids - Synthetic"
    ),
    "Rail" = c(
      "FE|Transport|Pass|Rail|Electricity" = "Electricity",
      "FE|Transport|Pass|Rail|Gases" = "Gases",
      "FE|Transport|Pass|Rail|Hydrogen" = "Hydrogen",
#      "FE|Transport|Pass|Rail|Liquids" = "Liquids",
        "FE|Transport|Pass|Rail|Liquids|Fossil" = "Liquids - Fossil",
        "FE|Transport|Pass|Rail|Liquids|Biomass" = "Liquids - Biomass",
        "FE|Transport|Pass|Rail|Liquids|Hydrogen" = "Liquids - Synthetic",
      "FE|Transport|Freight|Rail|Electricity" = "Electricity",
      "FE|Transport|Freight|Rail|Gases" = "Gases",
      "FE|Transport|Freight|Rail|Hydrogen" = "Hydrogen",
#      "FE|Transport|Freight|Rail|Liquids" = "Liquids"
        "FE|Transport|Freight|Rail|Liquids|Fossil" = "Liquids - Fossil",
        "FE|Transport|Freight|Rail|Liquids|Biomass" = "Liquids - Biomass",
        "FE|Transport|Freight|Rail|Liquids|Hydrogen" = "Liquids - Synthetic"
    ),
    "Aviation" = c(
      "FE|Transport|Pass|Aviation|Domestic|Electricity" = "Electricity",
      "FE|Transport|Pass|Aviation|Domestic|Gases" = "Gases",
      "FE|Transport|Pass|Aviation|Domestic|Hydrogen" = "Hydrogen",
#      "FE|Transport|Pass|Aviation|Domestic|Liquids" = "Liquids"
        "FE|Transport|Pass|Aviation|Domestic|Liquids|Fossil" = "Liquids - Fossil",
        "FE|Transport|Pass|Aviation|Domestic|Liquids|Biomass" = "Liquids - Biomass",
        "FE|Transport|Pass|Aviation|Domestic|Liquids|Hydrogen" = "Liquids - Synthetic"
    )
  ),
  "International" = list(
    "Aviation" = c(
      "FE|Transport|Pass|Aviation|International|Electricity" = "Electricity",
      "FE|Transport|Pass|Aviation|International|Gases" = "Gases",
      "FE|Transport|Pass|Aviation|International|Hydrogen" = "Hydrogen",
#      "FE|Transport|Pass|Aviation|International|Liquids" = "Liquids"
        "FE|Transport|Pass|Aviation|International|Liquids|Fossil" = "Liquids - Fossil",
        "FE|Transport|Pass|Aviation|International|Liquids|Biomass" = "Liquids - Biomass",
        "FE|Transport|Pass|Aviation|International|Liquids|Hydrogen" = "Liquids - Synthetic"
    ),
    "Shipping" = c(
      "FE|Transport|Freight|Navigation|Electricity" = "Electricity",
      "FE|Transport|Freight|Navigation|Gases" = "Gases",
      "FE|Transport|Freight|Navigation|Hydrogen" = "Hydrogen",
#      "FE|Transport|Freight|Navigation|Liquids" = "Liquids",
        "FE|Transport|Freight|Navigation|Liquids|Fossil" = "Liquids - Fossil",
        "FE|Transport|Freight|Navigation|Liquids|Biomass" = "Liquids - Biomass",
        "FE|Transport|Freight|Navigation|Liquids|Hydrogen" = "Liquids - Synthetic",
      "FE|Transport|Freight|International Shipping|Electricity" = "Electricity",
      "FE|Transport|Freight|International Shipping|Gases" = "Gases",
      "FE|Transport|Freight|International Shipping|Hydrogen" = "Hydrogen",
#      "FE|Transport|Freight|International Shipping|Liquids" = "Liquids"
        "FE|Transport|Freight|International Shipping|Liquids|Fossil" = "Liquids - Fossil",
        "FE|Transport|Freight|International Shipping|Liquids|Biomass" = "Liquids - Biomass",
        "FE|Transport|Freight|International Shipping|Liquids|Hydrogen" = "Liquids - Synthetic"
    )
  )
)

FEvarList <- unlist(c(FEvars$Domestic$LDV,FEvars$Domestic$Buses,FEvars$Domestic$Trucks,FEvars$Domestic$Rail,FEvars$Domestic$Aviation,FEvars$International$Aviation,FEvars$International$Shipping))

typeVar <- c()
for(national in names(FEvars)){
  for(mode in names(FEvars[[national]])){
    for(var in names(FEvars[[national]][[mode]])){
      v <- mode
      names(v) <- var
      typeVar <- c(typeVar,v)
    }
  }
}

FEplotData <- df %>% 
  filter(region == "EU27", period >= 2020, period <= 2060, scenario == mainScen, variable %in% names(FEvarList))

FEplotData <- FEplotData %>%
  mutate(national = ifelse(variable %in% names(c(FEvars$Domestic$LDV,FEvars$Domestic$Buses,FEvars$Domestic$Trucks,FEvars$Domestic$Rail,FEvars$Domestic$Aviation)), "Domestic", "International"),
         name = FEvarList[as.character(variable)],
         type="Final Energy",
         mode = typeVar[as.character(variable)]) %>%
  group_by(model,scenario,region,unit,period,national,name,type,mode) %>%
  summarize(value=sum(value),.groups = "keep")

FEplotData$mode <- factor(FEplotData$mode, levels=c("LDV","Buses","Trucks","Rail","Shipping","Aviation"))

FEplotData$name <- factor(FEplotData$name, levels=c("Electricity","Hydrogen","Liquids","Liquids - Biomass","Liquids - Synthetic","Liquids - Fossil","Gases"))

plotColor <- c(
  "Electricity" = "#F3E16B",
  "Hydrogen" = "#82c9d9",
  "Liquids" = "#282828",
  "Liquids - Fossil" = "#282828",
  "Liquids - Biomass" = "#005900",
  "Liquids - Synthetic" = "#b56576",
  "Gases" = "#999959"
)

g[["transport fe"]] <- ggplot(FEplotData,aes(x=period,y=value)) +
   geom_bar(aes(fill=name),stat='identity') +
   facet_wrap(type~national+mode,nrow=1) +
   theme_singleColumn() +
   theme(axis.title.y=element_blank()) +
   scale_fill_manual(values = plotColor, guide = guide_legend(nrow = 2)) + #, guide = guide_legend(nrow = 2)
   scale_x_continuous(breaks = c(2020,2040,2060)) + 
  theme(
    strip.background = element_blank(),
    strip.text.x = element_blank()) +
  theme(legend.text=element_text(size=theme.single.text.size)) +
  theme(legend.key.height = unit(0.7, "cm"),
    legend.key.width = unit(0.7,"cm"))
#+ scale_y_continuous(name = "Final Energy")

ggsave(paste0(outFolder,"/transport fe.svg"), g[["transport fe"]] , device="svg", width = 12, height = 5, dpi=100, units = "in")
ggsave(paste0(outFolder,"/transport fe.png"), g[["transport fe"]] , device="png", width = 12, height = 5, dpi=300, units = "in")

```

```{r, echo=FALSE, fig.width=12, fig.height=5}

g[["transport fe"]]

```

:::: {style="display: flex;"}

::: {}

(b) LDV sales and ICE stock (Millions of vehicles)

```{r, echo=FALSE}

#LDV

Sales <- list(
  "BEV" = "Sales|Transport|LDV|BEV",
  "FCEV" = "Sales|Transport|LDV|FCEV",
  "Hybrid Electric" =	"Sales|Transport|LDV|Hybrid Electric",
  "Liquids" = "Sales|Transport|LDV|Liquids",
  "Gases" = "Sales|Transport|LDV|NG"
)

Stock <- list(
  "BEV" = "Stock|Transport|LDV|BEV",
  "FCEV" = "Stock|Transport|LDV|FCEV",
  "Hybrid Electric" =	"Stock|Transport|LDV|Hybrid Electric",
  "Liquids" = "Stock|Transport|LDV|Liquids",
  "Gases" = "Stock|Transport|LDV|NG"
)

plotColor <- c(
  "BEV" = "#F3E16B",
  "FCEV" = "#82c9d9",
  "Hybrid Electric" =	"#F37A5E",
  "Liquids" = "#282828",
  "Gases" = "#F7EFE8" 
)

vars <- Sales

plotData <- df %>% 
  filter(region == "EU27", period >= 2020, period <= 2060, scenario == mainScen, variable %in% unlist(vars)) %>% 
  mutate(variable = factor(variable, levels=vars))

labels <- setNames(names(vars),vars[names(vars)])

plotData$labels <- labels[as.character(plotData$variable)]

#stock
stockData <- df %>% 
  filter(region == "EU27", period >= 2020, period <= 2060, scenario == mainScen, variable %in% unlist(Stock)) %>% 
  mutate(variable = factor(variable, levels=Stock)) %>%
  filter(variable %in% c("Stock|Transport|LDV|Liquids","Stock|Transport|LDV|NG")) %>%
  group_by(period) %>%
  summarize(value = sum(value),.groups = "keep") %>%
  mutate(variable = "ICE Stock")
stockData$value <- stockData$value/10

g[["LDV Sales"]] <-
  ggplot(plotData,aes(x=period,y=value)) +
  geom_bar(aes(fill=variable),stat='identity') +
  geom_line(data=stockData, aes(color=variable),linetype=11, lwd = 1) +
# geom_line(data=h,linetype="dashed", linewidth = 1, color="black") +
 geom_hline(yintercept=0, color = "black", linewidth=2, alpha = 1) +
 geom_label_repel(data=data.frame(period=2030,value=stockData %>% filter(period == 2030) %>% pull(value)), label = "Stock",
                  #box.padding = 0.1,
                  #segment.curvature = -0.1,
                  #segment.square = TRUE,
                  #point.padding = 0.6,
                  nudge_x = 5,
                  nudge_y = 5,
                  #nudge_y = -0.02,
                  #force = 0.5,
                  hjust = 0,
                  direction="x", size = geom.text.double.size*0.9) +
  theme_doubleColumn() +
 #ylab(expression(paste("Energy and Industrial Processes Emissions - Mt" , CO[2], "eq/yr"))) +
 scale_fill_manual(values = setNames(plotColor[names(vars)],vars), labels = setNames(names(vars), vars), guide = guide_legend(nrow = 2, order = 1)) + #, guide = guide_legend(nrow = 2)
 scale_color_manual(values = c("ICE Stock"="black"), guide = guide_legend(order = 2, keywidth = unit(2, 'cm'))) +
 scale_x_continuous(breaks=c(2020,2030,2040,2050,2060)) +
 # theme(legend.margin=margin(l = -2, unit='cm')) + 
 #theme(legend.position="right", legend.justification = "bottom") +
 #theme(axis.title.y = element_blank()) +
 scale_y_continuous(name = "Sales", # - Million vehicles", 
                     #breaks = c(round(emi1990), round(emi1990*0.6), round(emi1990*0.45), round(emi1990*0.3), round(emi1990*0.15), 0),
                     sec.axis = sec_axis( trans=~.*10, name="Stock")) +
                                          #, labels = scales::percent, breaks = c(1,0.4,0.55,0.7,0.85, 0))) +
  theme(legend.text=element_text(size=theme.double.text.size),
        legend.spacing.x = unit(0.2, 'cm'))  +
  guides(color = "none")
                                                      #+
 # theme(legend.key.height = unit(0.7, "cm"),
 #   legend.key.width = unit(0.4,"cm"))

ggsave(paste0(outFolder,"/LDV Sales.svg"), g[["LDV Sales"]] , device="svg", width = 6, height = 5, dpi=100, units = "in")
ggsave(paste0(outFolder,"/LDV Sales.png"), g[["LDV Sales"]] , device="png", width = 6, height = 5, dpi=300, units = "in")

```

```{r, echo=FALSE, fig.width=6, fig.height=5}

   g[["LDV Sales"]]

```

(e) Industry CO2 emissions per sub-sector (Mt CO2/yr)

```{r, echo=FALSE}

# Industry Emissions
vars <- c("Cement" = "Emi|CO2|Energy|Demand|Industry|Cement",
          "Steel"="Emi|CO2|Energy|Demand|Industry|Steel",
          "Chemicals"="Emi|CO2|Energy|Demand|Industry|Chemicals",
          "Other Industry"="Emi|CO2|Energy|Demand|Industry|Other Industry")

plotColor <- c(
  "Cement" = "#355070",
  "Steel" = "#219ebc",
  "Chemicals" = "#eaac8b",
  "Other Industry" = "#e56b6f"
)

plotData <- df %>% 
  filter(region == "EU27", period %in% c(2020:2060), tgt2030 %in% tgt2030Scen, variable %in% unlist(vars)) %>% 
  mutate(variable = factor(variable, levels=vars)) %>%
  group_by(model,region,variable,period) %>% 
  filter(tgt2050 != "Npi") %>% 
  mutate(ribbon.max=max(value),
         ribbon.min=min(value))

# historical data #missing non-energy
#h <- hist[hist$model == "UNFCCC" & hist$region == "EU27" & hist$variable %in% vars,] %>% drop_na(value)

#View(hist[hist$model == "UNFCCC" & hist$region == "EU27" & hist$variable == "Emi|CO2|Industrial Processes",] %>% drop_na(value))
# Emi|CO2|Energy|Demand|Industry
# Emi|CO2|Energy|Demand|Industry|Cement
# Emi|CO2|Energy|Demand|Industry|Chemicals
# Emi|CO2|Energy|Demand|Industry|Other Industry
# Emi|CO2|Energy|Demand|Industry|Other Industry|AFOFI
# Emi|CO2|Energy|Demand|Industry|Steel
# Emi|CO2|Industrial Processes
# Emi|CO2|Industrial Processes|Chemicals
# Emi|CO2|Industrial Processes|Metals
# Emi|CO2|Industrial Processes|Minerals
# Emi|GHG|Energy|Demand|Industry
# Emi|GHG|Industrial Processes

g[["Industry Emissions"]] <- ggplot(plotData,aes(x=period,y=value)) +
    geom_hline(yintercept=0, color = "black", linewidth=2, alpha = 0.5) +
    geom_vline(xintercept=2020, color = "black", linewidth=1.5, alpha = 0.5,linetype="dashed") +
    geom_segment(aes(x=2040, y=0, xend=2040, yend=max(plotData %>% filter(scenario == mainScen, period == 2040) %>% pull(value))), linetype="dashed", color="black") +
    geom_line(data=plotData %>% filter(scenario == mainScen), aes(color=variable), linewidth=1.5) +
    #geom_line(data=h,linewidth=1,linetype="dashed") +
    geom_ribbon(aes(ymax = ribbon.max, ymin = ribbon.min, fill=variable), alpha = 0.3) + 
    geom_point(data=plotData %>% filter(scenario == mainScen, period == 2040), aes(color=variable),size=3,shape=21,fill="white",stroke=1.5) +
    scale_color_manual(values = c(setNames(plotColor[names(vars)],vars)), labels = setNames(names(vars), vars)) +
    scale_fill_manual(values = c(setNames(plotColor[names(vars)],vars)), labels = setNames(names(vars), vars)) +
    theme_doubleColumn() + 
    expand_limits(y=c(0,1)) +
    guides(color=guide_legend(nrow=2)) +
  theme(axis.title.y = element_blank())

ggsave(paste0(outFolder,"/Industry Emissions.svg"), g[["Industry Emissions"]] , device="svg", width = 6, height = 6, dpi=100, units = "in")
ggsave(paste0(outFolder,"/Industry Emissions.png"), g[["Industry Emissions"]] , device="png", width = 6, height = 6, dpi=300, units = "in")

```

```{r, echo=FALSE, fig.width=6, fig.height=6}

g[["Industry Emissions"]]

```

:::

::: {}

(c) Buildings useful energy
    Electricity based heating (EJ/yr)

```{r, echo=FALSE}

vars <- c(
  "Heat pump" = "UE|Buildings|Heating|Electricity|Heat pump",
  #"District\nHeating" = "UE|Buildings|Heating|District Heating",
  "Resistance" = "UE|Buildings|Heating|Electricity|Resistance"
#  "non-Heating Electricity" = "UE|Buildings|non-Heating|Electricity|Conventional"
)

plotData <- df %>% 
  filter(region == "EU27", period %in% c(2020,2030,2040,2050), tgt2030 %in% tgt2030Scen, variable %in% unlist(vars)) %>% 
  mutate(variable = factor(variable, levels=vars))  %>% 
  filter(tgt2050 != "Npi")  %>%
  group_by(model,scenario,region,period) %>% 
  mutate(cumsum = cumsum(value),
         sum = sum(value)) %>% 
  group_by(model,region,variable,period) %>% 
  mutate(ribbon.max=max(cumsum),
         ribbon.min=min(cumsum)) %>% 
  group_by(model,region,period) %>% 
  mutate(ribbon.max.total=max(sum),
         ribbon.min.total=min(sum))

colors <- c(
  "Heat pump" = "#D62828",
  "District\nHeating" = "#E07A5F",
  "Resistance" = "#F77F00",
  "Heating - Gases" = as.vector(options$aesthetics$legend$color["Gases"]),
  "Heating - Hydrogen" = as.vector(options$aesthetics$legend$color["Hydrogen"]),
  "Heating - Liquids" = as.vector(options$aesthetics$legend$color["Liquids"]),
  "Heating - Solids" = as.vector(options$aesthetics$legend$color["Solids"]),
  "non-Heating Electricity" = as.vector(options$aesthetics$legend$color["Electricity"])
)

g[["Buildings Heating - Useful energy - Electricity Based"]] <- 
  ggplot(plotData,aes(x=period,y=value)) +
  geom_col(data=plotData %>% filter(scenario == mainScen), aes(fill = variable)) +
  geom_errorbar(data=plotData %>% filter(period != 2020, variable == vars[1], scenario == mainScen), aes(ymin=ribbon.min.total, ymax=ribbon.max.total), width=4, colour="black", alpha=0.9, size=1,position = position_dodge(width=2)) +
  # scale_y_continuous(labels = scales::percent, breaks = c(0,1,plotData[plotData$period == "2030",]$value,plotData[plotData$period == "2040",]$value,plotData[plotData$period == "2050",]$value)) +
  coord_flip() +
  theme_doubleColumn() +
  theme(panel.grid.minor.x = element_blank(),
        panel.grid.major.y = element_blank()) +
  #guides(y = "none") +
  theme(axis.title.y=element_blank()) +
  #theme(legend.position = "none") #+
  scale_fill_manual(values = setNames(colors[names(vars)],vars), labels = setNames(names(vars), vars), guide = guide_legend(reverse = TRUE, nrow = 1)) +
  theme(legend.text=element_text(size=theme.double.text.size)) +
  theme(legend.key.height = unit(0.7, "cm"),
    legend.key.width = unit(0.7,"cm")) +
  scale_y_continuous(breaks = c(0,round(sum(plotData %>% filter(scenario == mainScen, period == "2020") %>% pull(value)),1),round(sum(plotData %>% filter(scenario == mainScen, period == "2030") %>% pull(value)),1),round(sum(plotData %>% filter(scenario == mainScen, period == "2040") %>% pull(value)),1),round(sum(plotData %>% filter(scenario == mainScen, period == "2050") %>% pull(value)),1))) #+
  #scale_x_reverse()

ggsave(paste0(outFolder,"/Buildings Heating - Useful energy - Electricity Based.svg"), g[["Buildings Heating - Useful energy - Electricity Based"]] , device="svg", width = 6, height = 4, dpi=100, units = "in")
ggsave(paste0(outFolder,"/Buildings Heating - Useful energy - Electricity Based.png"), g[["Buildings Heating - Useful energy - Electricity Based"]] , device="png", width = 6, height = 4, dpi=300, units = "in")

```

```{r, echo=FALSE, fig.width=6, fig.height=4}

g[["Buildings Heating - Useful energy - Electricity Based"]]

```

(f) Carbon Management (Mt CO2/yr)

```{r, echo=FALSE}

vars <- list(
  "atmospheric/biogenic CO2 (CDR)" = "Carbon Management|atmospheric/biogenic CO2 - CDR",
  "CCS of fossil origin" = "Carbon Management|CCS of fossil origin",
  "DAC" = "Carbon Management|Storage|DAC",
  "BECCS" = "Carbon Management|Storage|Biomass|Pe2Se", 
  "Fossil CCS" = "Carbon Management|Storage|Fossil|Pe2Se", 
  "Industry CCS" = "Carbon Management|Storage|Industry CCS"
    # "Industry Energy" = "Carbon Management|Storage|Industry Energy",
    # "Industry Process" = "Carbon Management|Storage|Industry Process"
)

plotColor <- c(
  "atmospheric/biogenic CO2 (CDR)" = "#c1d8f0",
  "CCS of fossil origin" = "#6b5b4b",
  "DAC" = "#CC6666",
  "BECCS" = "#8bc18b",
  "Industry CCS" = "#5f6769",
  "Fossil CCS" = "#47331f"
)

plotData <- df %>% 
  filter(region == "EU27", period %in% c(2030, 2040, 2050), tgt2030 %in% tgt2030Scen, variable %in% unlist(vars)) %>% 
  mutate(CDRGroup = ifelse(as.character(variable) %in% c("Carbon Management|CCS of fossil origin","Carbon Management|atmospheric/biogenic CO2 - CDR"), "CDR", "CCS"), 
         variable = factor(variable, levels=vars),
         value = ifelse(grepl("Emi",variable,fixed = TRUE), -value, value))  %>% 
  mutate(CDRGroup = factor(CDRGroup, levels=c("CDR","CCS"))) %>%
  group_by(model,scenario,region,period, CDRGroup) %>% 
  mutate(cumsum = cumsum(value),
         sum = sum(value)) %>% 
  group_by(model,region,variable,period, CDRGroup) %>% 
  mutate(ribbon.max=max(cumsum),
         ribbon.min=min(cumsum)) %>% 
  group_by(model,region,period, CDRGroup) %>% 
  mutate(ribbon.max.total=max(sum),
         ribbon.min.total=min(sum)) %>% 
  mutate(period = factor(period, levels=c(2050,2040,2030)))

 g[["Carbon Management"]] <- ggplot(plotData,aes(x=CDRGroup,y=value)) +
  #geom_area(aes(fill=variable)) +
  geom_bar(data=plotData %>% filter(scenario == mainScen), aes(fill=variable),stat='identity') + 
  geom_errorbar(data=plotData %>% filter(variable == vars[1], scenario == mainScen), aes(ymin=ribbon.min.total, ymax=ribbon.max.total), width=0.4, colour="black", alpha=0.9, size=1,position = position_dodge(width=2)) +
  facet_wrap(period ~ ., ncol=1, strip.position="left") +
  coord_flip() +
  theme_doubleColumn() +
  scale_fill_manual(values = setNames(plotColor[names(vars)],vars), labels = setNames(names(vars), vars), guide = guide_legend(nrow = 4, reverse = TRUE)) + 
  theme(legend.margin=margin(l = -2, unit='cm')) +
  theme(axis.title.x=element_blank()) +
    theme(legend.title = element_blank()) +
  theme(axis.title.y = element_blank()) +
  theme(legend.text=element_text(size=theme.double.text.size)) +
  theme(legend.key.height = unit(0.7, "cm"),
    legend.key.width = unit(0.7,"cm")) +
  theme(legend.margin=margin(l = -2.5, unit='cm')) + 
  theme(strip.text.y.left = element_text(angle = 0)) +
  # theme(
  #   strip.background = element_blank(),
  #   strip.text.x = element_blank()
  # ) +
  theme(axis.text.y=element_blank(), #remove x axis labels
        axis.ticks.y=element_blank(), #remove x axis ticks
        )

 ggsave(paste0(outFolder,"/Carbon Management.svg"), g[["Carbon Management"]] , device="svg", width = 6, height = 6, dpi=100, units = "in")
 ggsave(paste0(outFolder,"/Carbon Management.png"), g[["Carbon Management"]] , device="png", width = 6, height = 6, dpi=300, units = "in")

```

```{r, echo=FALSE, fig.width=6, fig.height=6}

g[["Carbon Management"]]

```

:::

::::

---

# Figure 5. Targets update

```{r, echo=FALSE}

# “existent targets”
# emission reduction - 2030 (FitFor55) - 2030 scenarios - 2040 scenarios
# energy savings  - 2030 (FitFor55 and REPowerEU) - 2030 scenarios - 2040 scenarios
# renewables target - 2030 (RED II, RED II revised, REPowerEU) - 2030 scenarios - 2040 scenarios
# solar photovoltaics capacity - 2030 (REPowerEU) - 2030 scenarios - 2040 scenarios
# CCS capacity target - 2030 EU-wide 50 MtCO2/y “Net-Zero Industry Act"
# “implicit targets”
# electrification? -> fe electrification share -> 45%
# “carbon-content in electricity” ->“carbon-free electricity” (>95%) 
# Total Residual Fossil emissions
# sectoral emissions -> Transport emissions -> 23% of 2020 
#                                   -> Industry and buildings emissions -> 27% of 2020
# “residual fossil”     -> Complete coal-phase by 2040
#                                  -> natural gas
#                                 -> oil
# “net trade” -> net-supply independency of gas by 2040
#                       -> one third of oil imports by 2040

#Emissions Reduction
plotData <- rbind(
    df %>% 
      filter(region == "EU27", period == 2030, tgt2030 %in% tgt2030Scen) %>%
      calc_addVariable( "`Emissions with LULUCF and with intra-EU aviation`" = "`Emi|GHG|LULUCF national accounting` + `Emi|CO2|Energy|Demand|Transport|International Bunkers|Intra-region`" , units = "Mt CO2/yr", only.new=T) %>% 
      mutate(value = round(1-(value / tgt %>% filter(target == "Emissions with LULUCF and with intra-EU aviation") %>% pull(reference)) ,2)) %>% 
      select(-variable)
    ,
    df %>% 
      filter(region == "EU27", period == 2040, tgt2030 %in% tgt2030Scen) %>%
      calc_addVariable( "`Emissions with LULUCF and with international transport`" = "`Emi|GHG|LULUCF national accounting` + `Emi|CO2|Energy|Demand|Transport|International Bunkers`" , units = "Mt CO2/yr", only.new=T) %>% 
      mutate(value = round(1-(value / tgt %>% filter(target == "Emissions with LULUCF and with international transport") %>% pull(reference)) ,2)) %>% 
      select(-variable)
  ) %>%
  filter(tgt2050 != "Npi") %>%
  mutate(target = "Emissions\nreductions") %>%
  group_by(period) %>% 
  mutate(min = min(value),
        max = max(value))

g[["Emissions reductions"]] <- ggplot(plotData, aes(x = target, y = value, group=period)) +
  geom_errorbar(aes(ymin=min, ymax=max), width=0.3, colour="black", alpha=0.9, linewidth=1) +
  geom_boxplot(outlier.shape = NA, width=0.6) +
         #stat_summary(geom = "text", fun = quantile,
        #              aes(label = paste0(round(..y..*100),"%")),
        #       #position=position_nudge(x=c(0.1,0.35,0.35,0.35,0.1)), 
         #      hjust=-1, angle = 90) +
  geom_hline(yintercept=0.55, color = "black", linewidth=1, alpha = 0.5, linetype = "dashed") +
  annotate("text", x = "Emissions\nreductions", y = 0.55, label = "FitFor55", angle=90, vjust = -0.5, hjust = 1, color = "black", alpha = 0.5) +
  facet_grid(target~period, scales="free") +
  theme_doubleColumn() +
  coord_flip(clip = 'off') +
  theme(axis.title.y=element_blank()) +
  theme(strip.text.y = element_blank()) +
  scale_y_continuous(breaks=seq(0, 1, 0.2), labels = scales::percent) +
  expand_limits(y=c(0,1)) + 
  theme(panel.spacing = unit(4, "lines")) +
  theme(axis.text.x = element_text(size = theme.double.text.size))

# Energy savings
plotData <- df %>% filter(region == "EU27", period %in% c(2030,2040), tgt2030 %in% tgt2030Scen, variable == "FE|w/o Non-energy Use w/o Bunkers") %>% select(-variable) %>%
  mutate(value = round(value * 23.8845896627496)) %>%
  filter(tgt2050 != "Npi") %>%
  mutate(target = "Energy\nsavings") %>%
  group_by(period) %>% 
  mutate(min = min(value),
        max = max(value))

g[["Energy savings"]] <- ggplot(plotData, aes(x = target, y = value, group=period)) +
  geom_errorbar(aes(ymin=min, ymax=max), width=0.3, colour="black", alpha=0.9, linewidth=1) +   
  geom_boxplot(outlier.shape = NA, width=0.6) +
         #stat_summary(geom = "text", fun = quantile,
        #              aes(label = paste0(round(..y..*100),"%")),
        #       #position=position_nudge(x=c(0.1,0.35,0.35,0.35,0.1)), 
         #      hjust=-1, angle = 90) +
  geom_hline(yintercept=750, color = "black", linewidth=1, alpha = 0.5, linetype = "dashed") +
  annotate("text", x = "Energy\nsavings", y = 750, label = "FitFor55", angle=90, vjust = -0.5, hjust = 1, color = "black", alpha = 0.5) +
  geom_hline(yintercept=787, color = "black", linewidth=1, alpha = 0.5, linetype = "dashed") +
  annotate("text", x = "Energy\nsavings", y = 787, label = "REPowerEU", angle=90, vjust = 1.5, hjust = 0.7, color = "black", alpha = 0.5) +
  facet_grid(target~period, scales="free") +
  theme_doubleColumn() +
  coord_flip(clip = 'off') +
  theme(axis.title.y=element_blank()) +
  theme(strip.text.y = element_blank()) +
  theme(strip.text.x = element_blank()) +
  expand_limits(y=c(400,1000)) + 
  theme(panel.spacing = unit(4, "lines")) +
  theme(axis.text.x = element_text(size = theme.double.text.size))

# Renewables share
plotData <- df %>% filter(region == "EU27", period %in% c(2030,2040), tgt2030 %in% tgt2030Scen, variable == "SE|Renewables Share") %>% select(-variable) %>%
  filter(tgt2050 != "Npi") %>%
  mutate(target = "Renewables\nshare")

plotData$value <- plotData$value*(
  addHist$res %>% filter(type == "RES", period == 2020) %>% pull(value) /
   df %>% filter(region == "EU27", period == 2020, variable == "SE|Renewables Share", scenario == mainScen) %>% pull(value) 
)

plotData <- plotData %>%
  group_by(period) %>% 
  mutate(min = min(value),
        max = max(value))

g[["Renewables share"]] <- ggplot(plotData, aes(x = target, y = value, group=period)) +
  geom_errorbar(aes(ymin=min, ymax=max), width=0.3, colour="black", alpha=0.9, linewidth=1) +   
  geom_boxplot(outlier.shape = NA, width=0.6) +
         #stat_summary(geom = "text", fun = quantile,
        #              aes(label = paste0(round(..y..*100),"%")),
        #       #position=position_nudge(x=c(0.1,0.35,0.35,0.35,0.1)), 
         #      hjust=-1, angle = 90) +
  geom_hline(yintercept=0.4, color = "black", linewidth=1, alpha = 0.5, linetype = "dashed") +
  annotate("text", x = "Renewables\nshare", y = 0.4, label = "RED II revised", angle=90, vjust = -0.5, hjust = 0.5, color = "black", alpha = 0.5) +
  geom_hline(yintercept=0.45, color = "black", linewidth=1, alpha = 0.5, linetype = "dashed") +
  annotate("text", x = "Renewables\nshare", y = 0.45, label = "REPowerEU", angle=90, vjust = 1.5, hjust = 0.6, color = "black", alpha = 0.5) +
  facet_grid(target~period, scales="free") +
  theme_doubleColumn() +
  coord_flip(clip = 'off') +
  theme(axis.title.y=element_blank()) +
  theme(strip.text.y = element_blank()) +
  theme(strip.text.x = element_blank()) +
  scale_y_continuous(breaks=seq(0, 1, 0.2), labels = scales::percent) +
  expand_limits(y=c(0,1)) + 
  theme(panel.spacing = unit(4, "lines")) +
  theme(axis.text.x = element_text(size = theme.double.text.size))

#

# Solar photovoltaics
plotData <- df %>% filter(region == "EU27", period %in% c(2030,2040), tgt2030 %in% tgt2030Scen, variable == "Cap|Electricity|Solar|PV") %>% select(-variable) %>%
  filter(tgt2050 != "Npi") %>%
  mutate(target = "Solar\nphotovoltaics") %>%
  group_by(period) %>% 
  mutate(min = min(value),
        max = max(value))

#target <- data.frame(target=c("REPowerEU","REPowerEU"),period=c(2025,2030),value=c(320,600)) 
g[["Solar photovoltaics"]] <- ggplot(plotData, aes(x = target, y = value, group=period)) +
  geom_errorbar(aes(ymin=min, ymax=max), width=0.3, colour="black", alpha=0.9, linewidth=1) +   
  geom_boxplot(outlier.shape = NA, width=0.6) +
         #stat_summary(geom = "text", fun = quantile,
        #              aes(label = paste0(round(..y..*100),"%")),
        #       #position=position_nudge(x=c(0.1,0.35,0.35,0.35,0.1)), 
         #      hjust=-1, angle = 90) +
  geom_hline(yintercept=600, color = "black", linewidth=1, alpha = 0.5, linetype = "dashed") +
  annotate("text", x = "Solar\nphotovoltaics", y = 600, label = "REPowerEU", angle=90, vjust = -0.5, hjust = 0.6, color = "black", alpha = 0.5) +
  facet_grid(target~period, scales="free") +
  theme_doubleColumn() +
  coord_flip(clip = 'off') +
  theme(axis.title.y=element_blank()) +
  theme(strip.text.y = element_blank()) +
  theme(strip.text.x = element_blank()) +
  expand_limits(y=c(0,2500)) + 
  theme(panel.spacing = unit(4, "lines")) +
  theme(axis.text.x = element_text(size = theme.double.text.size))

# Carbon Capture and Storage 
plotData <- df %>% filter(region == "EU27", period %in% c(2030,2040), tgt2030 %in% tgt2030Scen, variable == "Carbon Management|Storage") %>% select(-variable) %>%
  filter(tgt2050 != "Npi") %>%
  mutate(target = "Carbon\nCapture and\nStorage") %>%
  group_by(period) %>% 
  mutate(min = min(value),
        max = max(value))

g[["Carbon Capture and Storage"]] <- ggplot(plotData, aes(x = target, y = value, group=period)) +
  geom_errorbar(aes(ymin=min, ymax=max), width=0.3, colour="black", alpha=0.9, linewidth=1) +   
  geom_boxplot(outlier.shape = NA, width=0.6) +
         #stat_summary(geom = "text", fun = quantile,
        #              aes(label = paste0(round(..y..*100),"%")),
        #       #position=position_nudge(x=c(0.1,0.35,0.35,0.35,0.1)), 
         #      hjust=-1, angle = 90) +
  geom_hline(yintercept=50, color = "black", linewidth=1, alpha = 0.5, linetype = "dashed") +
  annotate("text", x = "Carbon\nCapture and\nStorage", y = 50, label = "Net-Zero\nIndustry Act", angle=90, vjust = -0.2, hjust = 0.6, color = "black", alpha = 0.5) +
  facet_grid(target~period, scales="free") +
  theme_doubleColumn() +
  coord_flip(clip = 'off') +
  theme(axis.title.y=element_blank()) +
  theme(strip.text.y = element_blank()) +
  theme(strip.text.x = element_blank()) +
  #expand_limits(y=c(0,ceiling(max(plotData$value)/50)*50)) + 
  expand_limits(y=c(0,400)) + 
  theme(panel.spacing = unit(4, "lines")) +
  theme(axis.text.x = element_text(size = theme.double.text.size))

# Novel CDR = atmospheric/biogenic CO2 (CDR) 
plotData <- df %>% filter(region == "EU27", period %in% c(2030,2040), tgt2030 %in% tgt2030Scen, variable == "Carbon Management|atmospheric/biogenic CO2 - CDR") %>% select(-variable) %>%
  filter(tgt2050 != "Npi") %>%
  mutate(target = "Novel CDR") %>%
  group_by(period) %>% 
  mutate(min = min(value),
        max = max(value))

g[["Novel CDR"]] <- ggplot(plotData, aes(x = target, y = value, group=period)) +
  geom_errorbar(aes(ymin=min, ymax=max), width=0.3, colour="black", alpha=0.9, linewidth=1) +   
  geom_boxplot(outlier.shape = NA, width=0.6) +
         #stat_summary(geom = "text", fun = quantile,
        #              aes(label = paste0(round(..y..*100),"%")),
        #       #position=position_nudge(x=c(0.1,0.35,0.35,0.35,0.1)), 
         #      hjust=-1, angle = 90) +
  facet_grid(target~period, scales="free") +
  theme_doubleColumn() +
  coord_flip(clip = 'off') +
  theme(axis.title.y=element_blank()) +
  theme(strip.text.y = element_blank()) +
  theme(strip.text.x = element_blank()) +
  expand_limits(y=c(0,250)) + 
  #expand_limits(y=c(0,ceiling(max(plotData$value)/50)*50)) + 
  theme(panel.spacing = unit(4, "lines")) +
  theme(axis.text.x = element_text(size = theme.double.text.size))

# Direct Electrification
plotData <- df %>% filter(region == "EU27", period %in% c(2030,2040), tgt2030 %in% tgt2030Scen, variable == "Final Energy|Electricity Share") %>% select(-variable) %>%
  filter(tgt2050 != "Npi") %>%
  mutate(target = "Direct\nElectrification") %>%
  group_by(period) %>% 
  mutate(min = min(value),
        max = max(value))

g[["Direct Electrification"]] <- ggplot(plotData, aes(x = target, y = value, group=period)) +
    geom_errorbar(aes(ymin=min, ymax=max), width=0.3, colour="black", alpha=0.9, linewidth=1) +   
  geom_boxplot(outlier.shape = NA, width=0.6) +
         #stat_summary(geom = "text", fun = quantile,
        #              aes(label = paste0(round(..y..*100),"%")),
        #       #position=position_nudge(x=c(0.1,0.35,0.35,0.35,0.1)), 
         #      hjust=-1, angle = 90) +
  facet_grid(target~period, scales="free") +
  theme_doubleColumn() +
  coord_flip(clip = 'off') +
  theme(axis.title.y=element_blank()) +
  theme(strip.text.y = element_blank()) +
  theme(strip.text.x = element_blank()) +
  scale_y_continuous(breaks=seq(0, 1, 0.2), labels = scales::percent) +
  expand_limits(y=c(0,1)) + 
  theme(panel.spacing = unit(4, "lines")) +
  theme(axis.text.x = element_text(size = theme.double.text.size))

# Carbon-free electricity share
plotData <- df %>% filter(region == "EU27", period %in% c(2030,2040), tgt2030 %in% tgt2030Scen, variable == "SE|Electricity|carbon-free") %>% select(-variable) %>%
  filter(tgt2050 != "Npi") %>%
  mutate(target = "Carbon-free\nelectricity\nshare") %>%
  group_by(period) %>% 
  mutate(min = min(value),
        max = max(value))

g[["Carbon-free electricity share"]] <- ggplot(plotData, aes(x = target, y = value, group=period)) +
    geom_errorbar(aes(ymin=min, ymax=max), width=0.3, colour="black", alpha=0.9, linewidth=1) +   
  geom_boxplot(outlier.shape = NA, width=0.6) +
         #stat_summary(geom = "text", fun = quantile,
        #              aes(label = paste0(round(..y..*100),"%")),
        #       #position=position_nudge(x=c(0.1,0.35,0.35,0.35,0.1)), 
         #      hjust=-1, angle = 90) +
  facet_grid(target~period, scales="free") +
  theme_doubleColumn() +
  coord_flip(clip = 'off') +
  theme(axis.title.y=element_blank()) +
  theme(strip.text.y = element_blank()) +
  theme(strip.text.x = element_blank()) +
  scale_y_continuous(breaks=seq(0, 1, 0.2), labels = scales::percent) +
  expand_limits(y=c(0,1)) + 
  theme(panel.spacing = unit(4, "lines")) +
  theme(axis.text.x = element_text(size = theme.double.text.size))

# Residual Fossil Emissions
plotData <- df %>% filter(region == "EU27", period %in% c(2030,2040), tgt2030 %in% tgt2030Scen, variable == "Emi|CO2|Residual Fossil") %>% select(-variable) %>%
  filter(tgt2050 != "Npi") %>%
  mutate(target = "Residual\nFossil\nEmissions") %>%
  group_by(period) %>% 
  mutate(min = min(value),
        max = max(value))

g[["Residual Fossil Emissions"]] <- ggplot(plotData, aes(x = target, y = value, group=period)) +
  geom_errorbar(aes(ymin=min, ymax=max), width=0.3, colour="black", alpha=0.9, linewidth=1) +   
  geom_boxplot(outlier.shape = NA, width=0.6) +
         #stat_summary(geom = "text", fun = quantile,
        #              aes(label = paste0(round(..y..*100),"%")),
        #       #position=position_nudge(x=c(0.1,0.35,0.35,0.35,0.1)), 
         #      hjust=-1, angle = 90) +
  facet_grid(target~period, scales="free") +
  theme_doubleColumn() +
  coord_flip(clip = 'off') +
  theme(axis.title.y=element_blank()) +
  theme(strip.text.y = element_blank()) +
  theme(strip.text.x = element_blank()) +
  expand_limits(y=c(0,2500)) + 
  theme(panel.spacing = unit(4, "lines")) +
  theme(axis.text.x = element_text(size = theme.double.text.size))

g[["Targets"]] <- gridExtra::arrangeGrob(
  g[["Emissions reductions"]] + theme(strip.text.x = element_blank()) + theme(plot.margin=margin(b=20, l = 30)), #+ theme(plot.margin=margin(10,10,30,10)),margin(t = 0, r = 0, b = 0, l = 0, unit = "pt")
  g[["Energy savings"]] + theme(plot.margin=margin(b=20, l = 57)),
  g[["Renewables share"]] + theme(plot.margin=margin(b=20, l = 15)),
  g[["Solar photovoltaics"]] + theme(plot.margin=margin(b=20, l = 10)),
  g[["Carbon Capture and Storage"]] + theme(plot.margin=margin(b=20, l = 15)),
  g[["Novel CDR"]] + theme(plot.margin=margin(b=20, l = 25)),
  g[["Direct Electrification"]] + theme(plot.margin=margin(b=20, l = 19)),
  g[["Carbon-free electricity share"]] + theme(plot.margin=margin(b=20, l = 24)),
  g[["Residual Fossil Emissions"]] + theme(plot.margin=margin(b=20, l = 39)),
  ncol = 1
)

ggsave(paste0(outFolder,"/targets.svg"), g[["Targets"]] , device="svg", width = 12, height = 17, dpi=100, units = "in")
ggsave(paste0(outFolder,"/targets.png"), g[["Targets"]] , device="png", width = 12, height = 17, dpi=300, units = "in")

```

```{r, echo=FALSE, fig.width=12, fig.height=17}

grid.newpage()
grid.draw(g[["Targets"]])

```