### 2023-06-22 paperCharts.Rmd
additional paper numbers and target charts

---
title: "2040 greenhouse gas reduction targets"
output:
  html_document:
    theme: paper
    toc: true
    toc_float:
      collapsed: false
---

<style>
  .main-container {
    max-width: 95% !important;
  }
  .toc-content {
    padding-left: 0px !important;
  }
  .svg-container {
    margin: auto !important;
  }
</style>

Description:
Charts created for the paper:

Rodrigues, R., Pietzcker, R., Luderer, G., Sitarz, J., Merfort, A., Hasse, R., Hoppe, J., Pehl, M., Ershad, A. M., Baumstark, L. 2040 greenhouse gas reduction targets and energy transitions in line with the EU Green Deal. in preparation (2023).

<!-- Options -->
```{r options, echo=FALSE}

forceUpdateData = FALSE

mainScen <- "Nzero_55_bio7p5" # "Nzero_55_ff55Eff_bio7p5" #Nzero_55_ff55Eff_bio12" # #Nzero_55_bio12" #"Nzero_55_ff55Eff_bio12"

tgt2030Scen <- c(55,57,59)

```

<!-- Loading required packages -->
```{r Loading_required_packages, echo=FALSE}

# install.packages("remotes")
#remotes::install_github("coolbutuseless/ggpattern")

  packagesList <- c("quitte","dplyr","ggplot2","geomtextpath","tidyr","grid","stringr","gridExtra","ggrepel","kableExtra")
  packages <- suppressWarnings(suppressMessages(lapply(packagesList, function(x){ if(!require(x, character.only = T, quietly = T)){ install.packages(x); return(paste0("installed missing package: ", x)) } else paste0("required package is installed: ", x) } )))
```

<!-- Setting default chunk options -->
```{r chunkOptions, echo=FALSE}
#knitr::opts_chunk$set(echo = TRUE)
# setting global R chunk options (https://yihui.name/knitr/options/#chunk_options)
  knitr::opts_chunk$set(dev='svglite', # svg, png,...
                        fig.ext = ".svg",
                        #fig.asp = .8 # default aspect ratio
                        fig.width = 12,
                        fig.height = 6,
                        dpi=100
                        )
```

<!-- Aesthetics -->
```{r loading_aesthetics, echo=FALSE, include=FALSE}

#loading theme and colors
source("./R/aesthetics.R")

```

<!-- load data -->
```{r loading_data, echo=FALSE, include=FALSE}

#loading data
source("./R/loadData.R")

```

<!-- loading reference emissions -->
```{r loading_referenceEmissions, echo=FALSE}

#loading reference emissions
source("./R/loadReferenceEmissions.R")

```

<!-- harmonizing data -->
```{r harmonizingData, echo=FALSE}

#harmonizing data (scale emissions)
source("./R/harmonizeData.R")

```

<!-- efficiency target -->
```{r loading_efficiencyTarget, echo=FALSE}

#loading reference emissions
source("./R/efficiencyTarget.R")

```

---

# Data {.tabset}

<!-- Outputs -->

## Reference emissions

```{r, echo=FALSE}

selectedTgt %>%
  replace(is.na(.),"") %>% 
  select(target,region,period,tgt55,tgt57,tgt59,tgt61,tgt63,tgt65) %>% 
  kbl() %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"))

```

---

## Efficiency targets

```{r, echo=FALSE}

effTgt %>%
  kbl() %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"))

```

---

## 2040 Emission Reductions

```{r, echo=FALSE}

reductions2040 <- rbind(
  remindEmi %>% 
  filter(variable == "Emissions with LULUCF and with international transport", region == "EU27", period == 2040, tgt2050 != "Npi") %>% 
  mutate(value = 1-(value / tgt %>% filter(target == "Emissions with LULUCF and with international transport") %>% pull(reference)))
  ,
  remindEmi %>% 
  filter(variable == "Emissions without LULUCF and with international transport", region == "EU27", period == 2040, tgt2050 != "Npi") %>% 
  mutate(value = 1-(value / tgt %>% filter(target == "Emissions without LULUCF and with international transport") %>% pull(reference)))
  ,
  remindEmi %>% 
  filter(variable == "Emissions with LULUCF and with intra-EU aviation", region == "EU27", period == 2040, tgt2050 != "Npi") %>% 
  mutate(value = 1-(value / tgt %>% filter(target == "Emissions with LULUCF and with intra-EU aviation") %>% pull(reference)))
  ,
  remindEmi %>% 
  filter(variable == "Emissions without LULUCF and with intra-EU aviation", region == "EU27", period == 2040, tgt2050 != "Npi") %>% 
  mutate(value = 1-(value / tgt %>% filter(target == "Emissions with LULUCF and with intra-EU aviation") %>% pull(reference)))
)

reductions2040 %>% 
  filter(tgt2030 %in% tgt2030Scen) %>% 
  group_by(variable) %>%
  summarize(minTarget = min(value), maxTarget=max(value),.groups = "keep") %>%
  mutate(minTarget = paste0(round(minTarget*100), "%"),
         maxTarget = paste0(round(maxTarget*100), "%"),
         period = 2040) %>% 
  rename(TargetReference = variable) %>%
  relocate(period, .after=TargetReference) %>%
  kbl() %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"))

bdgt <- remindEmi %>% 
  filter(region == "EU27", period %in% c(2020:2050), tgt2030 %in% tgt2030Scen) %>%
    mutate(weight = ifelse(period %in% c(2020,2050),2.5,5),
           budget = value*weight) %>%
    group_by(model,scenario,region,variable,tgt2050,tgt2030,efficiency,bioLim,limCCS) %>%
    summarize(budget=round(sum(budget)/1000,2),.groups = "keep")

budget <- data.frame(
  Budget=c("Emissions with LULUCF and with international transport","Emissions without LULUCF and with international transport","Emissions with LULUCF and with intra-EU aviation","Emissions without LULUCF and with intra-EU aviation"),
  reference_value=c(round(bdgt %>% filter(variable == "Emissions with LULUCF and with international transport", scenario == mainScen) %>% pull(budget)),
                    round(bdgt %>% filter(variable == "Emissions without LULUCF and with international transport", scenario == mainScen) %>% pull(budget)),
                    round(bdgt %>% filter(variable == "Emissions with LULUCF and with intra-EU aviation", scenario == mainScen) %>% pull(budget)),
                    round(bdgt %>% filter(variable == "Emissions without LULUCF and with intra-EU aviation", scenario == mainScen) %>% pull(budget))),
  min = c(round(min(bdgt %>% filter(variable == "Emissions with LULUCF and with international transport") %>% pull(budget))),
                    round(min(bdgt %>% filter(variable == "Emissions without LULUCF and with international transport") %>% pull(budget))),
                    round(min(bdgt %>% filter(variable == "Emissions with LULUCF and with intra-EU aviation") %>% pull(budget))),
                    round(min(bdgt %>% filter(variable == "Emissions without LULUCF and with intra-EU aviation") %>% pull(budget)))),
  max = c(round(max(bdgt %>% filter(variable == "Emissions with LULUCF and with international transport") %>% pull(budget))),
                    round(max(bdgt %>% filter(variable == "Emissions without LULUCF and with international transport") %>% pull(budget))),
                    round(max(bdgt %>% filter(variable == "Emissions with LULUCF and with intra-EU aviation") %>% pull(budget))),
                    round(max(bdgt %>% filter(variable == "Emissions without LULUCF and with intra-EU aviation") %>% pull(budget))))
)

budget %>% 
  kbl() %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"))

```

---

## Results

```{r, echo=FALSE}

indirect_electrification <- data.frame(
  indirect_electrification_EJ = c("total","buildings","transport","industry"),
  reference_value=c(
    round(df %>% filter(region == "EU27", period == 2040, variable %in% c("FE|Hydrogen","FE|Gases|Hydrogen","FE|Liquids|Hydrogen"), scenario == mainScen) %>% group_by(scenario,tgt2030) %>% summarize(value = sum(value),.groups = "keep") %>% pull(value),1),
    round(df %>% filter(region == "EU27", period == 2040, variable %in% c("FE|Buildings|Hydrogen","FE|Buildings|Gases|Hydrogen","FE|Buildings|Liquids|Hydrogen"), scenario == mainScen) %>% group_by(scenario,tgt2030) %>% summarize(value = sum(value),.groups = "keep") %>% pull(value),1),
    round(df %>% filter(region == "EU27", period == 2040, variable %in% c("FE|Transport|Hydrogen","FE|Transport|Gases|Hydrogen","FE|Transport|Liquids|Hydrogen"), scenario == mainScen) %>% group_by(scenario,tgt2030) %>% summarize(value = sum(value),.groups = "keep") %>% pull(value),1),
    round(df %>% filter(region == "EU27", period == 2040, variable %in% c("FE|Industry|Hydrogen","FE|Industry|Gases|Hydrogen","FE|Industry|Liquids|Hydrogen"), scenario == mainScen) %>% group_by(scenario,tgt2030) %>% summarize(value = sum(value),.groups = "keep") %>% pull(value),1)
  ),
  min=c(
    round(min(df %>% filter(region == "EU27", period == 2040, variable %in% c("FE|Hydrogen","FE|Gases|Hydrogen","FE|Liquids|Hydrogen"), tgt2030 %in% tgt2030Scen) %>% group_by(scenario,tgt2030) %>% summarize(value = sum(value),.groups = "keep") %>% pull(value)),1),
    round(min(df %>% filter(region == "EU27", period == 2040, variable %in% c("FE|Buildings|Hydrogen","FE|Buildings|Gases|Hydrogen","FE|Buildings|Liquids|Hydrogen"), tgt2030 %in% tgt2030Scen) %>% group_by(scenario,tgt2030) %>% summarize(value = sum(value),.groups = "keep") %>% pull(value)),1),
    round(min(df %>% filter(region == "EU27", period == 2040, variable %in% c("FE|Transport|Hydrogen","FE|Transport|Gases|Hydrogen","FE|Transport|Liquids|Hydrogen"), tgt2030 %in% tgt2030Scen) %>% group_by(scenario,tgt2030) %>% summarize(value = sum(value),.groups = "keep") %>% pull(value)),1),
    round(min(df %>% filter(region == "EU27", period == 2040, variable %in% c("FE|Industry|Hydrogen","FE|Industry|Gases|Hydrogen","FE|Industry|Liquids|Hydrogen"), tgt2030 %in% tgt2030Scen) %>% group_by(scenario,tgt2030) %>% summarize(value = sum(value),.groups = "keep") %>% pull(value)),1)
  ),
  max=c(
    round(max(df %>% filter(region == "EU27", period == 2040, variable %in% c("FE|Hydrogen","FE|Gases|Hydrogen","FE|Liquids|Hydrogen"), tgt2030 %in% tgt2030Scen) %>% group_by(scenario,tgt2030) %>% summarize(value = sum(value),.groups = "keep") %>% pull(value)),1),
    round(max(df %>% filter(region == "EU27", period == 2040, variable %in% c("FE|Buildings|Hydrogen","FE|Buildings|Gases|Hydrogen","FE|Buildings|Liquids|Hydrogen"), tgt2030 %in% tgt2030Scen) %>% group_by(scenario,tgt2030) %>% summarize(value = sum(value),.groups = "keep") %>% pull(value)),1),
    round(max(df %>% filter(region == "EU27", period == 2040, variable %in% c("FE|Transport|Hydrogen","FE|Transport|Gases|Hydrogen","FE|Transport|Liquids|Hydrogen"), tgt2030 %in% tgt2030Scen) %>% group_by(scenario,tgt2030) %>% summarize(value = sum(value),.groups = "keep") %>% pull(value)),1),
    round(max(df %>% filter(region == "EU27", period == 2040, variable %in% c("FE|Industry|Hydrogen","FE|Industry|Gases|Hydrogen","FE|Industry|Liquids|Hydrogen"), tgt2030 %in% tgt2030Scen) %>% group_by(scenario,tgt2030) %>% summarize(value = sum(value),.groups = "keep") %>% pull(value)),1)
  )
)

electricity_demand_increase <- left_join(
  df %>%
  filter(region == "EU27", period %in% c(2030,2040), variable %in% c("FE|Electricity"), tgt2050 != "Npi")
  ,
  df %>%
  filter(region == "EU27", period == 2020, variable %in% c("FE|Electricity"), tgt2050 != "Npi") %>%
  mutate(base = value,.groups = "keep") %>%
  select(scenario,base)
  , by = join_by(scenario == scenario)
) %>%
  mutate(upscale = value/base)

electricity_demand_upscale <- data.frame(
  wind_and_Solar_gen_upscale=c("2030","2040"),
  reference_value=c(
    round(electricity_demand_increase %>% filter(scenario == mainScen, period == "2030") %>% pull(upscale)*100),
    round(electricity_demand_increase %>% filter(scenario == mainScen, period == "2040") %>% pull(upscale)*100)
    ),
  min = c(
   round(min(electricity_demand_increase %>% filter(tgt2030 %in% tgt2030Scen, period == "2030") %>% pull(upscale)*100)),
   round(min(electricity_demand_increase %>% filter(tgt2030 %in% tgt2030Scen, period == "2040") %>% pull(upscale)*100))
   ),
  max = c(
   round(max(electricity_demand_increase %>% filter(tgt2030 %in% tgt2030Scen, period == "2030") %>% pull(upscale)*100)),
   round(max(electricity_demand_increase %>% filter(tgt2030 %in% tgt2030Scen, period == "2040") %>% pull(upscale)*100))
   )
)

electricity_share_in_FE <- data.frame(
  electricity_share_in_FE = c("2040"),
  reference_value=c(
    round(df %>% filter(region == "EU27", period == 2040, variable %in% c("Final Energy|Electricity Share"), scenario == mainScen) %>% pull(value)*100)
  ),
  min=c(
    round(min(df %>% filter(region == "EU27", period == 2040, variable %in% c("Final Energy|Electricity Share"), tgt2030 %in% tgt2030Scen, tgt2050 != "Npi") %>% pull(value)*100))
  ),
  max=c(
    round(max(df %>% filter(region == "EU27", period == 2040, variable %in% c("Final Energy|Electricity Share"), tgt2030 %in% tgt2030Scen, tgt2050 != "Npi") %>% pull(value)*100))
  )
)

wind_and_Solar_gen_upscale <- left_join(
  df %>%
  filter(region == "EU27", period %in% c(2030,2040), variable %in% c("SE|Electricity|Wind|Offshore", "SE|Electricity|Wind|Onshore", "SE|Electricity|Solar|PV", "SE|Electricity|Solar|CSP"), tgt2050 != "Npi") %>%
  group_by(scenario,tgt2030,period) %>%
  summarize(value = sum(value),.groups = "keep")
  ,
  df %>%
  filter(region == "EU27", period == 2020, variable %in% c("SE|Electricity|Wind|Offshore", "SE|Electricity|Wind|Onshore", "SE|Electricity|Solar|PV", "SE|Electricity|Solar|CSP"), tgt2050 != "Npi") %>%
  group_by(scenario,tgt2030) %>%
  summarize(base = sum(value),.groups = "keep")
  , by = join_by(scenario == scenario, tgt2030 == tgt2030)
) %>%
  mutate(upscale = value/base)

gen_upscale <- data.frame(
  wind_and_Solar_gen_upscale=c("2030","2040"),
  reference_value=c(
    round(wind_and_Solar_gen_upscale %>% filter(scenario == mainScen, period == "2030") %>% pull(upscale),1),
    round(wind_and_Solar_gen_upscale %>% filter(scenario == mainScen, period == "2040") %>% pull(upscale),1)
    ),
  min = c(
   round(min(wind_and_Solar_gen_upscale %>% filter(tgt2030 %in% tgt2030Scen, period == "2030") %>% pull(upscale)),1),
   round(min(wind_and_Solar_gen_upscale %>% filter(tgt2030 %in% tgt2030Scen, period == "2040") %>% pull(upscale)),1)
   ),
  max = c(
   round(max(wind_and_Solar_gen_upscale %>% filter(tgt2030 %in% tgt2030Scen, period == "2030") %>% pull(upscale)),1),
   round(max(wind_and_Solar_gen_upscale %>% filter(tgt2030 %in% tgt2030Scen, period == "2040") %>% pull(upscale)),1)
   )
)

coal_in_generation <- left_join(
  df %>%
  filter(region == "EU27", period == 2040, variable %in% c("SE|Electricity|Coal")) %>%  
  select(scenario,tgt2030,value)
  ,
  df %>% 
  filter(region == "EU27", period == 2040, variable %in% c("SE|Electricity")) %>%
  mutate(total = value) %>%  
  select(scenario,tgt2030,total)
  , by = join_by(scenario == scenario, tgt2030 == tgt2030)
  ) %>%
  mutate(percentage = value/total)

gas_in_generation <- left_join(
  df %>%
  filter(region == "EU27", period == 2040, variable %in% c("SE|Electricity|Gas|w/o CC","SE|Electricity|Gas|w/ CC")) %>%
  group_by(scenario,tgt2030) %>%
  summarize(value = sum(value),.groups = "keep")
  ,
  df %>% 
  filter(region == "EU27", period == 2040, variable %in% c("SE|Electricity")) %>%
  mutate(total = value) %>%  
  select(scenario,tgt2030,total)
  , by = join_by(scenario == scenario, tgt2030 == tgt2030)
  ) %>%
  mutate(percentage = value/total)

wind_and_Solar_gen <- left_join(
  df %>%
  filter(region == "EU27", period == 2040, variable %in% c("SE|Electricity|Wind|Offshore", "SE|Electricity|Wind|Onshore", "SE|Electricity|Solar|PV", "SE|Electricity|Solar|CSP")) %>%
  group_by(scenario,tgt2030) %>%
  summarize(value = sum(value),.groups = "keep")
  ,
  df %>% 
  filter(region == "EU27", period == 2040, variable %in% c("SE|Electricity")) %>%
  mutate(total = value) %>%  
  select(scenario,tgt2030,total)
  , by = join_by(scenario == scenario, tgt2030 == tgt2030)
  ) %>%
  mutate(percentage = value/total)

wind_and_Solar_cap <- df %>%
  filter(region == "EU27", period == 2040, variable %in% c("Cap|Electricity|Wind|Offshore", "Cap|Electricity|Wind|Onshore", "Cap|Electricity|Solar|PV", "Cap|Electricity|Solar|CSP")) %>%
  group_by(scenario,tgt2030) %>%
  summarize(value = sum(value),.groups = "keep")

nuclear_in_generation <- left_join(
  df %>%
  filter(region == "EU27", period == 2040, variable %in% c("SE|Electricity|Nuclear")) %>%  
  select(scenario,tgt2030,value)
  ,
  df %>% 
  filter(region == "EU27", period == 2040, variable %in% c("SE|Electricity")) %>%
  mutate(total = value) %>%  
  select(scenario,tgt2030,total)
  , by = join_by(scenario == scenario, tgt2030 == tgt2030)
  ) %>%
  mutate(percentage = value/total)

Generation <- data.frame(
  Electricity_Generation=c("coal_in_generation_(%)","gas_in_generation_(%)","nuclear_in_generation_(%)","wind_and_Solar_gen_(%)","wind_and_Solar_cap_(TW)"),
  reference_value=c(
    round(coal_in_generation %>% filter(scenario == mainScen) %>% pull(percentage)*100,1),
    round(gas_in_generation %>% filter(scenario == mainScen) %>% pull(percentage)*100,1),
    round(nuclear_in_generation %>% filter(scenario == mainScen) %>% pull(percentage)*100,1),
    round(wind_and_Solar_gen %>% filter(scenario == mainScen) %>% pull(percentage)*100,1),
    round(wind_and_Solar_cap %>% filter(scenario == mainScen) %>% pull(value)/ 1000,1)

    ),
  min = c(
   round(min(coal_in_generation %>% filter(tgt2030 %in% tgt2030Scen) %>% pull(percentage)*100),1),
   round(min(gas_in_generation %>% filter(tgt2030 %in% tgt2030Scen) %>% pull(percentage)*100),1),
   round(min(nuclear_in_generation %>% filter(tgt2030 %in% tgt2030Scen) %>% pull(percentage)*100),1),
   round(min(wind_and_Solar_gen %>% filter(tgt2030 %in% tgt2030Scen) %>% pull(percentage)*100),1),
   round(min(wind_and_Solar_cap %>% filter(tgt2030 %in% tgt2030Scen) %>% pull(value)/ 1000),1)
   ),
  max = c(
   round(max(coal_in_generation %>% filter(tgt2030 %in% tgt2030Scen) %>% pull(percentage)*100),1),
   round(max(gas_in_generation %>% filter(tgt2030 %in% tgt2030Scen) %>% pull(percentage)*100),1),
   round(max(nuclear_in_generation %>% filter(tgt2030 %in% tgt2030Scen) %>% pull(percentage)*100),1),
   round(max(wind_and_Solar_gen %>% filter(tgt2030 %in% tgt2030Scen) %>% pull(percentage)*100),1),
   round(max(wind_and_Solar_cap %>% filter(tgt2030 %in% tgt2030Scen) %>% pull(value)/ 1000),1)
   )
)

share_of_fossil_in_electricity <- data.frame(
  FE=c("share_of_fossil_without_CCS_in_electricity_(%)"),
  reference_value_2020=c(
    round((1-df %>% filter(region == "EU27", period == 2020, variable == "SE|Electricity|carbon-free", tgt2030 %in% tgt2030Scen, scenario == mainScen) %>% pull(value))*100 )),
  reference_value_2040=c(
    round((1-df %>% filter(region == "EU27", period == 2040, variable == "SE|Electricity|carbon-free", tgt2030 %in% tgt2030Scen, scenario == mainScen) %>% pull(value))*100 )),
  min = c(round(min((1-df %>% filter(region == "EU27", period == 2040, variable == "SE|Electricity|carbon-free", tgt2030 %in% tgt2030Scen) %>% pull(value))*100))),
  max = c(round(max((1-df %>% filter(region == "EU27", period == 2040, variable == "SE|Electricity|carbon-free", tgt2030 %in% tgt2030Scen) %>% pull(value))*100)))
)

pe <- df %>% 
  filter(region == "EU27", period %in% c(2020:2060), tgt2030 %in% tgt2030Scen, variable %in% c("PE|Biomass","PE|Coal","PE|Gas","PE|Oil"))

primaryEnergy <- data.frame(
  Primary_Energy=c("Biomass","Coal","Gas","Oil"),
  relative_to_2020=c(
    round((pe %>% filter(period == 2040, scenario == mainScen, variable == "PE|Biomass") %>% mutate(value = value / (pe %>% filter(period == 2020, scenario == mainScen, variable == "PE|Biomass") %>% pull(value)))%>% pull(value))*100, 1),
    round((pe %>% filter(period == 2040, scenario == mainScen, variable == "PE|Coal") %>% mutate(value = value / (pe %>% filter(period == 2020, scenario == mainScen, variable == "PE|Coal") %>% pull(value)))%>% pull(value))*100, 1),
    round((pe %>% filter(period == 2040, scenario == mainScen, variable == "PE|Gas") %>% mutate(value = value / (pe %>% filter(period == 2020, scenario == mainScen, variable == "PE|Gas") %>% pull(value)))%>% pull(value))*100, 1),
    round((pe %>% filter(period == 2040, scenario == mainScen, variable == "PE|Oil") %>% mutate(value = value / (pe %>% filter(period == 2020, scenario == mainScen, variable == "PE|Oil") %>% pull(value)))%>% pull(value))*100, 1))
)

fe_transport_electricity <- left_join(
  df %>%
  filter(region == "EU27", period == 2040, variable %in% c("FE|Transport|Electricity")) %>%  
  select(scenario,tgt2030,value)
  ,
  df %>% 
  filter(region == "EU27", period == 2030, variable %in% c("FE|Transport|Electricity")) %>%
  mutate(base = value) %>%  
  select(scenario,tgt2030,base)
  , by = join_by(scenario == scenario, tgt2030 == tgt2030)
  ) %>%
  mutate(percentage = value/base)

FE <- data.frame(
  FE=c("transport_electricity_(%)"),
  reference_value=c(round(fe_transport_electricity %>% filter(scenario == mainScen) %>% pull(percentage)*100,1)),
  min = c(round(min(fe_transport_electricity %>% filter(tgt2030 %in% tgt2030Scen) %>% pull(percentage)*100),1)),
  max = c(round(max(fe_transport_electricity %>% filter(tgt2030 %in% tgt2030Scen) %>% pull(percentage)*100),1))
)

Bunkers <- list(
  "Aviation" = c(
      "FE|Transport|Pass|Aviation|International|Electricity" = "Electricity",
      "FE|Transport|Pass|Aviation|International|Gases" = "Gases",
      "FE|Transport|Pass|Aviation|International|Hydrogen" = "Hydrogen",
#     "FE|Transport|Pass|Aviation|International|Liquids" = "Liquids"
      "FE|Transport|Pass|Aviation|International|Liquids|Fossil" = "Liquids - Fossil",
      "FE|Transport|Pass|Aviation|International|Liquids|Biomass" = "Liquids - Biomass",
      "FE|Transport|Pass|Aviation|International|Liquids|Hydrogen" = "Liquids - Synthetic"
    ),
    "Shipping" = c(
      "FE|Transport|Freight|Navigation|Electricity" = "Electricity",
      "FE|Transport|Freight|Navigation|Gases" = "Gases",
      "FE|Transport|Freight|Navigation|Hydrogen" = "Hydrogen",
#     "FE|Transport|Freight|Navigation|Liquids" = "Liquids",
      "FE|Transport|Freight|Navigation|Liquids|Fossil" = "Liquids - Fossil",
      "FE|Transport|Freight|Navigation|Liquids|Biomass" = "Liquids - Biomass",
      "FE|Transport|Freight|Navigation|Liquids|Hydrogen" = "Liquids - Synthetic",
      "FE|Transport|Freight|International Shipping|Electricity" = "Electricity",
      "FE|Transport|Freight|International Shipping|Gases" = "Gases",
      "FE|Transport|Freight|International Shipping|Hydrogen" = "Hydrogen",
#     "FE|Transport|Freight|International Shipping|Liquids" = "Liquids"
      "FE|Transport|Freight|International Shipping|Liquids|Fossil" = "Liquids - Fossil",
      "FE|Transport|Freight|International Shipping|Liquids|Biomass" = "Liquids - Biomass",
      "FE|Transport|Freight|International Shipping|Liquids|Hydrogen" = "Liquids - Synthetic"
    )
)

cleanAviation <- c("FE|Transport|Pass|Aviation|International|Electricity", "FE|Transport|Pass|Aviation|International|Hydrogen", "FE|Transport|Pass|Aviation|International|Liquids|Biomass", "FE|Transport|Pass|Aviation|International|Liquids|Hydrogen")

AviationData <- left_join(
  df %>% 
  filter(region == "EU27", period == 2040, variable %in% cleanAviation) %>%
  group_by(scenario,tgt2030) %>%
  summarize(value = sum(value),.groups = "keep")
  ,
  df %>% 
  filter(region == "EU27", period == 2040, variable %in% c(names(Bunkers$Aviation))) %>%
  group_by(scenario,tgt2030) %>%
  summarize(total = sum(value),.groups = "keep")
  , by = join_by(scenario == scenario, tgt2030 == tgt2030)
) %>%
  mutate(percentage = value/total)

cleanShipping <- c("FE|Transport|Freight|Navigation|Electricity", "FE|Transport|Freight|Navigation|Hydrogen", "FE|Transport|Freight|Navigation|Liquids|Biomass",  "FE|Transport|Freight|Navigation|Liquids|Hydrogen", "FE|Transport|Freight|International Shipping|Electricity", "FE|Transport|Freight|International Shipping|Hydrogen", "FE|Transport|Freight|International Shipping|Liquids|Biomass", "FE|Transport|Freight|International Shipping|Liquids|Hydrogen")

ShippingData <- left_join(
  df %>% 
  filter(region == "EU27", period == 2040, variable %in% cleanShipping) %>%
  group_by(scenario,tgt2030) %>%
  summarize(value = sum(value),.groups = "keep")
  ,
  df %>% 
  filter(region == "EU27", period == 2040, variable %in% c(names(Bunkers$Shipping))) %>%
  group_by(scenario,tgt2030) %>%
  summarize(total = sum(value),.groups = "keep")
  , by = join_by(scenario == scenario, tgt2030 == tgt2030)
) %>%
  mutate(percentage = value/total)

share_of_sustainable_fuel <- data.frame(
  FE=c("share_of_sustainable_fuel_in_aviation_(%)","share_of_sustainable_fuel_in_shipping_(%)"),
  reference_value=c(round(AviationData %>% filter(scenario == mainScen) %>% pull(percentage)*100,1),
                    round(ShippingData %>% filter(scenario == mainScen) %>% pull(percentage)*100,1)),
  min = c(round(min(AviationData %>% filter(tgt2030 %in% tgt2030Scen) %>% pull(percentage)*100),1),
          round(min(ShippingData %>% filter(tgt2030 %in% tgt2030Scen) %>% pull(percentage)*100),1)),
  max = c(round(max(AviationData %>% filter(tgt2030 %in% tgt2030Scen) %>% pull(percentage)*100),1),
          round(max(ShippingData %>% filter(tgt2030 %in% tgt2030Scen) %>% pull(percentage)*100),1))
)

build_UE_heat <- c(
  "Heat pump" = "UE|Buildings|Heating|Electricity|Heat pump",
  "District Heating" = "UE|Buildings|Heating|District Heating",
  "Resistance" = "UE|Buildings|Heating|Electricity|Resistance",
  "Hydrogen" = "UE|Buildings|Heating|Hydrogen",
  "Gases" = "UE|Buildings|Heating|Gases",
  "Liquids" = "UE|Buildings|Heating|Liquids",
  "Solids" = "UE|Buildings|Heating|Solids"#,
#  "non-Heating Electricity" = "UE|Buildings|non-Heating|Electricity|Conventional"
)

build_UE_heating_data <- left_join(
  df %>% 
  filter(region == "EU27", period == 2040, variable %in% build_UE_heat[c("Heat pump","District Heating","Resistance")], tgt2030 %in% tgt2030Scen) %>%
  group_by(scenario) %>%    
  summarize(value = sum(value),.groups = "keep")
  ,
  df %>% 
  filter(region == "EU27", period == 2040, variable %in% build_UE_heat, tgt2030 %in% tgt2030Scen) %>%
  group_by(scenario) %>%  
  summarize(total = sum(value)) %>% 
  select(scenario,total)  
  ,
  by = join_by(scenario == scenario)
) %>% 
  mutate(pctg = value/total)

build_UE_liq_gas_data <- left_join(
  df %>% 
  filter(region == "EU27", period == 2040, variable %in% build_UE_heat[c("Gases","Hydrogen","Liquids")], tgt2030 %in% tgt2030Scen) %>%
  group_by(scenario) %>%    
  summarize(value = sum(value),.groups = "keep")
  ,
  df %>% 
  filter(region == "EU27", period == 2040, variable %in% build_UE_heat, tgt2030 %in% tgt2030Scen) %>%
  group_by(scenario) %>%  
  summarize(total = sum(value)) %>% 
  select(scenario,total)  
  ,
  by = join_by(scenario == scenario)
) %>% 
  mutate(pctg = value/total)

build_UE_solids_data <- left_join(
  df %>% 
  filter(region == "EU27", period == 2040, variable %in% build_UE_heat[c("Solids")], tgt2030 %in% tgt2030Scen) %>%
  group_by(scenario) %>%    
  summarize(value = sum(value),.groups = "keep")
  ,
  df %>% 
  filter(region == "EU27", period == 2040, variable %in% build_UE_heat, tgt2030 %in% tgt2030Scen) %>%
  group_by(scenario) %>%  
  summarize(total = sum(value)) %>% 
  select(scenario,total)  
  ,
  by = join_by(scenario == scenario)
) %>% 
  mutate(pctg = value/total)

share_of_build_UE_heating <- data.frame(
  FE=c("share_of_electric_and_district_buildings_heating_useful energy_(%)","share_of_liquids_and_gases_buildings_heating_useful_energy_(%)","share_of_solids_buildings_heating_useful energy_(%)"),
  reference_value=c(round(build_UE_heating_data %>% filter(scenario == mainScen) %>% summarize(total = sum(pctg),.groups = "keep") %>% pull(total)*100 ),
                    round(build_UE_liq_gas_data %>% filter(scenario == mainScen) %>% summarize(total = sum(pctg),.groups = "keep") %>% pull(total)*100 ),
                    round(build_UE_solids_data %>% filter(scenario == mainScen) %>% summarize(total = sum(pctg),.groups = "keep") %>% pull(total)*100 )),
  min = c(round(min(build_UE_heating_data %>% pull(pctg) *100)),
          round(min(build_UE_liq_gas_data %>% pull(pctg) *100)),
          round(min(build_UE_solids_data %>% pull(pctg) *100))),
  max = c(round(max(build_UE_heating_data %>% pull(pctg) *100)),
          round(max(build_UE_liq_gas_data %>% pull(pctg) *100)),
          round(max(build_UE_solids_data %>% pull(pctg) *100)))
)

build_FE_data <- left_join(
  df %>% 
  filter(region == "EU27", period == 2040, variable == "FE|Buildings|Heating", tgt2030 %in% tgt2030Scen) %>% 
  select(scenario,value)  
  ,
  df %>% 
  filter(region == "EU27", period == 2020, variable == "FE|Buildings|Heating", tgt2030 %in% tgt2030Scen) %>%
  mutate(base = value) %>% 
  select(scenario,base)  
  ,
  by = join_by(scenario == scenario)
) %>% 
  mutate(pctg = value/base)

share_of_build_FE_in_relation_to_2020 <- data.frame(
  FE=c("share_of_build_FE_in_relation_to_2020_(%)"),
  reference_value=c(round((1-build_FE_data %>% filter(scenario == mainScen) %>% pull(pctg))*100 )),
  min = c(round(min((1-build_FE_data %>% pull(pctg)) *100))),
  max = c(round(max((1-build_FE_data %>% pull(pctg)) *100)))
)

```

```{r, echo=FALSE, fig.width=12, fig.height=5}

indirect_electrification %>% 
  kbl() %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"))

electricity_demand_upscale %>% 
  kbl() %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"))

electricity_share_in_FE %>% 
  kbl() %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"))

gen_upscale %>% 
  kbl() %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"))

Generation %>% 
  kbl() %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"))

share_of_fossil_in_electricity %>% 
  kbl() %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"))

primaryEnergy %>% 
  kbl() %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"))

FE %>% 
  kbl() %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"))

share_of_sustainable_fuel %>% 
  kbl() %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"))

share_of_build_UE_heating %>% 
  kbl() %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"))

share_of_build_FE_in_relation_to_2020 %>% 
  kbl() %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"))

```

---

## Carbon Management Results

```{r, echo=FALSE}

#Carbon Management|Storage in 2030 (Minimum and Maximum of your scenario range)

#Carbon Management|Storage in 2040 (Minimum and Maximum of your scenario range)

#tenth root (because the period is 10 years) and then the division of 2040 CCS to 2030 CCS, i.e.:
#10√(Carbon Management|Storage in 2040 / Carbon Management|Storage in 2030)
#(here also minimum and maximum value across your scenario range)

#Here again we need the tenth (because 10 years) root of the ratio of CDR in 2040 divided by CDR in 2030
#For CDR quantity calculation see next comment :)

#Total CDR = (Emi|CO2|CDR|BECCS + Emi|CO2|CDR|DACCS + Emi|CO2|CDR|Industry CCS|Synthetic Fuels) in 2040
#(and then span the scenario range)

totalCDR <- df %>% filter(region == "EU27", period %in% c(2030,2040), variable %in% c("Emi|CO2|CDR|BECCS","Emi|CO2|CDR|DACCS","Emi|CO2|CDR|Industry CCS|Synthetic Fuels"), tgt2030 %in% tgt2030Scen) %>% group_by(period,scenario) %>% summarize(value = sum(-value),.groups = "keep")

rateCCS <- left_join(
  df %>%
    filter(region == "EU27", period == 2040, variable == "Carbon Management|Storage", tgt2030 %in% tgt2030Scen) %>%
    mutate(numerator=value) %>%
    select(scenario,numerator)
  ,
  df %>%
    filter(region == "EU27", period == 2030, variable == "Carbon Management|Storage", tgt2030 %in% tgt2030Scen) %>%
    mutate(denominator=value) %>%
    select(scenario,denominator)
  , by = join_by(scenario == scenario)
) %>%
  mutate(percentage = numerator/denominator,
         rate = percentage ^ (1/10))

rateCDR <- left_join(
  totalCDR %>%
    filter(period == 2040) %>%
    mutate(numerator=value) %>%
    ungroup() %>%
    select(scenario,numerator)
  ,
  totalCDR %>%
    filter(period == 2030) %>%
    mutate(denominator=value) %>%
    ungroup() %>%
    select(scenario,denominator)
  , by = join_by(scenario == scenario)
) %>%
  mutate(percentage = numerator/denominator,
         rate = percentage ^ (1/10))

CDR <- data.frame(
  result=c("CCS_deployment_in_2030_MtCO2","CCS_deployment_in_2040_MtCO2","CCS_upscaling_rates_2030_2040_%","novel_CDR_upscaling_rates_2030_2040_%","novel_CDR_deployment_in_2040_MtCO2"),
  reference_value=c(
    round(df %>% filter(region == "EU27", period == 2030, variable == "Carbon Management|Storage", scenario == mainScen) %>% pull(value)),
    round(df %>% filter(region == "EU27", period == 2040, variable == "Carbon Management|Storage", scenario == mainScen) %>% pull(value)),
    round((rateCCS %>% filter(scenario == mainScen) %>% pull(rate)-1)*100),
    round((rateCDR %>% filter(scenario == mainScen) %>% pull(rate)-1)*100),
    round(totalCDR %>% filter(period == 2040, scenario == mainScen) %>% pull(value))
  ),
  min = c(
    round(min(df %>% filter(region == "EU27", period == 2030, variable == "Carbon Management|Storage", tgt2030 %in% tgt2030Scen) %>% pull(value))),
    round(min(df %>% filter(region == "EU27", period == 2040, variable == "Carbon Management|Storage", tgt2030 %in% tgt2030Scen) %>% pull(value))),
    round((min(rateCCS %>% pull(rate))-1)*100),
    round((min(rateCDR %>% pull(rate))-1)*100),
    round(min(totalCDR %>% pull(value)))
  ),
  max = c(
    round(max(df %>% filter(region == "EU27", period == 2030, variable == "Carbon Management|Storage", tgt2030 %in% tgt2030Scen) %>% pull(value))),
    round(max(df %>% filter(region == "EU27", period == 2040, variable == "Carbon Management|Storage", tgt2030 %in% tgt2030Scen) %>% pull(value))),
    round((max(rateCCS %>% pull(rate))-1)*100),
    round((max(rateCDR %>% pull(rate))-1)*100),
    round(max(totalCDR %>% pull(value)))
  )
)

```

```{r, echo=FALSE, fig.width=12, fig.height=5}

CDR %>% 
  kbl() %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"))

```

---

<!-- Paper Figures -->
```{r, echo=FALSE}
### Initialize plot objects
g <- NULL

```

# Figure 1. Emission results panel

(a) Total GHG emissions (Mt CO2eq/yr)                    (b) Emission reductions (%)

```{r, echo=FALSE}

# all emission reductions were calculated relative to 1990 emissions with LULUCF and international transport (4814 Mt CO2e), except for 2030 emission reduction calculated relative to 1990 emissions with LULUCF and intra-EU aviation only (4713 Mt CO2e).

plotData <- df %>% 
  filter(region == "EU27", period %in% c(2020:2060), tgt2030 %in% tgt2030Scen) %>% 
  calc_addVariable( "`Emissions with LULUCF and with international transport`" = "`Emi|GHG|LULUCF national accounting` + `Emi|CO2|Energy|Demand|Transport|International Bunkers`" , units = "Mt CO2/yr", only.new=T)  %>%
  group_by(model,region,variable,period) %>% 
  filter(tgt2050 != "Npi") %>% 
  mutate(ribbon.max=max(value),
         ribbon.min=min(value))

minValue <- min(plotData$value)

h <- refEmi %>% filter(variable == "Emissions with LULUCF and with international transport", region == "EU27", pollutant == "All greenhouse gases - (CO2 equivalent)") %>% ungroup() %>% select(region,variable,period,value)

emi1990 <- tgt %>% filter(target == "Emissions with LULUCF and with international transport") %>% pull(reference)

g[["GHG emi"]] <- ggplot(plotData, aes(x=period,y=value)) +
    geom_hline(yintercept=0, color = "black", linewidth=2, alpha = 0.5) +
    geom_vline(xintercept=2020, color = "black", linewidth=1.5, alpha = 0.5,linetype="dashed") +
     geom_ribbon(aes(ymax = ribbon.max, ymin = ribbon.min), alpha = 0.5, fill = "#545454") + 
     geom_segment(data=plotData %>% filter(scenario == mainScen, period == 2040), aes(x=2040, xend=2040, y=0, yend=value), linetype="dashed", color="black") +
     geom_line(data=plotData %>% filter(scenario == mainScen), linewidth=1.5) +
     #geom_line(data=plotData, color= "#545454", linewidth=0.2, alpha=0.2, aes(group=scenario)) +
     geom_line(data=h,linewidth=0.7,linetype="dashed") +
     geom_point(data=plotData %>% filter(scenario == mainScen, period == 2040),size=3,shape=21,fill="white",stroke=1.5) +
     theme_doubleColumn() +
#    ggtitle("GHG Emissions") +
    theme(axis.title.y=element_blank()) +
     expand_limits(y=emi1990) +
#     ylab(expression(paste("Mt ", CO[2],"eq/yr"))) +
     scale_y_continuous(breaks = c(0,1000,2000,3000,4000,5000)) +
  scale_x_continuous(breaks = seq(1990,2060,10))

plotData <- rbind(
  df %>% 
    filter(region == "EU27", period %in% c(2030:2040), tgt2030 %in% tgt2030Scen) %>%
    calc_addVariable( "`Emissions with LULUCF and with intra-EU aviation`" = "`Emi|GHG|LULUCF national accounting` + `Emi|CO2|Energy|Demand|Transport|International Bunkers|Intra-region`" , units = "Mt CO2/yr", only.new=T) %>% 
    mutate(reduction = round(value / tgt %>% filter(target == "Emissions with LULUCF and with intra-EU aviation") %>% pull(reference) ,2)) %>% 
    select(-variable)
  ,
  df %>% 
    filter(region == "EU27", period %in% c(2045:2050), tgt2030 %in% tgt2030Scen) %>%
    calc_addVariable( "`Emissions with LULUCF and with international transport`" = "`Emi|GHG|LULUCF national accounting` + `Emi|CO2|Energy|Demand|Transport|International Bunkers`" , units = "Mt CO2/yr", only.new=T) %>% 
    mutate(reduction = round(value / tgt %>% filter(target == "Emissions with LULUCF and with international transport") %>% pull(reference) ,2)) %>% 
    select(-variable)
  ) %>%
  filter(tgt2050 != "Npi") %>%
  mutate(x=ifelse(reduction>0,period+1,NA),
         xend=ifelse(reduction>0,period+1,NA),
         y=ifelse(reduction>0,1,NA),
         yend=ifelse(reduction>0,ifelse(reduction<=0,0,reduction),NA),
         bar = ifelse(period==2040,"2040","emi"),
         #label = ifelse(value>0,paste0( "- ", (1-value) * 100, "% (-", (1-intraEURed) * 100 ,"%)"),NA),
         labelY = reduction + (1-reduction)/2,
         labelX=period+1) %>%
  group_by(model,region,period) %>% 
  mutate(max=max(reduction),
         min=min(reduction),
         range=max-min,
         label= paste0(round((1-max) * 100), "% to ", round((1-min) * 100) ,"%"))

g[["target"]] <- ggplot(data=plotData,aes(x=period,y=reduction,fill=bar)) +
  geom_bar(data=plotData %>% filter(scenario == mainScen), stat='identity', show.legend = FALSE, alpha = 0.7, color="black") +
  geom_errorbar(aes(ymin=min, ymax=max), width=1, colour="black", alpha=0.9, linewidth=1.3) +
  geom_hline(yintercept=0, color = "black", linewidth=2, alpha = 0.5) +
  theme_doubleColumn() +
  scale_fill_manual(values = c("emi" = "black", "2040" = "#5a788b")) + 
  guides(y = "none") +
  theme(axis.title.y=element_blank()) +
  geom_segment(data=plotData %>% filter(scenario == mainScen), aes(x=x, y=y, xend=xend, yend=yend), arrow = arrow(length=unit(.5, 'cm')), na.rm=TRUE) +
  geom_text(data=plotData %>% filter(scenario == mainScen), aes(label = label, y=labelY, x=labelX, color=factor(period)), vjust=-1, hjust=0.5, na.rm=TRUE, size = geom.text.single.size, angle = 90) +
  scale_color_manual(values = c("2030" = "black", "2035" = "#777777","2040" = "black", "2045" = "#777777","2050" = "black"), breaks=c("2030", "2035","2040", "2045","2050")) +
  expand_limits(y=minValue/emi1990) +
  annotate("text", x=2052, y=1, label= "1990 emissions", vjust=-1, hjust=1, size = geom.text.single.size) +
  geom_hline(yintercept=1, color = "black", linetype="dashed", linewidth=2, alpha = 0.5) +
  coord_cartesian(clip = 'off') +
  scale_y_continuous(breaks = c(0,1000/emi1990,2000/emi1990,3000/emi1990,4000/emi1990,5000/emi1990)) + 
  theme(legend.position = "none")

g[["GHG Emission reductions"]] <- gridExtra::arrangeGrob(
  g[["GHG emi"]], #+ theme(plot.margin=margin(10,10,30,10)),
  g[["target"]], # + theme(plot.margin=margin(10,10,10,10)),
  layout_matrix = rbind(c(1, 1, 1, 1, 2, 2),
                          c(1, 1, 1, 1, 2, 2))
   )

 ggsave(paste0(outFolder,"/GHG Emission reductions.svg"), g[["GHG Emission reductions"]] , device="svg", width = 12, height = 5, dpi=100, units = "in")
 ggsave(paste0(outFolder,"/GHG Emission reductions.png"), g[["GHG Emission reductions"]] , device="png", width = 12, height = 5, dpi=300, units = "in")

```

```{r, echo=FALSE, fig.width=12, fig.height=5}

grid.newpage()
grid.draw(g[["GHG Emission reductions"]])

```

(c) Breakdown of residual CO2 emissions from energy and industrial processes (% in relation to 2020)

```{r, echo=FALSE}

#df <- rbind(df,df %>% calc_addVariable("`Emi|CO2|Gross|Other Energy Conversion`" = "`Emi|CO2|Gross|Energy|Supply|Hydrogen` + `Emi|CO2|Gross|Energy|Supply|Liquids` #+ `Emi|CO2|Gross|Energy|Supply|Liquids` + `Emi|CO2|Gross|Energy|Supply|Solids`", units = "Mt CO2/yr", only.new=T))

vars <- list(
  "Industrial\nProcesses" = "Emi|GHG|Industrial Processes",
  "Industry" = "Emi|GHG|Gross|Energy|Demand|Industry",
  "Buildings" ="Emi|GHG|Energy|Demand|Buildings",
  "Transportation" ="Emi|GHG|Energy|Demand|Transport",
  "Bunkers" = "Emi|CO2|Energy|Demand|Transport|International Bunkers",
#  "CDR" = "Emi|GHG|Gross|Energy|Demand|CDR",
#  "AFOFI" = "Emissions|CO2|Energy|Demand|AFOFI",
#  "Other Sector" ="Emissions|CO2|Energy|Demand|Other Sector",
  "Heat"="Emi|CO2|Gross|Energy|Supply|Heat",
  "Other Energy\nConversion" = "Emi|CO2|Gross|Other Energy Conversion",
    #"Hydrogen" ="Emi|CO2|Gross|Energy|Supply|Hydrogen",
    #"Liquids"="Emi|CO2|Gross|Energy|Supply|Liquids",
    #"Gases"="Emi|CO2|Gross|Energy|Supply|Gases",
    #"Solids"="Emi|CO2|Gross|Energy|Supply|Solids",
#  "Other"="Emissions|CO2|Energy|Supply|Other",
  "Electricity"="Emi|GHG|Gross|Energy|Supply|Electricity"
)
# "Emi|GHG|Agriculture",
#   "Emi|GHG|Land-Use Change",
#   "Emi|GHG|Waste",
#   "Emi|CO2|CDR|BECCS",
#   "Emi|CO2|CDR|Industry CCS|Synthetic Fuels",
#   "Emi|CO2|CDR|DACCS",
#   "Emi|CO2|CDR|EW"

plotColor <- c(
  "Electricity"= "#ffb200",
  "Heat"= "#cc0000",
  "Other Energy\nConversion" = "#0000cc",
  #"Gases"= "#999959",
  #"Liquids"= "#0000cc",
  #"Solids"= "#0c0c0c",
  #"Other" = "#79a6d2",
  #"Hydrogen" = "#5ed5b0",
  "Transportation" = "#a1dd70",
  "Bunkers" = "#ffc0cb",
  "Buildings" = "#5edfff",
  "Other Sector" = "#999959",
  "AFOFI" = "#005900",
  "Industry" = "#5f6769",
  "Industrial\nProcesses" = "#5c5c5c",
  "Other" = "#5b0d8f",
  "Waste" = "#A52A2A",
  "AFOLU" = "#4DAF4A",
  "Agriculture" = "#357933",
  "LULUCF" = "#6ec16b"
)

labels <- setNames(names(vars),vars[names(vars)])

fulldata <- df %>%
  filter(region == "EU27",
         period %in% c(2020,2030,2040),
         tgt2030 %in% tgt2030Scen,
         variable %in% unlist(vars)) %>%
  mutate(labels = factor(labels[as.character(variable)],levels = names(vars)),
         variable = factor(variable, levels = vars)) %>%
  group_by(period,variable) %>%
  mutate(min = min(value),
         max = max(value))

plotData <- fulldata %>%
  filter(scenario == mainScen) %>%
  mutate(complement = FALSE) %>%
  group_by(variable) %>%
  mutate(pct = paste0(round(value / value[period==2020],2)*100, "%")) %>%
  bind_rows(
    fulldata %>%
      filter(scenario == mainScen) %>%
      mutate(complement = TRUE) %>%
      group_by(variable) %>%
      mutate(value = value[period==2020] - value) %>%
      mutate(labels=NA)
  ) %>%
  mutate(ext_Var = ifelse(complement, paste0(as.character(variable),"_"), as.character(variable))) %>%
  arrange(period,ext_Var) %>%
  group_by(period) %>%
  arrange(variable) %>%
  mutate(ymid = as.numeric(sub("\\D+", "", period)),
         ymax = ymid + 4, ymin = ymid - 4,
         xmin = c(0, head(cumsum(value), -1)),
         xmax = cumsum(value),
         xmid = (xmax + xmin) / 2,
         ) %>%
  ungroup() %>%
  filter(complement == FALSE)

g[["2040 Emissions - Energy and Industrial Processes - horizontal bar"]] <- plotData %>% 
  ggplot(aes(xmid, ymid, fill = ext_Var)) +
  geom_rect(aes(xmin = xmin, xmax = xmax, ymin = ymin, ymax = ymax,
                alpha = factor(period), color = variable)) +
  geom_errorbar(data=plotData %>% filter(period == 2040), aes(xmin=xmax-value+min, xmax=xmax-value+max), width=4, colour="black", alpha=0.9, linewidth=0.8) +
  geom_text(data=plotData %>% filter(period == 2020), aes(y = 2015, label = labels, group = labels, hjust=1, angle = 90), vjust=0.5, hjust=1.1, na.rm=TRUE, size = geom.text.single.size, angle = 90, lineheight = 0.8) +
  geom_text(data=plotData %>% filter(period == 2040), aes(x=xmid , y = 2045, label = pct), vjust=0.5, hjust=-0.1, na.rm=TRUE, size = geom.text.single.size, angle = 90) +
  scale_fill_manual(values = c(setNames(plotColor[names(vars)],vars)), labels = setNames(names(vars), vars)) + 
  scale_alpha_manual(values = c(0.2, 0.3, 1)) +
  scale_color_manual(values = c(setNames(plotColor[names(vars)],vars)), labels = setNames(names(vars), vars)) +
  theme_doubleColumn() +
  theme(legend.position = "none") +
  scale_y_continuous(breaks=c(2020,2030,2040), labels=c("today",2030,2040), minor_breaks = NULL, limits=c(2008,2047)) +
  theme(axis.title.x=element_blank(),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank(),
        axis.title.y=element_blank()) +
  scale_x_continuous(breaks = NULL) +
  coord_cartesian(clip="off") +
  theme(plot.margin = margin(t = 30, r = 0, b = 85, l = 0, unit = "pt"))

ggsave(paste0(outFolder,"/2040 Emissions - Energy and Industrial Processes - horizontal bar.svg"), g[["2040 Emissions - Energy and Industrial Processes - horizontal bar"]] , device="svg", width = 12, height = 3.2, dpi=100, units = "in")
ggsave(paste0(outFolder,"/2040 Emissions - Energy and Industrial Processes - horizontal bar.png"), g[["2040 Emissions - Energy and Industrial Processes - horizontal bar"]] , device="png", width = 12, height = 3.2, dpi=300, units = "in")

```

```{r, echo=FALSE, fig.width=12, fig.height=3.2}

  g[["2040 Emissions - Energy and Industrial Processes - horizontal bar"]]

```

Figure 1. Emission results panel. Reference scenario: Net-zero by 2050, 55% reduction by 2030, REMIND future final energy projections, biomass availability limited to 7.5 EJ/yr and default CCS assumption.

---

# Figure 2. Electricity panel

:::: {style="display: flex;"}

::: {}

(a) Electricity Generation (%)

```{r, echo=FALSE}

vars <- c(
  "Net Imports" = "SE|Electricity|Net Imports",
  "Coal w/ CC" = "SE|Electricity|Coal|w/ CC",
  "Coal w/o CC" = "SE|Electricity|Coal|w/o CC",
  "Oil" = "SE|Electricity|Oil",
  "Gas w/ CC" = "SE|Electricity|Gas|w/ CC",
  "Gas w/o CC" = "SE|Electricity|Gas|w/o CC",
  "Hydrogen" = "SE|Electricity|Hydrogen",
  "Nuclear" = "SE|Electricity|Nuclear",
  "Geothermal" = "SE|Electricity|Geothermal",
  "Hydro" = "SE|Electricity|Hydro",
  "Biomass w/ CC" = "SE|Electricity|Biomass|w/ CC",
  "Biomass w/o CC" = "SE|Electricity|Biomass|w/o CC",
  "Solar CSP" = "SE|Electricity|Solar|CSP",
  "Solar PV" = "SE|Electricity|Solar|PV",
  "Wind Onshore" = "SE|Electricity|Wind|Onshore",
  "Wind Offshore" = "SE|Electricity|Wind|Offshore")

inv_vars <- setNames(names(vars),vars)

color <- c(
  "Net Imports"   = "#472b62",
  "Coal w/ CC"    = "#b2b2b2",
  "Coal w/o CC"   = "#0c0c0c",
  "Oil"           = "#b30000",
  "Gas w/ CC"     = "#b79e38",
  "Gas w/o CC"    = "#999959",
  "Hydrogen"      = "#5ed5b0",
  "Nuclear"       = "#ff33ff",
  "Geothermal"    = "#e51900",
  "Hydro"         = "#191999",
  "Biomass w/ CC" = "#33ff00",
  "Biomass w/o CC"= "#005900",
  "Solar CSP"     = "#ffcc00",
  "Solar PV"      = "#ffe600",
  "Wind Onshore"  = "#193F7F",
  "Wind Offshore" = "#337fff")

plotData <- df %>% 
  filter(region == "EU27", period >= 2020, period <= 2060, scenario  == mainScen, variable %in% unlist(vars)) %>% 
  mutate(variable = factor(variable, levels=vars)) %>% 
  arrange(factor(variable, levels = rev(vars))) %>%
  group_by(model,scenario,region,period) %>%
  mutate(cumValue = cumsum(value),
         n = sum(value),
         percentage = value / n,
         cumPercentage = cumValue / n)

plotData2040 <- plotData[plotData$period == 2040,]
#plotData$value <- plotData$value / plotData[plotData$period == "2020",]$value
#plotData$period <- as.character(plotData$period)
#plotData$period <- factor(plotData$period , levels = rev(c("2020","2030","2040","2050")))

plotData2040$xPos <- 2042.5
#plotData2040[plotData2040$variable %in% c("SE|Electricity|Hydro"),]$xPos <- 2028.5
#plotData2040[plotData2040$variable %in% c("SE|Electricity|Biomass|w/o CC"),]$xPos <- 2020
plotData2040[plotData2040$variable %in% c("SE|Electricity|Nuclear","SE|Electricity|Biomass|w/o CC"),]$xPos <- 2035

plotDataRenewables <- plotData %>%
  filter(variable %in% c("SE|Electricity|Geothermal","SE|Electricity|Hydro","SE|Electricity|Biomass|w/ CC","SE|Electricity|Biomass|w/o CC","SE|Electricity|Solar|CSP","SE|Electricity|Solar|PV","SE|Electricity|Wind|Onshore","SE|Electricity|Wind|Offshore","SE|Electricity|Nuclear")) %>%
  group_by(model,scenario,region,period) %>%
  summarize(percentage = sum(percentage),.groups = "keep")

g[["Electricity Generation"]] <- 
  ggplot(plotData,aes(x=period,y=percentage)) +
  geom_area(aes(fill=variable),alpha=0.4) +
  geom_line(data=plotDataRenewables, linetype = "dashed", linewidth = 1.5, color="#666666") +
  geom_bar(data=plotData2040,aes(fill=variable),stat='identity', show.legend = FALSE, alpha = 0.7, color="white", width = 3, linewidth = 1.5) +
  #ggrepel::geom_text_repel(data=plotData2040[plotData2040$percentage >= 0.01,],aes(label = paste0(round(percentage,2)*100,"%"), y=cumPercentage-percentage/2, x = 2043.5), vjust=0.5, hjust=0.5, na.rm=TRUE, color = "black", size = 12) +
  #geom_text(data=plotData2040[plotData2040$percentage >= 0.01,],aes(label = paste0(inv_vars[variable], " (",round(percentage,2)*100,"%)"), y=cumPercentage-percentage/2, x = xPos), vjust=0.5, hjust=0, na.rm=TRUE, color = "black", size = 12) +
  geom_text(data=plotData2040[plotData2040$percentage >= 0.02,],aes(label = paste0(round(percentage,2)*100,"%", ifelse(inv_vars[variable] %in% c("Solar PV","Wind Onshore","Wind Offshore"),paste0(" - ", inv_vars[variable]), "")), y=cumPercentage-percentage/2, x = xPos), vjust=0.5, hjust=0, na.rm=TRUE, color = "black", size=geom.text.double.size) +
  theme_doubleColumn()  +
  scale_y_continuous(minor_breaks = seq(0.2, 0.8, 0.2), breaks=c(0,1),labels = scales::percent) +
  scale_x_continuous(breaks=c(2020,2040,2060)) +
  scale_fill_manual(values = setNames(color[names(vars)],vars), labels = setNames(names(vars), vars)) +
  theme(axis.title.y=element_blank()) +
  guides(fill=guide_legend(nrow=8,override.aes = list(alpha=1), byrow = TRUE)) +
  #annotate(geom = "text", x = 2022, y = plotDataRenewables[plotDataRenewables$period == 2020,]$percentage + 0.06, label = "renewables", color = "#888888",angle = 28, hjust=0, vjust=-0.5, size = 12) +
  #annotate(geom = "text", x = 2023, y = plotDataRenewables[plotDataRenewables$period == 2020,]$percentage - 0.06, label = "and nuclear", color = "#888888",angle = 28, hjust=0, vjust=-0.5, size = 12) +
  annotate(geom = "text", x = 2020, y = plotDataRenewables[plotDataRenewables$period == 2020,]$percentage, label = "renewables\nand nuclear", color = "#666666",angle = 26, hjust=0, vjust=1, size=geom.text.double.size) +
  theme(axis.text.y=element_blank()) +
  theme(legend.key.height = unit(0.8, "cm"),
    legend.key.width = unit(0.8,"cm")) +
  theme(legend.spacing.y = unit(0.1, 'cm')) +
  theme() +
  annotate(geom = "text", label = "100%", x = 2020, y = 1, hjust=1.2, color = "#666666", size=geom.text.double.size) +
  annotate(geom = "text", label = "0%", x = 2020, y = 0, hjust=1.2, color = "#666666", size=geom.text.double.size) +
  coord_cartesian(clip = 'off') +
  theme(plot.margin = margin(t = 0, r = 0, b = 0, l = 20, unit = "pt")
        #,legend.margin=margin(l = -2, unit='cm')
        )

ggsave(paste0(outFolder,"/Electricity Generation.svg"), g[["Electricity Generation"]] , device="svg", width = 6, height = 10, dpi=100, units = "in")
ggsave(paste0(outFolder,"/Electricity Generation.png"), g[["Electricity Generation"]] , device="png", width = 6, height = 10, dpi=300, units = "in")

```

```{r, echo=FALSE, fig.width=6, fig.height=10}

g[["Electricity Generation"]]

```

:::

::: {}

(b) Solar Photovoltaics capacity (GW)

```{r, echo=FALSE}

# Solar photovoltaics roll out
plotData <- df %>% 
  filter(region == "EU27", period %in% c(2020:2060), tgt2030 %in% tgt2030Scen, variable == "Cap|Electricity|Solar|PV") %>% 
  filter(tgt2050 != "Npi") %>% 
  group_by(model,region,variable,period) %>% 
  mutate(ribbon.max=max(value),
         ribbon.min=min(value))

target <- data.frame(target=c("REPowerEU","REPowerEU"),period=c(2025,2030),value=c(320,600))

#value <- data.frame(period=c(2025,2030,2040),value=c(plotData[plotData$period==2025,]$value,plotData[plotData$period==2030,]$value,plotData[plotData$period==2040,]$value))

# historical data
h <- hist %>% filter(model == "IRENA", region == "EU27", variable == "Cap|Electricity|Solar|PV", period >= 2005) %>% drop_na(value)

g[["Solar Photovoltaics capacity"]] <- ggplot(plotData,aes(x=period,y=value)) +
    geom_hline(yintercept=0, color = "black", linewidth=2, alpha = 0.5) +
    geom_vline(xintercept=2020, color = "black", linewidth=1.5, alpha = 0.5,linetype="dashed") +
    geom_segment(aes(x=2040, y=0, xend=2040, yend=plotData %>% filter(scenario == mainScen, period == 2040) %>% pull(value)), linetype="dashed", color="black") +
    geom_ribbon(aes(ymax = ribbon.max, ymin = ribbon.min), alpha = 0.5, fill = "#545454") + 
    geom_line(data=plotData %>% filter(scenario == mainScen),linewidth=1.5) +
    geom_line(data=h,linewidth=1,linetype="dashed") +
    geom_point(data=plotData %>% filter(scenario == mainScen, period %in% c(2030,2040)),size=3,shape=21,fill="white",stroke=1.5) +
    geom_point(data=target, size=3, shape=4, stroke = 1.5) +
    geom_label_repel(data=target, aes(label = paste0(round(value),"GW - ", target)),box.padding = 0.1,
                  #segment.curvature = -0.1,
                  segment.square = TRUE,
                  point.padding = 0.6,
                  nudge_x = 5,
                  #nudge_y = -0.02,
                  #force = 0.5,
                  hjust = 0,
                  direction="y", size = geom.text.double.size*0.9, xlim = c(2020, 2070)) +
  geom_label_repel(data=plotData %>% filter(scenario == mainScen, period == 2030), aes(label = paste0(round(value),"GW")),box.padding = 0.1,
                  point.padding = 1,
                  nudge_x = -2,
                  nudge_y = 500,
                  hjust = 1, direction="y",
                  size = geom.text.double.size) +
  geom_label_repel(data=plotData %>% filter(scenario == mainScen, period == 2040), aes(label = paste0(round(value),"GW")),box.padding = 0.1,
                  point.padding = 1,
                  nudge_x = -2,
                  nudge_y =500,
                  #hjust = 0, 
                  direction="y",
                  size = geom.text.double.size) +
    theme_doubleColumn() + 
    theme(axis.title.y = element_blank())+
    coord_cartesian(clip = 'off') +
    expand_limits(y = 3000)

ggsave(paste0(outFolder,"/Solar Photovoltaics capacity.svg"), g[["Solar Photovoltaics capacity"]] , device="svg", width = 6, height = 4, dpi=100, units = "in")
ggsave(paste0(outFolder,"/Solar Photovoltaics capacity.png"), g[["Solar Photovoltaics capacity"]] , device="png", width = 6, height = 4, dpi=300, units = "in")

```

```{r, echo=FALSE, fig.width=6, fig.height=4}

g[["Solar Photovoltaics capacity"]]

```

(c) Electricity share in final energy (%)

```{r, echo=FALSE}

# Electricity share
vars <- c("Total" = "Final Energy|Electricity Share",
          "Transport"="Final Energy|Transport|Electricity Share",
          "Industry"="Final Energy|Industry|Electricity Share",
          "Buildings"="Final Energy|Buildings|Electricity Share")

plotData <- df %>% 
  filter(region == "EU27", period %in% c(2020:2060), tgt2030 %in% tgt2030Scen, variable %in% unlist(vars)) %>% 
  group_by(model,region,variable,period) %>% 
  filter(tgt2050 != "Npi") %>% 
  mutate(ribbon.max=max(value),
         ribbon.min=min(value))

h_Orig <- hist[hist$model == "IEA" & hist$region == "EU27" & hist$variable %in% c("FE","FE|Electricity","FE|Transport","FE|Transport|Electricity","FE|Industry", "FE|Industry|Electricity","FE|Buildings","FE|Buildings|Electricity"),] %>% drop_na(value)

h <- rbind(
  h_Orig %>% calc_addVariable("`Final Energy|Electricity Share`" = "`FE|Electricity` / `FE`", units = "%", only.new=T),
  h_Orig %>% calc_addVariable("`Final Energy|Transport|Electricity Share`" = "`FE|Transport|Electricity` / `FE|Transport`", units = "%", only.new=T),
  h_Orig %>% calc_addVariable("`Final Energy|Industry|Electricity Share`"  = "`FE|Industry|Electricity`  / `FE|Industry`", units = "%", only.new=T),
  h_Orig %>% calc_addVariable("`Final Energy|Buildings|Electricity Share`" = "`FE|Buildings|Electricity` / `FE|Buildings`", units = "%", only.new=T))

g[["Electricity Share"]] <- ggplot(plotData,aes(x=period,y=value)) +
    geom_hline(yintercept=0, color = "black", linewidth=2, alpha = 0.5) +
    geom_vline(xintercept=2020, color = "black", linewidth=1.5, alpha = 0.5,linetype="dashed") +
    geom_segment(aes(x=2040, y=0, xend=2040, yend=max(plotData %>% filter(scenario == mainScen, period == 2040) %>% pull(value))), linetype="dashed", color="black") +
    geom_ribbon(aes(ymax = ribbon.max, ymin = ribbon.min, fill=variable), alpha = 0.3) + 
    geom_line(data=plotData %>% filter(scenario == mainScen), linewidth=1.5,aes(color=variable)) +
    #geom_line(data=h,linewidth=1,linetype="dashed") +
    geom_point(data=plotData %>% filter(scenario == mainScen, period == 2040),size=3,shape=21,fill="white",stroke=1.5) +
    geom_line(data=h %>% filter(period >= 2005),linetype="dashed", linewidth = 1, aes(color=variable)) +
    geom_label_repel(data= plotData %>% filter(scenario == mainScen, period == 2040),
                         aes(
                           label = paste0(round(value*100),"%")),
                      box.padding = 3,
                      #segment.curvature = -0.1,
                      segment.square = TRUE,
                      point.padding = 0.6,
                      nudge_x = 5,
                      nudge_y = c(0.55,0.27,-0.05,-0.6),
                      #force = 0.5,
                      #hjust = 0,
                      direction="y"
                      , size = geom.text.double.size*0.9
                      #, xlim = c(2020, 2070)
                      ) +
    scale_colour_manual(values = c("Final Energy|Electricity Share"="black","Final Energy|Transport|Electricity Share"="#a1dd70","Final Energy|Industry|Electricity Share"="#5f6769","Final Energy|Buildings|Electricity Share"="#5edfff"), labels = setNames(names(vars), vars)) +
    scale_fill_manual(values = c("Final Energy|Electricity Share"="black","Final Energy|Transport|Electricity Share"="#a1dd70","Final Energy|Industry|Electricity Share"="#5f6769","Final Energy|Buildings|Electricity Share"="#5edfff"), labels = setNames(names(vars), vars)) +
    theme_doubleColumn() + 
    scale_y_continuous(breaks = seq(0, 1, 0.2),labels = scales::percent) +
    expand_limits(y=c(0,1)) +
    guides(color=guide_legend(nrow=2)) +
  theme(axis.title.y = element_blank())

ggsave(paste0(outFolder,"/Electricity Share.svg"), g[["Electricity Share"]] , device="svg", width = 6, height = 6, dpi=100, units = "in")
ggsave(paste0(outFolder,"/Electricity Share.png"), g[["Electricity Share"]] , device="png", width = 6, height = 6, dpi=300, units = "in")

```

```{r, echo=FALSE, fig.width=6, fig.height=6}

g[["Electricity Share"]]

```

:::

::::

---

# Figure 3. Energy supply panel

:::: {style="display: flex;"}

::: {}

(a) Primary energy (EJ/yr)

```{r, echo=FALSE}

# Primary Energy
vars <- c("Biomass" = "PE|Biomass",
          "Coal"="PE|Coal",
          "Gas"="PE|Gas",
          "Oil"="PE|Oil")

plotData <- df %>% 
  filter(region == "EU27", period %in% c(2020:2060), tgt2030 %in% tgt2030Scen, variable %in% unlist(vars)) %>% 
  group_by(model,region,variable,period) %>% 
  filter(tgt2050 != "Npi") %>% 
  mutate(ribbon.max=max(value),
         ribbon.min=min(value))

# historical data
h <- hist %>% filter(model == "IEA", region == "EU27", variable %in% vars, period >= 2005, period <= 2020) %>% drop_na(value) %>% filter(value != 0)

g[["Primary Energy"]] <- ggplot(plotData,aes(x=period,y=value)) +
    geom_hline(yintercept=0, color = "black", linewidth=2, alpha = 0.5) +
    geom_vline(xintercept=2020, color = "black", linewidth=1.5, alpha = 0.5,linetype="dashed") +
    geom_segment(aes(x=2040, y=0, xend=2040, yend=max(plotData %>% filter(scenario == mainScen, period == 2040) %>% pull(value))), linetype="dashed", color="black") +
    geom_line(data=plotData %>% filter(scenario == mainScen), aes(color=variable), linewidth=1.5) +
    geom_line(data=h,aes(color=variable),linewidth=1,linetype="dashed") +
    geom_ribbon(aes(ymax = ribbon.max, ymin = ribbon.min, fill=variable), alpha = 0.3) + 
    #geom_point(data=plotData[plotData$period %in% c(2020,2030,2040),],size=4,shape=21,fill="white") +
    geom_point(data=plotData %>% filter(scenario == mainScen, period == 2040),aes(color=variable),size=3,shape=21,fill="white",stroke=1.5) +
    #annotate(geom = "text", x = 2020, y = plotDataRenewables[plotDataRenewables$period == 2020,]$percentage, label = "renewables\nand nuclear", color = "#666666",angle = 26, hjust=0, vjust=1, size=geom.text.double.size) +
    scale_color_manual(values = c(setNames(options$aesthetics$legend$color[names(vars)],vars)), labels = setNames(names(vars), vars)) +
    scale_fill_manual(values = c(setNames(options$aesthetics$legend$color[names(vars)],vars)), labels = setNames(names(vars), vars)) +
    theme_doubleColumn() + 
    expand_limits(y=c(0,1)) +
    guides(color=guide_legend(nrow=2)) +
    theme(axis.title.y = element_blank()) +
    expand_limits(y = 30)
    #guides(fill = FALSE)

ggsave(paste0(outFolder,"/Primary Energy.svg"), g[["Primary Energy"]] , device="svg", width = 6, height = 6, dpi=100, units = "in")
ggsave(paste0(outFolder,"/Primary Energy.png"), g[["Primary Energy"]] , device="png", width = 6, height = 6, dpi=300, units = "in")

```

```{r, echo=FALSE, fig.width=6, fig.height=6}

g[["Primary Energy"]]

```

(c) Natural gas and syngas final energy demand (EJ/yr)

```{r, echo=FALSE}

vars <- c(
  "Hydrogen"="FE|Hydrogen",
  "Gases"="FE|Gases"
  #"Syngas" = "FE|Gases|Hydrogen",
  #"Biogas" = "FE|Gases|Biomass",
  #"Natural gas" = "FE|Gases|Fossil"
)

plotData <- df %>% 
  filter(region == "EU27", period %in% c(2020,2030,2040,2050), tgt2030 %in% tgt2030Scen, variable %in% unlist(vars), tgt2050 != "Npi") %>% 
  mutate(variable = factor(variable, levels=vars))  %>% 
  group_by(model,scenario,region,period) %>% 
  mutate(cumsum = cumsum(value),
         sum = sum(value)) %>% 
  group_by(model,region,variable,period) %>% 
  mutate(ribbon.max=max(cumsum),
         ribbon.min=min(cumsum)) %>% 
  group_by(model,region,period) %>% 
  mutate(ribbon.max.total=max(sum),
         ribbon.min.total=min(sum))

g[["FE Gases"]] <- 
  ggplot(plotData,aes(x=period,y=value)) +
  geom_col(data=plotData %>% filter(scenario == mainScen), aes(fill = variable, alpha=factor(period))) +
  #geom_errorbar(data=plotData %>% filter(period != 2020), aes(ymin=ribbon.min, ymax=ribbon.max, group=variable), width=8, colour="black", alpha=0.9, size=1,position = position_dodge(width=2)) +
  geom_errorbar(data=plotData %>% filter(period != 2020, variable == vars[1], scenario == mainScen), aes(ymin=ribbon.min.total, ymax=ribbon.max.total), width=4, colour="black", alpha=0.9, linewidth=1,position = position_dodge(width=2)) +
  # scale_y_continuous(labels = scales::percent, breaks = c(0,1,plotData[plotData$period == "2030",]$value,plotData[plotData$period == "2040",]$value,plotData[plotData$period == "2050",]$value)) +
  coord_flip() +
  theme_doubleColumn() +
  theme(panel.grid.minor.x = element_blank(),
        panel.grid.major.y = element_blank()) +
  #guides(y = "none") +
  theme(axis.title.y=element_blank()) +
  #theme(legend.position = "none") #+
  scale_fill_manual(values = setNames(options$aesthetics$legend$color[names(vars)],vars), labels = setNames(names(vars), vars), guide = guide_legend(reverse = TRUE, nrow = 1)) +
  scale_alpha_manual(values = c("2020"=0.7,"2030"=0.7,"2040"=1,"2050"=0.7), guide = 'none') +
  theme(legend.text=element_text(size=theme.double.text.size)) +
  theme(legend.key.height = unit(0.7, "cm"),
    legend.key.width = unit(0.7,"cm")) +
  scale_y_continuous(breaks = c(0,round(sum(plotData %>% filter(scenario == mainScen, period == 2020) %>% pull(value)),1),round(sum(plotData %>% filter(scenario == mainScen, period == 2030) %>% pull(value)),1),round(sum(plotData %>% filter(scenario == mainScen, period == 2040) %>% pull(value)),1))) +
  #scale_x_reverse()
  theme(legend.margin=margin(l = -2, unit='cm'))

ggsave(paste0(outFolder,"/FE Gases.svg"), g[["FE Gases"]] , device="svg", width = 6, height = 4, dpi=100, units = "in")
ggsave(paste0(outFolder,"/FE Gases.png"), g[["FE Gases"]] , device="png", width = 6, height = 4, dpi=300, units = "in")

# vars <- c(
#   "Hydrogen"="FE|Hydrogen",
#   #"Gases"="FE|Gases"
#   "Syngas" = "FE|Gases|Hydrogen",
#   "Biogas" = "FE|Gases|Biomass",
#   "Natural gas" = "FE|Gases|Fossil"
# )
# 
# plotData <- df %>% 
#   filter(region == "EU27", period %in% c(2020,2030,2040,2050), tgt2030 %in% tgt2030Scen, variable %in% unlist(vars), tgt2050 != "Npi") %>% 
#   mutate(variable = factor(variable, levels=vars))  %>% 
#   group_by(model,scenario,region,period,bioLim) %>% 
#   mutate(cumsum = cumsum(value),
#          sum = sum(value)) %>% 
#   group_by(model,region,variable,period,bioLim) %>% 
#   mutate(ribbon.max=max(cumsum),
#          ribbon.min=min(cumsum)) %>% 
#   group_by(model,region,period) %>% 
#   mutate(ribbon.max.total=max(sum),
#          ribbon.min.total=min(sum))
# 
#   
# g[["FE Gases"]] <- 
#   ggplot(plotData,aes(x=period,y=value)) +
#   geom_col(data=plotData %>% filter(scenario %in% c(mainScen,"Nzero_55_bio12")), aes(fill = variable, alpha=factor(period))) +
#   #geom_errorbar(data=plotData %>% filter(period != 2020), aes(ymin=ribbon.min, ymax=ribbon.max, group=variable), width=8, colour="black", alpha=0.9, size=1,position = position_dodge(width=2)) +
#   geom_errorbar(data=plotData %>% filter(period != 2020, variable == vars[1], scenario %in%  c(mainScen,"Nzero_55_bio12")), aes(ymin=ribbon.min.total, ymax=ribbon.max.total), width=4, colour="black", alpha=0.9, linewidth=1,position = position_dodge(width=2)) +
#   # scale_y_continuous(labels = scales::percent, breaks = c(0,1,plotData[plotData$period == "2030",]$value,plotData[plotData$period == "2040",]$value,plotData[plotData$period == "2050",]$value)) +
#   coord_flip() +
#   facet_grid(bioLim~.) +
#   theme_doubleColumn() +
#   theme(panel.grid.minor.x = element_blank(),
#         panel.grid.major.y = element_blank()) +
#   #guides(y = "none") +
#   theme(axis.title.y=element_blank()) +
#   #theme(legend.position = "none") #+
#   scale_fill_manual(values = setNames(options$aesthetics$legend$color[names(vars)],vars), labels = setNames(names(vars), vars), guide = guide_legend(reverse = TRUE, nrow = 1)) +
#   scale_alpha_manual(values = c("2020"=0.7,"2030"=0.7,"2040"=1,"2050"=0.7), guide = 'none') +
#   theme(legend.text=element_text(size=theme.double.text.size)) +
#   theme(legend.key.height = unit(0.7, "cm"),
#     legend.key.width = unit(0.7,"cm")) +
#   scale_y_continuous(breaks = c(0,round(sum(plotData %>% filter(scenario == mainScen, period == 2020) %>% pull(value)),1),round(sum(plotData %>% filter(scenario == mainScen, period == 2030) %>% pull(value)),1),round(sum(plotData %>% filter(scenario == mainScen, period == 2040) %>% pull(value)),1))) +
#   #scale_x_reverse()
#   theme(legend.margin=margin(l = -2, unit='cm'))

```

```{r, echo=FALSE, fig.width=6, fig.height=4}

g[["FE Gases"]]

```

:::

::: {}

(b) Fossil trade (EJ/yr)

```{r, echo=FALSE}

vars <- list(
 "Coal" = c("Trade|Imports|Coal","Trade|Exports|Coal"),
 "Oil" = c("Trade|Imports|Oil","Trade|Exports|Oil"),
 "Gas" = c("Trade|Imports|Gas","Trade|Exports|Gas"),
 "Biomass" = c("Trade|Imports|Biomass","Trade|Exports|Biomass")
)

# plotData <- plotData %>% 
#   arrange(factor(variable, levels = rev(vars))) %>%
#   group_by(model,scenario,region,period) %>%
#   mutate(cumValue = cumsum(value),
#          n = sum(value),
#          percentage = value / n,
#          cumPercentage = cumValue / n)

plotData <- df %>% 
  filter(region == "EU27", period  %in% c(2020,2030,2040,2050), scenario == mainScen, variable %in% unlist(vars))

carrier <- c()
for(c in names(vars)){
  for(var in vars[[c]]){
    v <- c
    names(v) <- var
    carrier <- c(carrier,v)
  }
}

plotData <- plotData %>%
  mutate(type = ifelse(grepl("Imports",variable),"Imports","Exports"),
         carrier = carrier[as.character(variable)])

plotData[plotData$type == "Exports",]$value <- -plotData[plotData$type == "Exports",]$value

plotData$carrier <- factor(plotData$carrier, levels=c("Coal","Oil","Gas","Biomass"))

plotDataTotal <- plotData %>%
  group_by(model,scenario,region,unit,period,carrier) %>%
  summarize(value=sum(value),.groups = "keep") %>%
  mutate(type="EU-27 Net-trade")

plotColor <- c(
  "Imports"= "#e56b6f",
  "Exports"= "#355070"
)

# g[["security of supply"]] <- ggplot(plotData,aes(x=period,y=value)) +
#   geom_bar(aes(fill=type),stat='identity') +
#   geom_errorbar(data  = plotDataTotal, aes(y=value, ymax=value, ymin=value, color=type), size=1) +
#   facet_wrap(~carrier,nrow=1) +
#    theme_doubleColumn() +
#    theme(axis.title.y=element_blank()) +
#    scale_fill_manual(values = plotColor, guide = guide_legend(nrow = 1)) + #, guide = guide_legend(nrow = 2) setNames(options$aesthetics$legend$color[names(vars)],vars)
#   scale_color_manual(values = c("Total"="black")) +
#    scale_x_continuous(breaks = c(2020,2030,2040,2050)) + 
#   #theme(
#   #  strip.background = element_blank(),
#   #  strip.text.x = element_blank()) +
#   theme(legend.text=element_text(size=theme.single.text.size)) +
#   theme(legend.key.height = unit(0.7, "cm"),
#     legend.key.width = unit(0.7,"cm")) + 
#   theme(axis.text = element_text(size=theme.single.text.size)) #+                                                              
#   #theme(strip.text.x = element_text(size=0.8*theme.single.text.size)) +
#   #theme(strip.clip = "off")
# 
# 
#  ggsave(paste0(outFolder,"/security of supply.svg"), g[["security of supply"]] , device="svg", width = 6, height = 5, dpi=100, units = "in")
#  ggsave(paste0(outFolder,"/security of supply.png"), g[["security of supply"]] , device="png", width = 6, height = 5, dpi=300, units = "in")

g[["security of supply - fossil"]] <- ggplot(plotData %>% filter(period <= 2050, carrier != "Biomass"),aes(x=period,y=value)) +
  geom_bar(aes(fill=type),stat='identity', alpha = 0.7) +
  geom_errorbar(data  = plotDataTotal %>% filter(period <= 2050, carrier != "Biomass"), aes(y=value, ymax=value, ymin=value, color=type), linewidth=1.5) +
  facet_wrap(~carrier,nrow=1) +
  theme_doubleColumn()  +
  theme(axis.title.y=element_blank()) +
  scale_fill_manual(values = plotColor, guide = guide_legend(ncol = 1, order = 2)) + #, guide = guide_legend(nrow = 2) setNames(options$aesthetics$legend$color[names(vars)],vars)
  scale_color_manual(values = c("EU-27 Net-trade"="black"), guide = guide_legend(order = 1)) +
  scale_x_continuous(breaks = c(2020,2030,2040,2050)) + 
  #theme(
  #  strip.background = element_blank(),
  #  strip.text.x = element_blank()) +
  theme(legend.text=element_text(size=theme.single.text.size)) +
  theme(legend.key.height = unit(0.7, "cm"),
    legend.key.width = unit(1.5,"cm")) #+ 
  #theme(axis.text = element_text(size=theme.single.text.size)) #+                                                              
  #theme(strip.text.x = element_text(size=0.8*theme.single.text.size)) +
  #theme(strip.clip = "off")

ggsave(paste0(outFolder,"/security of supply - fossil.svg"), g[["security of supply - fossil"]] , device="svg", width = 6, height = 6, dpi=100, units = "in")
ggsave(paste0(outFolder,"/security of supply - fossil.png"), g[["security of supply - fossil"]] , device="png", width = 6, height = 6, dpi=300, units = "in")

```

```{r, echo=FALSE, fig.width=6, fig.height=5}

  g[["security of supply - fossil"]]

```

(d) Hydrogen production (EJ/yr)

```{r, echo=FALSE}

vars <- c(
  "Biomass w/ CC" = "SE|Hydrogen|Biomass|w/ CC",
  "Biomass w/o CC" = "SE|Hydrogen|Biomass|w/o CC",
  "Coal w/ CC" = "SE|Hydrogen|Coal|w/ CC",
  "Coal w/o CC" = "SE|Hydrogen|Coal|w/o CC",
  "Gas w/ CC" = "SE|Hydrogen|Gas|w/ CC",
  "Gas w/o CC" = "SE|Hydrogen|Gas|w/o CC",
  "Electrolysis" = "SE|Hydrogen|Electricity",
  #"VRE Storage Electrolysis" = "SE|Hydrogen|Electricity|VRE Storage Electrolysis",
  #"Standard Electrolysis" = "SE|Hydrogen|Electricity|Standard Electrolysis",
  "Net Imports" = "SE|Hydrogen|Net Imports")

#1EJ <- 7 million tons #https://hydrogencouncil.com/wp-content/uploads/2017/11/Hydrogen-Scaling-up_Hydrogen-Council_2017.compressed.pdf
#1EJ <- 8 million tons adrian

color <- setNames(options$aesthetics$legend$color[names(vars)],vars)
color["SE|Hydrogen|Biomass|w/ CC"] <- color["SE|Hydrogen|Biomass|w/o CC"]

target <- data.frame(target=c("REPowerEU","REPowerEU"),period=c(2030,2030),value=c(10,20)) %>%
  mutate(value = value/8)

plotData <- df %>% 
  filter(region == "EU27", period %in% c(2020,2030,2040,2050), tgt2030 %in% tgt2030Scen, variable %in% unlist(vars)) %>% 
  mutate(#value = value*8, #1EJ <- 8 million tons adrian
    variable = factor(variable, levels=vars))  %>% 
  filter(tgt2050 != "Npi")  %>%
  group_by(model,scenario,region,period) %>% 
  mutate(cumsum = cumsum(value),
         sum = sum(value)) %>% 
  group_by(model,region,variable,period) %>% 
  mutate(ribbon.max=max(cumsum),
         ribbon.min=min(cumsum)) %>% 
  group_by(model,region,period) %>% 
  mutate(ribbon.max.total=max(sum),
         ribbon.min.total=min(sum))

g[["Hydrogen"]] <- 
  ggplot(plotData,aes(x=period,y=value)) +
  geom_hline(yintercept= target[1,]$value, color = "black", linewidth=1, alpha = 0.5, linetype = "dashed") +
  annotate("text", x = 2015, y = target[1,]$value, label = "REPowerEU\ndomestic", angle=90, vjust = 0.4, hjust = 0, color = "black", alpha = 0.5, size = 0.3*theme.single.text.size) +
  geom_hline(yintercept= target[2,]$value, color = "black", linewidth=1, alpha = 0.5, linetype = "dashed") +
  annotate("text", x = 2015, y = target[2,]$value, label = "REPowerEU\ntotal", angle=90, vjust = 0.4, hjust = 0, color = "black", alpha = 0.5, size = 0.3*theme.single.text.size) +
  geom_col(data=plotData %>% filter(scenario == mainScen), aes(fill = variable, alpha=factor(period))) +
  geom_errorbar(data=plotData %>% filter(period != 2020, variable == vars[1], scenario == mainScen), aes(ymin=ribbon.min.total, ymax=ribbon.max.total), width=4, colour="black", alpha=0.9, linewidth=1,position = position_dodge(width=2)) +
  # scale_y_continuous(labels = scales::percent, breaks = c(0,1,plotData[plotData$period == "2030",]$value,plotData[plotData$period == "2040",]$value,plotData[plotData$period == "2050",]$value)) +
  coord_flip() +
  theme_doubleColumn() +
  theme(panel.grid.minor.x = element_blank(),
        panel.grid.major.y = element_blank()) +
  #guides(y = "none") +
  theme(axis.title.y=element_blank()) +
  #theme(legend.position = "none") #+
  scale_fill_manual(values = color,
                    labels = setNames(names(vars), vars)[c("SE|Hydrogen|Electricity", "SE|Hydrogen|Gas|w/o CC", "SE|Hydrogen|Biomass|w/ CC")],
                    breaks = c("SE|Hydrogen|Electricity", "SE|Hydrogen|Gas|w/o CC", "SE|Hydrogen|Biomass|w/ CC"), guide = guide_legend(nrow = 1)) +
  scale_alpha_manual(values = c("2020"=0.7,"2030"=0.7,"2040"=1,"2050"=0.7), guide = 'none') +
  theme(legend.text=element_text(size=theme.double.text.size)) +
  theme(legend.key.height = unit(0.7, "cm"),
    legend.key.width = unit(0.7,"cm")) +
  scale_y_continuous(breaks = c(0,plotData %>% filter(scenario == mainScen, period == 2030, variable == vars[1]) %>% pull(sum),plotData %>% filter(scenario == mainScen, period == 2040, variable == vars[1]) %>% pull(sum),plotData %>% filter(scenario == mainScen, period == 2050, variable == vars[1]) %>% pull(sum)),
                     labels = c(format(0,nsmall = 1),format(round(plotData %>% filter(scenario == mainScen, period == 2030, variable == vars[1]) %>% pull(sum),1),nsmall = 1),format(round(plotData %>% filter(scenario == mainScen, period == 2040, variable == vars[1]) %>% pull(sum),1),nsmall = 1),format(round(plotData %>% filter(scenario == mainScen, period == 2050, variable == vars[1]) %>% pull(sum),1),nsmall = 1))) +
  #scale_y_continuous(breaks = c(0,round(sum(plotData %>% filter(scenario == mainScen, period == 2030) %>% pull(value))),round(sum(plotData %>% filter(scenario == mainScen, period == 2040) %>% pull(value))),round(sum(plotData %>% filter(scenario == mainScen, period == 2050) %>% pull(value))))) +
  #scale_x_reverse() +
  theme(legend.margin=margin(l = -2, unit='cm'))

ggsave(paste0(outFolder,"/Hydrogen.svg"), g[["Hydrogen"]] , device="svg", width = 6, height = 4, dpi=100, units = "in")
ggsave(paste0(outFolder,"/Hydrogen.png"), g[["Hydrogen"]] , device="png", width = 6, height = 4, dpi=300, units = "in")

```

```{r, echo=FALSE, fig.width=6, fig.height=4}

g[["Hydrogen"]]

```

:::

::::

---

# Figure 4. Demand panel

(a) Transport final energy (EJ/yr)

```{r, echo=FALSE}

FEvars <- list(
  "Domestic" = list(
    "LDV" = c(
      "FE|Transport|Pass|Road|LDV|Electricity" = "Electricity",
      "FE|Transport|Pass|Road|LDV|Gases" = "Gases",
      "FE|Transport|Pass|Road|LDV|Hydrogen" = "Hydrogen",
#      "FE|Transport|Pass|Road|LDV|Liquids" = "Liquids"
       "FE|Transport|Pass|Road|LDV|Liquids|Fossil" = "Liquids - Fossil",
       "FE|Transport|Pass|Road|LDV|Liquids|Biomass" = "Liquids - Biomass",
        "FE|Transport|Pass|Road|LDV|Liquids|Hydrogen" = "Liquids - Synthetic"
    ),
    "Buses" = c(
      "FE|Transport|Pass|Road|Bus|Electricity" = "Electricity",
      "FE|Transport|Pass|Road|Bus|Gases" = "Gases",
      "FE|Transport|Pass|Road|Bus|Hydrogen" = "Hydrogen",
#      "FE|Transport|Pass|Road|Bus|Liquids" = "Liquids"
        "FE|Transport|Pass|Road|Bus|Liquids|Fossil" = "Liquids - Fossil",
        "FE|Transport|Pass|Road|Bus|Liquids|Biomass" = "Liquids - Biomass",
        "FE|Transport|Pass|Road|Bus|Liquids|Hydrogen" = "Liquids - Synthetic"

    ),
    "Trucks" = c(
      "FE|Transport|Freight|Road|Electricity" = "Electricity",
      "FE|Transport|Freight|Road|Gases" = "Gases",
      "FE|Transport|Freight|Road|Hydrogen" = "Hydrogen",
#      "FE|Transport|Freight|Road|Liquids" = "Liquids"
        "FE|Transport|Freight|Road|Liquids|Fossil" = "Liquids - Fossil",
        "FE|Transport|Freight|Road|Liquids|Biomass" = "Liquids - Biomass",
        "FE|Transport|Freight|Road|Liquids|Hydrogen" = "Liquids - Synthetic"
    ),
    "Rail" = c(
      "FE|Transport|Pass|Rail|Electricity" = "Electricity",
      "FE|Transport|Pass|Rail|Gases" = "Gases",
      "FE|Transport|Pass|Rail|Hydrogen" = "Hydrogen",
#      "FE|Transport|Pass|Rail|Liquids" = "Liquids",
        "FE|Transport|Pass|Rail|Liquids|Fossil" = "Liquids - Fossil",
        "FE|Transport|Pass|Rail|Liquids|Biomass" = "Liquids - Biomass",
        "FE|Transport|Pass|Rail|Liquids|Hydrogen" = "Liquids - Synthetic",
      "FE|Transport|Freight|Rail|Electricity" = "Electricity",
      "FE|Transport|Freight|Rail|Gases" = "Gases",
      "FE|Transport|Freight|Rail|Hydrogen" = "Hydrogen",
#      "FE|Transport|Freight|Rail|Liquids" = "Liquids"
        "FE|Transport|Freight|Rail|Liquids|Fossil" = "Liquids - Fossil",
        "FE|Transport|Freight|Rail|Liquids|Biomass" = "Liquids - Biomass",
        "FE|Transport|Freight|Rail|Liquids|Hydrogen" = "Liquids - Synthetic"
    ),
    "Aviation" = c(
      "FE|Transport|Pass|Aviation|Domestic|Electricity" = "Electricity",
      "FE|Transport|Pass|Aviation|Domestic|Gases" = "Gases",
      "FE|Transport|Pass|Aviation|Domestic|Hydrogen" = "Hydrogen",
#      "FE|Transport|Pass|Aviation|Domestic|Liquids" = "Liquids"
        "FE|Transport|Pass|Aviation|Domestic|Liquids|Fossil" = "Liquids - Fossil",
        "FE|Transport|Pass|Aviation|Domestic|Liquids|Biomass" = "Liquids - Biomass",
        "FE|Transport|Pass|Aviation|Domestic|Liquids|Hydrogen" = "Liquids - Synthetic"
    )
  ),
  "International" = list(
    "Aviation" = c(
      "FE|Transport|Pass|Aviation|International|Electricity" = "Electricity",
      "FE|Transport|Pass|Aviation|International|Gases" = "Gases",
      "FE|Transport|Pass|Aviation|International|Hydrogen" = "Hydrogen",
#      "FE|Transport|Pass|Aviation|International|Liquids" = "Liquids"
        "FE|Transport|Pass|Aviation|International|Liquids|Fossil" = "Liquids - Fossil",
        "FE|Transport|Pass|Aviation|International|Liquids|Biomass" = "Liquids - Biomass",
        "FE|Transport|Pass|Aviation|International|Liquids|Hydrogen" = "Liquids - Synthetic"
    ),
    "Shipping" = c(
      "FE|Transport|Freight|Navigation|Electricity" = "Electricity",
      "FE|Transport|Freight|Navigation|Gases" = "Gases",
      "FE|Transport|Freight|Navigation|Hydrogen" = "Hydrogen",
#      "FE|Transport|Freight|Navigation|Liquids" = "Liquids",
        "FE|Transport|Freight|Navigation|Liquids|Fossil" = "Liquids - Fossil",
        "FE|Transport|Freight|Navigation|Liquids|Biomass" = "Liquids - Biomass",
        "FE|Transport|Freight|Navigation|Liquids|Hydrogen" = "Liquids - Synthetic",
      "FE|Transport|Freight|International Shipping|Electricity" = "Electricity",
      "FE|Transport|Freight|International Shipping|Gases" = "Gases",
      "FE|Transport|Freight|International Shipping|Hydrogen" = "Hydrogen",
#      "FE|Transport|Freight|International Shipping|Liquids" = "Liquids"
        "FE|Transport|Freight|International Shipping|Liquids|Fossil" = "Liquids - Fossil",
        "FE|Transport|Freight|International Shipping|Liquids|Biomass" = "Liquids - Biomass",
        "FE|Transport|Freight|International Shipping|Liquids|Hydrogen" = "Liquids - Synthetic"
    )
  )
)

FEvarList <- unlist(c(FEvars$Domestic$LDV,FEvars$Domestic$Buses,FEvars$Domestic$Trucks,FEvars$Domestic$Rail,FEvars$Domestic$Aviation,FEvars$International$Aviation,FEvars$International$Shipping))

typeVar <- c()
for(national in names(FEvars)){
  for(mode in names(FEvars[[national]])){
    for(var in names(FEvars[[national]][[mode]])){
      v <- mode
      names(v) <- var
      typeVar <- c(typeVar,v)
    }
  }
}

FEplotData <- df %>% 
  filter(region == "EU27", period >= 2020, period <= 2060, scenario == mainScen, variable %in% names(FEvarList))

FEplotData <- FEplotData %>%
  mutate(national = ifelse(variable %in% names(c(FEvars$Domestic$LDV,FEvars$Domestic$Buses,FEvars$Domestic$Trucks,FEvars$Domestic$Rail,FEvars$Domestic$Aviation)), "Domestic", "International"),
         name = FEvarList[as.character(variable)],
         type="Final Energy",
         mode = typeVar[as.character(variable)]) %>%
  group_by(model,scenario,region,unit,period,national,name,type,mode) %>%
  summarize(value=sum(value),.groups = "keep")

FEplotData$mode <- factor(FEplotData$mode, levels=c("LDV","Buses","Trucks","Rail","Shipping","Aviation"))

FEplotData$name <- factor(FEplotData$name, levels=c("Electricity","Hydrogen","Liquids","Liquids - Biomass","Liquids - Synthetic","Liquids - Fossil","Gases"))

plotColor <- c(
  "Electricity" = "#F3E16B",
  "Hydrogen" = "#82c9d9",
  "Liquids" = "#282828",
  "Liquids - Fossil" = "#282828",
  "Liquids - Biomass" = "#005900",
  "Liquids - Synthetic" = "#b56576",
  "Gases" = "#999959"
)

g[["transport fe"]] <- ggplot(FEplotData,aes(x=period,y=value)) +
   geom_bar(aes(fill=name, alpha=factor(period)),stat='identity') +
   facet_wrap(type~national+mode,nrow=1) +
   theme_singleColumn() +
   theme(axis.title.y=element_blank()) +
   scale_fill_manual(values = plotColor, guide = guide_legend(nrow = 2)) + #, guide = guide_legend(nrow = 2)
   scale_alpha_manual(values = c("2020"=0.7,"2025"=0.7,"2030"=0.7,"2035"=0.7,"2040"=1,"2045"=0.7,"2050"=0.7,"2055"=0.7,"2060"=0.7), guide = 'none') +
   scale_x_continuous(breaks = c(2020,2040,2060)) + 
  theme(
    strip.background = element_blank(),
    strip.text.x = element_blank()) +
  theme(legend.text=element_text(size=theme.single.text.size)) +
  theme(legend.key.height = unit(0.7, "cm"),
    legend.key.width = unit(0.7,"cm"))
#+ scale_y_continuous(name = "Final Energy")

ggsave(paste0(outFolder,"/transport fe.svg"), g[["transport fe"]] , device="svg", width = 12, height = 5, dpi=100, units = "in")
ggsave(paste0(outFolder,"/transport fe.png"), g[["transport fe"]] , device="png", width = 12, height = 5, dpi=300, units = "in")

```

```{r, echo=FALSE, fig.width=12, fig.height=5}

g[["transport fe"]]

```

:::: {style="display: flex;"}

::: {}

(b) LDV sales and ICE stock (Millions of vehicles)

```{r, echo=FALSE}

#LDV

Sales <- list(
  "BEV" = "Sales|Transport|LDV|BEV",
  "FCEV" = "Sales|Transport|LDV|FCEV",
  "Hybrid Electric" =	"Sales|Transport|LDV|Hybrid Electric",
  "Liquids" = "Sales|Transport|LDV|Liquids",
  "Gases" = "Sales|Transport|LDV|NG"
)

Stock <- list(
  "BEV" = "Stock|Transport|LDV|BEV",
  "FCEV" = "Stock|Transport|LDV|FCEV",
  "Hybrid Electric" =	"Stock|Transport|LDV|Hybrid Electric",
  "Liquids" = "Stock|Transport|LDV|Liquids",
  "Gases" = "Stock|Transport|LDV|NG"
)

plotColor <- c(
  "BEV" = "#F3E16B",
  "FCEV" = "#82c9d9",
  "Hybrid Electric" =	"#F37A5E",
  "Liquids" = "#282828",
  "Gases" = "#F7EFE8" 
)

vars <- Sales

plotData <- df %>% 
  filter(region == "EU27", period >= 2020, period <= 2060, scenario == mainScen, variable %in% unlist(vars)) %>% 
  mutate(variable = factor(variable, levels=vars))

labels <- setNames(names(vars),vars[names(vars)])

plotData$labels <- labels[as.character(plotData$variable)]

#stock
stockData <- df %>% 
  filter(region == "EU27", period >= 2020, period <= 2060, scenario == mainScen, variable %in% unlist(Stock)) %>% 
  mutate(variable = factor(variable, levels=Stock)) %>%
  filter(variable %in% c("Stock|Transport|LDV|Liquids","Stock|Transport|LDV|NG")) %>%
  group_by(period) %>%
  summarize(value = sum(value),.groups = "keep") %>%
  mutate(variable = "ICE Stock")
stockData$value <- stockData$value/10

g[["LDV Sales"]] <-
  ggplot(plotData,aes(x=period,y=value)) +
  geom_bar(aes(fill=variable, alpha=factor(period)),stat='identity') +
  geom_line(data=stockData, aes(color=variable),linetype=11, lwd = 1) +
# geom_line(data=h,linetype="dashed", linewidth = 1, color="black") +
 geom_hline(yintercept=0, color = "black", linewidth=2, alpha = 1) +
 geom_label_repel(data=data.frame(period=2030,value=stockData %>% filter(period == 2030) %>% pull(value)), label = "Stock",
                  #box.padding = 0.1,
                  #segment.curvature = -0.1,
                  #segment.square = TRUE,
                  #point.padding = 0.6,
                  nudge_x = 5,
                  nudge_y = 5,
                  #nudge_y = -0.02,
                  #force = 0.5,
                  hjust = 0,
                  direction="x", size = geom.text.double.size*0.9) +
  theme_doubleColumn() +
 #ylab(expression(paste("Energy and Industrial Processes Emissions - Mt" , CO[2], "eq/yr"))) +
 scale_fill_manual(values = setNames(plotColor[names(vars)],vars), labels = setNames(names(vars), vars), guide = guide_legend(nrow = 2, order = 1)) + #, guide = guide_legend(nrow = 2)
 scale_color_manual(values = c("ICE Stock"="black"), guide = guide_legend(order = 2, keywidth = unit(2, 'cm'))) +
 scale_alpha_manual(values = c("2020"=0.7,"2025"=0.7,"2030"=0.7,"2035"=0.7,"2040"=1,"2045"=0.7,"2050"=0.7,"2055"=0.7,"2060"=0.7), guide = 'none') +
 scale_x_continuous(breaks=c(2020,2030,2040,2050,2060)) +
 # theme(legend.margin=margin(l = -2, unit='cm')) + 
 #theme(legend.position="right", legend.justification = "bottom") +
 #theme(axis.title.y = element_blank()) +
 scale_y_continuous(name = "Sales", # - Million vehicles", 
                     #breaks = c(round(emi1990), round(emi1990*0.6), round(emi1990*0.45), round(emi1990*0.3), round(emi1990*0.15), 0),
                     sec.axis = sec_axis( trans=~.*10, name="Stock")) +
                                          #, labels = scales::percent, breaks = c(1,0.4,0.55,0.7,0.85, 0))) +
  theme(legend.text=element_text(size=theme.double.text.size),
        legend.spacing.x = unit(0.2, 'cm'))  +
  guides(color = "none") +
  theme(axis.title.y = element_text(colour = "#666666"))

#+
 # theme(legend.key.height = unit(0.7, "cm"),
 #   legend.key.width = unit(0.4,"cm"))

ggsave(paste0(outFolder,"/LDV Sales.svg"), g[["LDV Sales"]] , device="svg", width = 6, height = 5, dpi=100, units = "in")
ggsave(paste0(outFolder,"/LDV Sales.png"), g[["LDV Sales"]] , device="png", width = 6, height = 5, dpi=300, units = "in")

```

```{r, echo=FALSE, fig.width=6, fig.height=5}

   g[["LDV Sales"]]

```

(e) Industry CO2 emissions per sub-sector (Mt CO2/yr)

```{r, echo=FALSE}

# Industry Emissions
vars <- c("Cement" = "Emi|CO2|Energy|Demand|Industry|Cement",
          "Steel"="Emi|CO2|Energy|Demand|Industry|Steel",
          "Chemicals"="Emi|CO2|Energy|Demand|Industry|Chemicals",
          "Other Industry"="Emi|CO2|Energy|Demand|Industry|Other Industry")

plotColor <- c(
  "Cement" = "#355070",
  "Steel" = "#219ebc",
  "Chemicals" = "#eaac8b",
  "Other Industry" = "#e56b6f"
)

plotData <- df %>% 
  filter(region == "EU27", period %in% c(2020:2060), tgt2030 %in% tgt2030Scen, variable %in% unlist(vars)) %>% 
  mutate(variable = factor(variable, levels=vars)) %>%
  group_by(model,region,variable,period) %>% 
  filter(tgt2050 != "Npi") %>% 
  mutate(ribbon.max=max(value),
         ribbon.min=min(value))

# historical data #missing non-energy
#h <- hist[hist$model == "UNFCCC" & hist$region == "EU27" & hist$variable %in% vars,] %>% drop_na(value)

#View(hist[hist$model == "UNFCCC" & hist$region == "EU27" & hist$variable == "Emi|CO2|Industrial Processes",] %>% drop_na(value))
# Emi|CO2|Energy|Demand|Industry
# Emi|CO2|Energy|Demand|Industry|Cement
# Emi|CO2|Energy|Demand|Industry|Chemicals
# Emi|CO2|Energy|Demand|Industry|Other Industry
# Emi|CO2|Energy|Demand|Industry|Other Industry|AFOFI
# Emi|CO2|Energy|Demand|Industry|Steel
# Emi|CO2|Industrial Processes
# Emi|CO2|Industrial Processes|Chemicals
# Emi|CO2|Industrial Processes|Metals
# Emi|CO2|Industrial Processes|Minerals
# Emi|GHG|Energy|Demand|Industry
# Emi|GHG|Industrial Processes

g[["Industry Emissions"]] <- ggplot(plotData,aes(x=period,y=value)) +
    geom_hline(yintercept=0, color = "black", linewidth=2, alpha = 0.5) +
    geom_vline(xintercept=2020, color = "black", linewidth=1.5, alpha = 0.5,linetype="dashed") +
    geom_segment(aes(x=2040, y=0, xend=2040, yend=max(plotData %>% filter(scenario == mainScen, period == 2040) %>% pull(value))), linetype="dashed", color="black") +
    geom_line(data=plotData %>% filter(scenario == mainScen), aes(color=variable), linewidth=1.5) +
    #geom_line(data=h,linewidth=1,linetype="dashed") +
    geom_ribbon(aes(ymax = ribbon.max, ymin = ribbon.min, fill=variable), alpha = 0.3) + 
    geom_point(data=plotData %>% filter(scenario == mainScen, period == 2040), aes(color=variable),size=3,shape=21,fill="white",stroke=1.5) +
    scale_color_manual(values = c(setNames(plotColor[names(vars)],vars)), labels = setNames(names(vars), vars)) +
    scale_fill_manual(values = c(setNames(plotColor[names(vars)],vars)), labels = setNames(names(vars), vars)) +
    theme_doubleColumn() + 
    expand_limits(y=c(0,1)) +
    guides(color=guide_legend(nrow=2)) +
  theme(axis.title.y = element_blank())

ggsave(paste0(outFolder,"/Industry Emissions.svg"), g[["Industry Emissions"]] , device="svg", width = 6, height = 6, dpi=100, units = "in")
ggsave(paste0(outFolder,"/Industry Emissions.png"), g[["Industry Emissions"]] , device="png", width = 6, height = 6, dpi=300, units = "in")

```

```{r, echo=FALSE, fig.width=6, fig.height=6}

g[["Industry Emissions"]]

```

:::

::: {}

(c) Buildings useful energy
    Electricity based heating (EJ/yr)

```{r, echo=FALSE}

vars <- c(
  "Heat pump" = "UE|Buildings|Heating|Electricity|Heat pump",
  #"District\nHeating" = "UE|Buildings|Heating|District Heating",
  "Resistance" = "UE|Buildings|Heating|Electricity|Resistance"
#  "non-Heating Electricity" = "UE|Buildings|non-Heating|Electricity|Conventional"
)

plotData <- df %>% 
  filter(region == "EU27", period %in% c(2020,2030,2040,2050), tgt2030 %in% tgt2030Scen, variable %in% unlist(vars)) %>% 
  mutate(variable = factor(variable, levels=vars))  %>% 
  filter(tgt2050 != "Npi")  %>%
  group_by(model,scenario,region,period) %>% 
  mutate(cumsum = cumsum(value),
         sum = sum(value)) %>% 
  group_by(model,region,variable,period) %>% 
  mutate(ribbon.max=max(cumsum),
         ribbon.min=min(cumsum)) %>% 
  group_by(model,region,period) %>% 
  mutate(ribbon.max.total=max(sum),
         ribbon.min.total=min(sum))

colors <- c(
  "Heat pump" = "#D62828",
  "District\nHeating" = "#E07A5F",
  "Resistance" = "#F77F00",
  "Heating - Gases" = as.vector(options$aesthetics$legend$color["Gases"]),
  "Heating - Hydrogen" = as.vector(options$aesthetics$legend$color["Hydrogen"]),
  "Heating - Liquids" = as.vector(options$aesthetics$legend$color["Liquids"]),
  "Heating - Solids" = as.vector(options$aesthetics$legend$color["Solids"]),
  "non-Heating Electricity" = as.vector(options$aesthetics$legend$color["Electricity"])
)

g[["Buildings Heating - Useful energy - Electricity Based"]] <- 
  ggplot(plotData,aes(x=period,y=value)) +
  geom_col(data=plotData %>% filter(scenario == mainScen), aes(fill = variable, alpha=factor(period))) +
  geom_errorbar(data=plotData %>% filter(period != 2020, variable == vars[1], scenario == mainScen), aes(ymin=ribbon.min.total, ymax=ribbon.max.total), width=4, colour="black", alpha=0.9, size=1,position = position_dodge(width=2)) +
  # scale_y_continuous(labels = scales::percent, breaks = c(0,1,plotData[plotData$period == "2030",]$value,plotData[plotData$period == "2040",]$value,plotData[plotData$period == "2050",]$value)) +
  coord_flip() +
  theme_doubleColumn() +
  theme(panel.grid.minor.x = element_blank(),
        panel.grid.major.y = element_blank()) +
  #guides(y = "none") +
  theme(axis.title.y=element_blank()) +
  #theme(legend.position = "none") #+
  scale_fill_manual(values = setNames(colors[names(vars)],vars), labels = setNames(names(vars), vars), guide = guide_legend(reverse = TRUE, nrow = 1)) +
  scale_alpha_manual(values = c("2020"=0.7,"2030"=0.7,"2040"=1,"2050"=0.7), guide = 'none') +
  theme(legend.text=element_text(size=theme.double.text.size)) +
  theme(legend.key.height = unit(0.7, "cm"),
    legend.key.width = unit(0.7,"cm")) +
  scale_y_continuous(breaks = c(0,round(sum(plotData %>% filter(scenario == mainScen, period == "2020") %>% pull(value)),1),round(sum(plotData %>% filter(scenario == mainScen, period == "2030") %>% pull(value)),1),round(sum(plotData %>% filter(scenario == mainScen, period == "2040") %>% pull(value)),1),round(sum(plotData %>% filter(scenario == mainScen, period == "2050") %>% pull(value)),1))) #+
  #scale_x_reverse()

ggsave(paste0(outFolder,"/Buildings Heating - Useful energy - Electricity Based.svg"), g[["Buildings Heating - Useful energy - Electricity Based"]] , device="svg", width = 6, height = 4, dpi=100, units = "in")
ggsave(paste0(outFolder,"/Buildings Heating - Useful energy - Electricity Based.png"), g[["Buildings Heating - Useful energy - Electricity Based"]] , device="png", width = 6, height = 4, dpi=300, units = "in")

```

```{r, echo=FALSE, fig.width=6, fig.height=4}

g[["Buildings Heating - Useful energy - Electricity Based"]]

```

(f) Carbon Management (Mt CO2/yr)

```{r, echo=FALSE}

vars <- list(
  "atmospheric/biogenic CO2 (CDR)" = "Carbon Management|atmospheric/biogenic CO2 - CDR",
  "CCS of fossil origin" = "Carbon Management|CCS of fossil origin",
  "DAC" = "Carbon Management|Storage|DAC",
  "BECCS" = "Carbon Management|Storage|Biomass|Pe2Se", 
  "Fossil CCS" = "Carbon Management|Storage|Fossil|Pe2Se", 
  "Industry CCS" = "Carbon Management|Storage|Industry CCS"
    # "Industry Energy" = "Carbon Management|Storage|Industry Energy",
    # "Industry Process" = "Carbon Management|Storage|Industry Process"
)

plotColor <- c(
  "atmospheric/biogenic CO2 (CDR)" = "#c1d8f0",
  "CCS of fossil origin" = "#6b5b4b",
  "DAC" = "#CC6666",
  "BECCS" = "#8bc18b",
  "Industry CCS" = "#5f6769",
  "Fossil CCS" = "#47331f"
)

plotData <- df %>% 
  filter(region == "EU27", period %in% c(2030, 2040, 2050), tgt2030 %in% tgt2030Scen, variable %in% unlist(vars)) %>% 
  mutate(CDRGroup = ifelse(as.character(variable) %in% c("Carbon Management|CCS of fossil origin","Carbon Management|atmospheric/biogenic CO2 - CDR"), "CDR", "CCS"), 
         variable = factor(variable, levels=vars),
         value = ifelse(grepl("Emi",variable,fixed = TRUE), -value, value))  %>% 
  mutate(CDRGroup = factor(CDRGroup, levels=c("CDR","CCS"))) %>%
  group_by(model,scenario,region,period, CDRGroup) %>% 
  mutate(cumsum = cumsum(value),
         sum = sum(value)) %>% 
  group_by(model,region,variable,period, CDRGroup) %>% 
  mutate(ribbon.max=max(cumsum),
         ribbon.min=min(cumsum)) %>% 
  group_by(model,region,period, CDRGroup) %>% 
  mutate(ribbon.max.total=max(sum),
         ribbon.min.total=min(sum)) %>% 
  mutate(period = factor(period, levels=c(2050,2040,2030)))

 g[["Carbon Management"]] <- ggplot(plotData,aes(x=CDRGroup,y=value)) +
  #geom_area(aes(fill=variable)) +
  geom_bar(data=plotData %>% filter(scenario == mainScen), aes(fill=variable, alpha=factor(period)),stat='identity') + 
  geom_errorbar(data=plotData %>% filter(variable == vars[1], scenario == mainScen), aes(ymin=ribbon.min.total, ymax=ribbon.max.total), width=0.4, colour="black", alpha=0.9, size=1) +
  facet_wrap(period ~ ., ncol=1, strip.position="left") +
  coord_flip() +
  theme_doubleColumn() +
  scale_fill_manual(values = setNames(plotColor[names(vars)],vars), labels = setNames(names(vars), vars), guide = guide_legend(nrow = 4, reverse = TRUE)) + 
  scale_alpha_manual(values = c("2020"=0.7,"2030"=0.7,"2040"=1,"2050"=0.7), guide = 'none') +
  theme(legend.margin=margin(l = -2, unit='cm')) +
  theme(axis.title.x=element_blank()) +
    theme(legend.title = element_blank()) +
  theme(axis.title.y = element_blank()) +
  theme(legend.text=element_text(size=theme.double.text.size)) +
  theme(legend.key.height = unit(0.7, "cm"),
    legend.key.width = unit(0.7,"cm")) +
  theme(legend.margin=margin(l = -2.5, unit='cm')) + 
  theme(strip.text.y.left = element_text(angle = 0)) +
  # theme(
  #   strip.background = element_blank(),
  #   strip.text.x = element_blank()
  # ) +
  theme(axis.text.y=element_blank(), #remove x axis labels
        axis.ticks.y=element_blank(), #remove x axis ticks
        )

 ggsave(paste0(outFolder,"/Carbon Management.svg"), g[["Carbon Management"]] , device="svg", width = 6, height = 6, dpi=100, units = "in")
 ggsave(paste0(outFolder,"/Carbon Management.png"), g[["Carbon Management"]] , device="png", width = 6, height = 6, dpi=300, units = "in")

```

```{r, echo=FALSE, fig.width=6, fig.height=6}

g[["Carbon Management"]]

```

:::

::::

---

# Figure 5. Targets update

```{r, echo=FALSE}

# “existent targets”
# emission reduction - 2030 (FitFor55) - 2030 scenarios - 2040 scenarios
# energy savings  - 2030 (FitFor55 and REPowerEU) - 2030 scenarios - 2040 scenarios
# renewables target - 2030 (RED II, RED II revised, REPowerEU) - 2030 scenarios - 2040 scenarios
# solar photovoltaics capacity - 2030 (REPowerEU) - 2030 scenarios - 2040 scenarios
# CCS capacity target - 2030 EU-wide 50 MtCO2/y “Net-Zero Industry Act"
# “implicit targets”
# electrification? -> fe electrification share -> 45%
# “carbon-content in electricity” ->“carbon-free electricity” (>95%) 
# Total Residual Fossil emissions
# sectoral emissions -> Transport emissions -> 23% of 2020 
#                                   -> Industry and buildings emissions -> 27% of 2020
# “residual fossil”     -> Complete coal-phase by 2040
#                                  -> natural gas
#                                 -> oil
# “net trade” -> net-supply independency of gas by 2040
#                       -> one third of oil imports by 2040

#Emissions Reduction
plotData <- rbind(
    df %>% 
      filter(region == "EU27", period == 2030, tgt2030 %in% tgt2030Scen) %>%
      calc_addVariable( "`Emissions with LULUCF and with intra-EU aviation`" = "`Emi|GHG|LULUCF national accounting` + `Emi|CO2|Energy|Demand|Transport|International Bunkers|Intra-region`" , units = "Mt CO2/yr", only.new=T) %>% 
      mutate(value = round(1-(value / tgt %>% filter(target == "Emissions with LULUCF and with intra-EU aviation") %>% pull(reference)) ,2)) %>% 
      select(-variable)
    ,
    df %>% 
      filter(region == "EU27", period == 2040, tgt2030 %in% tgt2030Scen) %>%
      calc_addVariable( "`Emissions with LULUCF and with international transport`" = "`Emi|GHG|LULUCF national accounting` + `Emi|CO2|Energy|Demand|Transport|International Bunkers`" , units = "Mt CO2/yr", only.new=T) %>% 
      mutate(value = round(1-(value / tgt %>% filter(target == "Emissions with LULUCF and with international transport") %>% pull(reference)) ,2)) %>% 
      select(-variable)
  ) %>%
  filter(tgt2050 != "Npi") %>%
  mutate(target = "Emissions\nreductions") %>%
  group_by(period) %>% 
  mutate(min = min(value),
        max = max(value))

g[["Emissions reductions"]] <- ggplot(plotData, aes(x = target, y = value, group=period)) +
  geom_errorbar(aes(ymin=min, ymax=max), width=0.3, colour="black", alpha=0.9, linewidth=1) +
  geom_boxplot(outlier.shape = NA, width=0.6) +
         #stat_summary(geom = "text", fun = quantile,
        #              aes(label = paste0(round(..y..*100),"%")),
        #       #position=position_nudge(x=c(0.1,0.35,0.35,0.35,0.1)), 
         #      hjust=-1, angle = 90) +
  geom_hline(yintercept=0.55, color = "black", linewidth=1, alpha = 0.5, linetype = "dashed") +
  annotate("text", x = "Emissions\nreductions", y = 0.55, label = "FitFor55", angle=90, vjust = -0.5, hjust = 1, color = "black", alpha = 0.5) +
  facet_grid(target~period, scales="free") +
  theme_doubleColumn() +
  coord_flip(clip = 'off') +
  theme(axis.title.y=element_blank()) +
  theme(strip.text.y = element_blank()) +
  scale_y_continuous(breaks=seq(0, 1, 0.2), labels = scales::percent) +
  expand_limits(y=c(0,1)) + 
  theme(panel.spacing = unit(4, "lines")) +
  theme(axis.text.x = element_text(size = 0.8*theme.double.text.size)) +
  theme(axis.text.y = element_text(size = 1.2*theme.double.text.size)) +
  geom_text(data=data.frame(period=c(2030,2040),label=c("(%)",NA)), aes(label = label,x = 0, y = 0), color = "#666666", size=0.8*geom.text.double.size, hjust=1.4, na.rm = TRUE)

# Final Energy Demand - Energy savings
plotData <- df %>% filter(region == "EU27", period %in% c(2030,2040), tgt2030 %in% tgt2030Scen, variable == "FE|w/o Non-energy Use w/o Bunkers") %>% select(-variable) %>%
  mutate(value = round(value * 23.8845896627496)) %>%
  filter(tgt2050 != "Npi") %>%
  mutate(target = "Final Energy\nDemand") %>%
  group_by(period) %>% 
  mutate(min = min(value),
        max = max(value))

g[["Energy savings"]] <- ggplot(plotData, aes(x = target, y = value, group=period)) +
  geom_errorbar(aes(ymin=min, ymax=max), width=0.3, colour="black", alpha=0.9, linewidth=1) +   
  geom_boxplot(outlier.shape = NA, width=0.6) +
         #stat_summary(geom = "text", fun = quantile,
        #              aes(label = paste0(round(..y..*100),"%")),
        #       #position=position_nudge(x=c(0.1,0.35,0.35,0.35,0.1)), 
         #      hjust=-1, angle = 90) +
  geom_hline(yintercept=750, color = "black", linewidth=1, alpha = 0.5, linetype = "dashed") +
  annotate("text", x = "Final Energy\nDemand", y = 787, label = "FitFor55", angle=90, vjust = 1.5, hjust = 1, color = "black", alpha = 0.5) +
  geom_hline(yintercept=787, color = "black", linewidth=1, alpha = 0.5, linetype = "dashed") +
  annotate("text", x = "Final Energy\nDemand", y = 750, label = "REPowerEU", angle=90, vjust = -0.5, hjust = 0.7, color = "black", alpha = 0.5) +
  facet_grid(target~period, scales="free") +
  theme_doubleColumn() +
  coord_flip(clip = 'off') +
  theme(axis.title.y=element_blank()) +
  theme(strip.text.y = element_blank()) +
  theme(strip.text.x = element_blank()) +
  expand_limits(y=c(400,1000)) + 
  theme(panel.spacing = unit(4, "lines")) +
  theme(axis.text.x = element_text(size = 0.8*theme.double.text.size)) +
  theme(axis.text.y = element_text(size = 1.2*theme.double.text.size)) +
  geom_text(data=data.frame(period=c(2030,2040),label=c("(Mtoe)",NA)), aes(label = label,x = 0, y = 0), color = "#666666", size=0.8*geom.text.double.size, hjust=1.2, na.rm = TRUE)

# Electricity Demand
plotData <- df %>% filter(region == "EU27", period %in% c(2030,2040), tgt2030 %in% tgt2030Scen, variable == "FE|Electricity") %>% select(-variable) %>%
  mutate(value = round(value * 1/0.0036)) %>%
  filter(tgt2050 != "Npi") %>%
  mutate(target = "Electricity\nDemand") %>%
  group_by(period) %>% 
  mutate(min = min(value),
        max = max(value))

g[["Electricity Demand"]] <- ggplot(plotData, aes(x = target, y = value, group=period)) +
  geom_errorbar(aes(ymin=min, ymax=max), width=0.3, colour="black", alpha=0.9, linewidth=1) +   
  geom_boxplot(outlier.shape = NA, width=0.6) +
  geom_hline(yintercept=2641, color = "black", linewidth=1, alpha = 0.5, linetype = "dashed") +
  annotate("text", x = "Electricity\nDemand", y = 2641, label = "2022", angle=90, vjust = -0.5, hjust = 1.2, color = "black", alpha = 0.5) +
  facet_grid(target~period, scales="free") +
  theme_doubleColumn() +
  coord_flip(clip = 'off') +
  theme(axis.title.y=element_blank()) +
  theme(strip.text.y = element_blank()) +
  theme(strip.text.x = element_blank()) +
  expand_limits(y=c(0,5000)) + 
  theme(panel.spacing = unit(4, "lines")) +
  theme(axis.text.x = element_text(size = 0.8*theme.double.text.size)) +
  theme(axis.text.y = element_text(size = 1.2*theme.double.text.size)) +
  geom_text(data=data.frame(period=c(2030,2040),label=c("(TWh)",NA)), aes(label = label,x = 0, y = 0), color = "#666666", size=0.8*geom.text.double.size, hjust=1.2, na.rm = TRUE)

# Renewables share
plotData <- df %>% filter(region == "EU27", period %in% c(2030,2040), tgt2030 %in% tgt2030Scen, variable == "SE|Renewables Share") %>% select(-variable) %>%
  filter(tgt2050 != "Npi") %>%
  mutate(target = "Renewables\nShare in\nFinal Energy")

plotData$value <- plotData$value*(
  addHist$res %>% filter(type == "RES", period == 2020) %>% pull(value) /
   df %>% filter(region == "EU27", period == 2020, variable == "SE|Renewables Share", scenario == mainScen) %>% pull(value) 
)

plotData <- plotData %>%
  group_by(period) %>% 
  mutate(min = min(value),
        max = max(value))

g[["Renewables Share"]] <- ggplot(plotData, aes(x = target, y = value, group=period)) +
  geom_errorbar(aes(ymin=min, ymax=max), width=0.3, colour="black", alpha=0.9, linewidth=1) +   
  geom_boxplot(outlier.shape = NA, width=0.6) +
         #stat_summary(geom = "text", fun = quantile,
        #              aes(label = paste0(round(..y..*100),"%")),
        #       #position=position_nudge(x=c(0.1,0.35,0.35,0.35,0.1)), 
         #      hjust=-1, angle = 90) +
  geom_hline(yintercept=0.4, color = "black", linewidth=1, alpha = 0.5, linetype = "dashed") +
  annotate("text", x = "Renewables\nShare in\nFinal Energy", y = 0.4, label = "RED II revised", angle=90, vjust = -0.5, hjust = 0.5, color = "black", alpha = 0.5) +
  geom_hline(yintercept=0.45, color = "black", linewidth=1, alpha = 0.5, linetype = "dashed") +
  annotate("text", x = "Renewables\nShare in\nFinal Energy", y = 0.45, label = "REPowerEU", angle=90, vjust = 1.5, hjust = 0.6, color = "black", alpha = 0.5) +
  facet_grid(target~period, scales="free") +
  theme_doubleColumn() +
  coord_flip(clip = 'off') +
  theme(axis.title.y=element_blank()) +
  theme(strip.text.y = element_blank()) +
  theme(strip.text.x = element_blank()) +
  scale_y_continuous(breaks=seq(0, 1, 0.2), labels = scales::percent) +
  expand_limits(y=c(0,1)) + 
  theme(panel.spacing = unit(4, "lines")) +
  theme(axis.text.x = element_text(size = 0.8*theme.double.text.size)) +
  theme(axis.text.y = element_text(size = 1.2*theme.double.text.size)) +
  geom_text(data=data.frame(period=c(2030,2040),label=c("(%)",NA)), aes(label = label,x = 0, y = 0), color = "#666666", size=0.8*geom.text.double.size, hjust=1.4, na.rm = TRUE)

# Wind
plotData <- df %>% filter(region == "EU27", period %in% c(2030,2040), tgt2030 %in% tgt2030Scen, variable == "Cap|Electricity|Wind") %>% select(-variable) %>%
  filter(tgt2050 != "Npi") %>%
  mutate(target = "Wind\ncapacity") %>%
  group_by(period) %>% 
  mutate(min = min(value),
        max = max(value))

#target <- data.frame(target=c("REPowerEU","REPowerEU"),period=c(2025,2030),value=c(320,600)) 
g[["Wind capacity"]] <- ggplot(plotData, aes(x = target, y = value, group=period)) +
  geom_errorbar(aes(ymin=min, ymax=max), width=0.3, colour="black", alpha=0.9, linewidth=1) +   
  geom_boxplot(outlier.shape = NA, width=0.6) +
  geom_hline(yintercept=255, color = "black", linewidth=1, alpha = 0.5, linetype = "dashed") +
  annotate("text", x = "Wind\ncapacity", y = 255, label = "2022", angle=90, vjust = -0.5, hjust = 1.2, color = "black", alpha = 0.5) +
  facet_grid(target~period, scales="free") +
  theme_doubleColumn() +
  coord_flip(clip = 'off') +
  theme(axis.title.y=element_blank()) +
  theme(strip.text.y = element_blank()) +
  theme(strip.text.x = element_blank()) +
  expand_limits(y=c(0,1000)) + 
  theme(panel.spacing = unit(4, "lines")) +
  theme(axis.text.x = element_text(size = 0.8*theme.double.text.size)) +
  theme(axis.text.y = element_text(size = 1.2*theme.double.text.size)) +
  geom_text(data=data.frame(period=c(2030,2040),label=c("(GW)",NA)), aes(label = label,x = 0, y = 0), color = "#666666", size=0.8*geom.text.double.size, hjust=1.3, na.rm = TRUE)

# Solar photovoltaics
plotData <- df %>% filter(region == "EU27", period %in% c(2030,2040), tgt2030 %in% tgt2030Scen, variable == "Cap|Electricity|Solar|PV") %>% select(-variable) %>%
  filter(tgt2050 != "Npi") %>%
  mutate(target = "Solar\nphotovoltaics") %>%
  group_by(period) %>% 
  mutate(min = min(value),
        max = max(value))

#target <- data.frame(target=c("REPowerEU","REPowerEU"),period=c(2025,2030),value=c(320,600)) 
g[["Solar photovoltaics"]] <- ggplot(plotData, aes(x = target, y = value, group=period)) +
  geom_errorbar(aes(ymin=min, ymax=max), width=0.3, colour="black", alpha=0.9, linewidth=1) +   
  geom_boxplot(outlier.shape = NA, width=0.6) +
         #stat_summary(geom = "text", fun = quantile,
        #              aes(label = paste0(round(..y..*100),"%")),
        #       #position=position_nudge(x=c(0.1,0.35,0.35,0.35,0.1)), 
         #      hjust=-1, angle = 90) +
  geom_hline(yintercept=600, color = "black", linewidth=1, alpha = 0.5, linetype = "dashed") +
  annotate("text", x = "Solar\nphotovoltaics", y = 600, label = "REPowerEU", angle=90, vjust = -0.5, hjust = 0.6, color = "black", alpha = 0.5) +
  geom_hline(yintercept=208.9, color = "black", linewidth=1, alpha = 0.5, linetype = "dashed") +
  annotate("text", x = "Solar\nphotovoltaics", y = 208.9, label = "2022", angle=90, vjust = -0.5, hjust = 1.2, color = "black", alpha = 0.5) +
  facet_grid(target~period, scales="free") +
  theme_doubleColumn() +
  coord_flip(clip = 'off') +
  theme(axis.title.y=element_blank()) +
  theme(strip.text.y = element_blank()) +
  theme(strip.text.x = element_blank()) +
  expand_limits(y=c(0,2500)) + 
  theme(panel.spacing = unit(4, "lines")) +
  theme(axis.text.x = element_text(size = 0.8*theme.double.text.size)) +
  theme(axis.text.y = element_text(size = 1.2*theme.double.text.size)) +
  geom_text(data=data.frame(period=c(2030,2040),label=c("(GW)",NA)), aes(label = label,x = 0, y = 0), color = "#666666", size=0.8*geom.text.double.size, hjust=1.3, na.rm = TRUE)

# Carbon Capture and Storage 
plotData <- df %>% filter(region == "EU27", period %in% c(2030,2040), tgt2030 %in% tgt2030Scen, variable == "Carbon Management|Storage") %>% select(-variable) %>%
  filter(tgt2050 != "Npi") %>%
  mutate(target = "Carbon\nCapture and\nStorage") %>%
  group_by(period) %>% 
  mutate(min = min(value),
        max = max(value))

g[["Carbon Capture and Storage"]] <- ggplot(plotData, aes(x = target, y = value, group=period)) +
  geom_errorbar(aes(ymin=min, ymax=max), width=0.3, colour="black", alpha=0.9, linewidth=1) +   
  geom_boxplot(outlier.shape = NA, width=0.6) +
         #stat_summary(geom = "text", fun = quantile,
        #              aes(label = paste0(round(..y..*100),"%")),
        #       #position=position_nudge(x=c(0.1,0.35,0.35,0.35,0.1)), 
         #      hjust=-1, angle = 90) +
  geom_hline(yintercept=50, color = "black", linewidth=1, alpha = 0.5, linetype = "dashed") +
  annotate("text", x = "Carbon\nCapture and\nStorage", y = 50, label = "Net-Zero\nIndustry Act", angle=90, vjust = -0.2, hjust = 0.6, color = "black", alpha = 0.5) +
  facet_grid(target~period, scales="free") +
  theme_doubleColumn() +
  coord_flip(clip = 'off') +
  theme(axis.title.y=element_blank()) +
  theme(strip.text.y = element_blank()) +
  theme(strip.text.x = element_blank()) +
  #expand_limits(y=c(0,ceiling(max(plotData$value)/50)*50)) + 
  expand_limits(y=c(0,400)) + 
  theme(panel.spacing = unit(4, "lines")) +
  theme(axis.text.x = element_text(size = 0.8*theme.double.text.size)) +
  theme(axis.text.y = element_text(size = 1.2*theme.double.text.size)) +
  geom_text(data=data.frame(period=c(2030,2040),label=c("(Mt CO2/yr)",NA)), aes(label = label,x = 0, y = 0), color = "#666666", size=0.8*geom.text.double.size, hjust=1.1, na.rm = TRUE)

# Novel CDR = atmospheric/biogenic CO2 (CDR) 
plotData <- df %>% filter(region == "EU27", period %in% c(2030,2040), tgt2030 %in% tgt2030Scen, variable == "Carbon Management|atmospheric/biogenic CO2 - CDR") %>% select(-variable) %>%
  filter(tgt2050 != "Npi") %>%
  mutate(target = "Novel CDR") %>%
  group_by(period) %>% 
  mutate(min = min(value),
        max = max(value))

g[["Novel CDR"]] <- ggplot(plotData, aes(x = target, y = value, group=period)) +
  geom_errorbar(aes(ymin=min, ymax=max), width=0.3, colour="black", alpha=0.9, linewidth=1) +   
  geom_boxplot(outlier.shape = NA, width=0.6) +
         #stat_summary(geom = "text", fun = quantile,
        #              aes(label = paste0(round(..y..*100),"%")),
        #       #position=position_nudge(x=c(0.1,0.35,0.35,0.35,0.1)), 
         #      hjust=-1, angle = 90) +
  facet_grid(target~period, scales="free") +
  theme_doubleColumn() +
  coord_flip(clip = 'off') +
  theme(axis.title.y=element_blank()) +
  theme(strip.text.y = element_blank()) +
  theme(strip.text.x = element_blank()) +
  expand_limits(y=c(0,250)) + 
  #expand_limits(y=c(0,ceiling(max(plotData$value)/50)*50)) + 
  theme(panel.spacing = unit(4, "lines")) +
  theme(axis.text.x = element_text(size = 0.8*theme.double.text.size)) +
  theme(axis.text.y = element_text(size = 1.2*theme.double.text.size)) +
  geom_text(data=data.frame(period=c(2030,2040),label=c("(Mt CO2/yr)",NA)), aes(label = label,x = 0, y = 0), color = "#666666", size=0.8*geom.text.double.size, hjust=1.1, na.rm = TRUE)

# Direct Electrification
plotData <- df %>% filter(region == "EU27", period %in% c(2030,2040), tgt2030 %in% tgt2030Scen, variable == "Final Energy|Electricity Share") %>% select(-variable) %>%
  filter(tgt2050 != "Npi") %>%
  mutate(target = "Electricity\nShare in\nFinal Energy") %>%
  group_by(period) %>% 
  mutate(min = min(value),
        max = max(value))

g[["Direct Electrification"]] <- ggplot(plotData, aes(x = target, y = value, group=period)) +
    geom_errorbar(aes(ymin=min, ymax=max), width=0.3, colour="black", alpha=0.9, linewidth=1) +   
  geom_boxplot(outlier.shape = NA, width=0.6) +
         #stat_summary(geom = "text", fun = quantile,
        #              aes(label = paste0(round(..y..*100),"%")),
        #       #position=position_nudge(x=c(0.1,0.35,0.35,0.35,0.1)), 
         #      hjust=-1, angle = 90) +
  facet_grid(target~period, scales="free") +
  theme_doubleColumn() +
  coord_flip(clip = 'off') +
  theme(axis.title.y=element_blank()) +
  theme(strip.text.y = element_blank()) +
  theme(strip.text.x = element_blank()) +
  scale_y_continuous(breaks=seq(0, 1, 0.2), labels = scales::percent) +
  expand_limits(y=c(0,1)) + 
  theme(panel.spacing = unit(4, "lines")) +
  theme(axis.text.x = element_text(size = 0.8*theme.double.text.size)) +
  theme(axis.text.y = element_text(size = 1.2*theme.double.text.size)) +
  geom_text(data=data.frame(period=c(2030,2040),label=c("(%)",NA)), aes(label = label,x = 0, y = 0), color = "#666666", size=0.8*geom.text.double.size, hjust=1.4, na.rm = TRUE)

# Carbon-free electricity share
plotData <- df %>% filter(region == "EU27", period %in% c(2030,2040), tgt2030 %in% tgt2030Scen, variable == "SE|Electricity|carbon-free") %>% select(-variable) %>%
  filter(tgt2050 != "Npi") %>%
  mutate(target = "Carbon-free\nelectricity\nshare") %>%
  group_by(period) %>% 
  mutate(min = min(value),
        max = max(value))

g[["Carbon-free electricity share"]] <- ggplot(plotData, aes(x = target, y = value, group=period)) +
    geom_errorbar(aes(ymin=min, ymax=max), width=0.3, colour="black", alpha=0.9, linewidth=1) +   
  geom_boxplot(outlier.shape = NA, width=0.6) +
         #stat_summary(geom = "text", fun = quantile,
        #              aes(label = paste0(round(..y..*100),"%")),
        #       #position=position_nudge(x=c(0.1,0.35,0.35,0.35,0.1)), 
         #      hjust=-1, angle = 90) +
  facet_grid(target~period, scales="free") +
  theme_doubleColumn() +
  coord_flip(clip = 'off') +
  theme(axis.title.y=element_blank()) +
  theme(strip.text.y = element_blank()) +
  theme(strip.text.x = element_blank()) +
  scale_y_continuous(breaks=seq(0, 1, 0.2), labels = scales::percent) +
  expand_limits(y=c(0,1)) + 
  theme(panel.spacing = unit(4, "lines")) +
  theme(axis.text.x = element_text(size = 0.8*theme.double.text.size)) +
  theme(axis.text.y = element_text(size = 1.2*theme.double.text.size)) +
  geom_text(data=data.frame(period=c(2030,2040),label=c("(%)",NA)), aes(label = label,x = 0, y = 0), color = "#666666", size=0.8*geom.text.double.size, hjust=1.4, na.rm = TRUE)

# Residual Fossil Emissions
plotData <- df %>% filter(region == "EU27", period %in% c(2030,2040), tgt2030 %in% tgt2030Scen, variable == "Emi|CO2|Residual Fossil") %>% select(-variable) %>%
  filter(tgt2050 != "Npi") %>%
  mutate(target = "Residual\nFossil\nEmissions") %>%
  group_by(period) %>% 
  mutate(min = min(value),
        max = max(value))

g[["Residual Fossil Emissions"]] <- ggplot(plotData, aes(x = target, y = value, group=period)) +
  geom_errorbar(aes(ymin=min, ymax=max), width=0.3, colour="black", alpha=0.9, linewidth=1) +   
  geom_boxplot(outlier.shape = NA, width=0.6) +
         #stat_summary(geom = "text", fun = quantile,
        #              aes(label = paste0(round(..y..*100),"%")),
        #       #position=position_nudge(x=c(0.1,0.35,0.35,0.35,0.1)), 
         #      hjust=-1, angle = 90) +
  facet_grid(target~period, scales="free") +
  theme_doubleColumn() +
  coord_flip(clip = 'off') +
  theme(axis.title.y=element_blank()) +
  theme(strip.text.y = element_blank()) +
  theme(strip.text.x = element_blank()) +
  expand_limits(y=c(0,2500)) + 
  theme(panel.spacing = unit(4, "lines")) +
  theme(axis.text.x = element_text(size = 0.8*theme.double.text.size)) +
  theme(axis.text.y = element_text(size = 1.2*theme.double.text.size)) +
  geom_text(data=data.frame(period=c(2030,2040),label=c("(Mt CO2/yr)",NA)), aes(label = label,x = 0, y = 0), color = "#666666", size=0.8*geom.text.double.size, hjust=1.05, na.rm = TRUE)

g[["Targets"]] <- gridExtra::arrangeGrob(
  g[["Emissions reductions"]] + theme(strip.text.x = element_blank()) + theme(plot.margin=margin(b=20, l = 30)), #+ theme(plot.margin=margin(10,10,30,10)),margin(t = 0, r = 0, b = 0, l = 0, unit = "pt")
  g[["Energy savings"]] + theme(plot.margin=margin(b=20, l = 10)),
  g[["Electricity Demand"]] + theme(plot.margin=margin(b=20, l = 35)),
  g[["Renewables Share"]] + theme(plot.margin=margin(b=20, l = 15)),
  g[["Wind capacity"]] + theme(plot.margin=margin(b=20, l = 50)),
  g[["Solar photovoltaics"]] + theme(plot.margin=margin(b=20, l = 10)),
  g[["Carbon Capture and Storage"]] + theme(plot.margin=margin(b=20, l = 15)),
  g[["Novel CDR"]] + theme(plot.margin=margin(b=20, l = 27)),
  g[["Direct Electrification"]] + theme(plot.margin=margin(b=20, l = 17)),
  g[["Carbon-free electricity share"]] + theme(plot.margin=margin(b=20, l = 24)),
  g[["Residual Fossil Emissions"]] + theme(plot.margin=margin(b=20, l = 39)),
  ncol = 1
)

ggsave(paste0(outFolder,"/targets.svg"), g[["Targets"]] , device="svg", width = 12, height = 17, dpi=100, units = "in")
ggsave(paste0(outFolder,"/targets.png"), g[["Targets"]] , device="png", width = 12, height = 17, dpi=300, units = "in")

```

```{r, echo=FALSE, fig.width=12, fig.height=17}

grid.newpage()
grid.draw(g[["Targets"]])

```