### 2024-08-30 report_plotlyCharts.Rmd

---
title: "plotly charts for sensitivity iteractive explorer html"
output:
  html_document:
    theme: paper
    toc: true
    toc_float:
      collapsed: false
---

<style>
  .main-container {
    max-width: 95% !important;
  }
  .toc-content {
    padding-left: 0px !important;
  }
  .svg-container {
    margin: auto !important;
  }
</style>

Description:
Plotly charts to be used in iteractive html

Rodrigues, R., Pietzcker, R., Sitarz, J., Merfort, A., Hasse, R., Hoppe, J., Pehl, M., Ershad, A. M., Baumstark, L., Luderer, G. 2040 greenhouse gas reduction targets and energy transitions in line with the EU Green Deal. under revision (2023).

<!-- Setting default chunk options -->
```{r chunkOptions, echo=FALSE, include = FALSE}
#knitr::opts_chunk$set(echo = TRUE)
# setting global R chunk options (https://yihui.name/knitr/options/#chunk_options)
  knitr::opts_chunk$set(dev='svglite', # svg, png,...
                        fig.ext = ".svg",
                        #fig.asp = .8 # default aspect ratio
                        fig.width = 12,
                        fig.height = 6,
                        dpi=100
                        )
```

```{r loading_packages, echo=FALSE, include = FALSE}

  #Loading required packages
  # install.packages("remotes")
  #remotes::install_github("coolbutuseless/ggpattern")
  #install.packages('rsvg')
  #remotes::install_github('coolbutuseless/ggsvg')
  packagesList <- c("quitte","ggplot2","geomtextpath","ggpattern","tidyr","grid","stringr","gridExtra","ggrepel","kableExtra","dplyr","remind2","rsvg","ggsvg","plotly")
  packages <- suppressWarnings(suppressMessages(lapply(packagesList, function(x){ if(!require(x, character.only = T, quietly = T)){ install.packages(x); return(paste0("installed missing package: ", x)) } else paste0("required package is installed: ", x) } )))

```

<!-- Preprocess: prepare and load the data for use in the charts and markdown -->

```{r preprocess, echo=FALSE, include = FALSE}

#Options
forceUpdateData = FALSE
loadHist = TRUE

# data and output folders
folders <- file.info(list.files(path="./data", full.names = T))
dataFolder <- folders %>% tibble::rownames_to_column(var = "path") %>% arrange(desc(path)) %>% slice(1) %>% pull(path)

source("./R/preprocess.R")

#create output folders
outFolder <- paste0("./output/",gsub("./data/","",dataFolder))
jsOutputFolder <- paste0(outFolder, "/report_sensitivityExplorer/charts")
dir.create(jsOutputFolder, recursive = TRUE, showWarnings = FALSE)

```

<!-- Function to convert plotly charts to js file  -->

```{r loadAuxFunction, echo=FALSE, include = FALSE}

source("./R/plotly2js.R")

```

<!-- Initializing charts object -->
```{r chart_initialization, echo=FALSE, include = FALSE}

g <- NULL

```

---

# Carbon price

```{r chart_carbonPrice, echo=FALSE, include = FALSE}

  vars <- c(
    "Carbon price" = "Price|Carbon"
  )

  plotData <- df_beforeFilter %>%
    filter(region == "EU27", period %in% c(2020:2050), tgt2030 %in% tgt2030Scen, variable %in% vars) %>%
    mutate(period = as.Date(paste0("01/01/", period),format="%m/%d/%Y"),
           value = 1.17*value) %>%
    mutate(value = round(value,2))

  ETS_price <- addHist$ETS_price %>%
    mutate(scenario = "ETS price")

  cc <- c("reference"="#ae0000", "EED 2018 Eff"="#db4646", "FitFor55 eff"="#0090ab", "RePowerEU eff"="#095786")
  tmp <- plotData %>% mutate(color = cc[efficiency])   
  scenarioColor <- setNames(tmp$color,tmp$scenario)

  ff <- c("BioLim4"="#fedc97", "BioLim7.5"="#b5b682", "BioLim12"="#7c9885", "BioLim20"="#28666e")
  tmp <- plotData %>% mutate(color = cc[bioLim])   
  scenarioFill <- setNames(tmp$color,tmp$scenario)

  g[["Carbon Price"]][["ggplot"]] <- ggplot(plotData,aes(x=period,y=value,group = scenario)) +
    geom_hline(yintercept=0, color = "black", linewidth=2, alpha = 0.5) +
    geom_vline(xintercept=2020, color = "black", linewidth=1.5, alpha = 0.5,linetype="dashed") +
    geom_line(linewidth = 0.5, aes(linetype=tgt2030, color = scenario)) +
    geom_point(size = 3, aes(shape = lim, fill = scenario)) +
    geom_line(data=ETS_price,aes(x=Date, y=Price),size=0.5, color = "#000000") +
    scale_x_date(date_labels = "%Y") +
    scale_linetype_manual(values = c("55"="twodash", "57"="solid", "59"="dashed"), guide = guide_legend(nrow =1, byrow = T, order = 1, override.aes = list(linewidth = 1.5))) + 
    scale_color_manual(values = scenarioColor, guide = guide_legend(nrow =4, byrow = T, order = 2, override.aes = list(linewidth = 1.5))) + 
    scale_fill_manual(values = scenarioFill, guide = guide_legend(nrow =2, byrow = T, override.aes = list(shape = 21, size = 8), order = 3)) +
    scale_shape_manual(values = c("default"=21,"limCCS"=22,"limH2"=23,"limVRE"=24), guide = guide_legend(nrow =2, byrow = T, order = 4, override.aes = list(size = 8))) +
    theme_doubleColumn() +
    theme(strip.text.y = element_blank()) +
    #guides(fill = guide_legend(nrow = 2, byrow = T)) +
    theme(legend.text=element_text(size=theme.single.text.size)) +
    theme(legend.key.height = unit(0.7, "cm"),
          legend.key.width = unit(2,"cm")) +
    theme(legend.position="right", legend.direction = "vertical", legend.box = "vertical") + 
    #theme(legend.key.width = unit(2,"cm")) +
    #ylab("€/t CO2") +
    theme(legend.spacing.y = unit(0.5, "lines")) + 
    theme(legend.position = "none")

  g[["Carbon Price"]][["plotly"]] <- ggplotly(g[["Carbon Price"]][["ggplot"]]) %>% layout(autosize = TRUE)

  #save plotly js
  plotly2js(
    g[["Carbon Price"]][["plotly"]],
    div.id = "carbonPrice", output.html = TRUE, output.file = "carbonPrice.js", 
    output.dir = jsOutputFolder, output.url = paste0(jsOutputFolder, "/carbonPrice.html"))

```

```{r, echo=FALSE, fig.width=8, fig.height=6}

  g[["Carbon Price"]][["plotly"]]

```

---

# Emissions

```{r chart_emissions, echo=FALSE, include = FALSE}

  h <- refEmi %>% filter(variable == "Emissions with LULUCF and with international transport", region == "EU27", pollutant == "All greenhouse gases - (CO2 equivalent)") %>% ungroup() %>% select(region,variable,period,value) %>%
    mutate(scenario = "historical") %>%
    mutate(value = round(value,2))

  plotData <- df_beforeFilter %>% 
    filter(region == "EU27", period %in% c(2020:2060), tgt2030 %in% tgt2030Scen) %>%
    calc_addVariable( "`emiIntTrnsp`" = "`Emi|GHG|LULUCF national accounting` + `Emi|CO2|Energy|Demand|Transport|International Bunkers`" , units = "Mt CO2/yr") %>%
    calc_addVariable( "`emiIntraEU`" = "`Emi|GHG|LULUCF national accounting` + `Emi|CO2|Energy|Demand|Transport|International Bunkers|Intra-region`" , units = "Mt CO2/yr") %>%
    filter(variable %in% c("emiIntTrnsp","emiIntraEU")) %>%
    pivot_wider(names_from = variable, values_from = value) %>%
    mutate(ref_emiIntTrnsp = tgt %>% filter(target == "Emissions with LULUCF and with intra-EU aviation") %>% pull(reference),
           ref_emiIntraEU = tgt %>% filter(target == "Emissions with LULUCF and with international transport") %>% pull(reference)) %>%
    mutate(red_emiIntTrnsp = emiIntTrnsp/ref_emiIntTrnsp,
           red_emiIntraEU = emiIntraEU/ref_emiIntraEU) %>%
    mutate(value = emiIntTrnsp,
           reduction = ifelse(period >= 2045, red_emiIntTrnsp, red_emiIntraEU)) %>%
    mutate(text = paste0("Emissions: ", round(value), " Mt CO2/yr", "\nReduction: ", round(1-reduction,3)*100, "%", "\nPeriod: ", period, "\nScenario: ", scenario, "\nTarget 2030: ", tgt2030, "%", "\nEfficiency: ", efficiency, "\nBiomass Limit: ", bioLim, "\nSensititivy limit: ", lim)) %>%
    mutate(value = round(value,2))

  g[["Emissions"]][["ggplot"]] <- ggplot(plotData,aes(x=period,y=value,group = scenario)) +
    geom_hline(yintercept=0, color = "black", linewidth=2, alpha = 0.5) +
    geom_vline(xintercept=2020, color = "black", linewidth=1.5, alpha = 0.5,linetype="dashed") +
    geom_line(size = 0.5, aes(linetype=tgt2030, color = scenario, text = text)) +
    geom_point(size = 3, aes(shape = lim, fill = scenario, text = text)) +
    geom_line(data=h,size=0.5, color = "#000000") +
    scale_linetype_manual(values = c("55"="twodash", "57"="solid", "59"="dashed"), guide = guide_legend(nrow =1, byrow = T, order = 1, override.aes = list(linewidth = 1.5))) +   scale_color_manual(values = scenarioColor, guide = guide_legend(nrow =4, byrow = T, order = 2, override.aes = list(linewidth = 1.5))) + 
    scale_fill_manual(values = scenarioFill, guide = guide_legend(nrow =2, byrow = T, override.aes = list(shape = 21, size = 8), order = 3)) +
    scale_shape_manual(values = c("default"=21,"limCCS"=22,"limH2"=23,"limVRE"=24), guide = guide_legend(nrow =2, byrow = T, order = 4, override.aes = list(size = 8))) +
    theme_doubleColumn() +
    theme(strip.text.y = element_blank()) +
    #guides(fill = guide_legend(nrow = 2, byrow = T)) +
    theme(legend.text=element_text(size=theme.single.text.size)) +
    theme(legend.key.height = unit(0.7, "cm"),
          legend.key.width = unit(2,"cm")) +
    theme(legend.position="right", legend.direction = "vertical", legend.box = "vertical") + 
    #theme(legend.key.width = unit(2,"cm")) +
    theme(legend.spacing.y = unit(0.5, "lines")) + 
    theme(legend.position = "none") +
    scale_x_continuous(breaks = seq(1990,2060,10))

  g[["Emissions"]][["plotly"]] <- ggplotly(g[["Emissions"]][["ggplot"]],tooltip = c("text")) %>% layout(autosize = TRUE)

  plotly2js(
    g[["Emissions"]][["plotly"]],
    div.id = "emissions", output.html = TRUE, output.file = "emissions.js", 
    output.dir = jsOutputFolder, output.url = paste0(jsOutputFolder, "/emissions.html"))

```

```{r, echo=FALSE, fig.width=8, fig.height=6}

  g[["Emissions"]][["plotly"]]

```

---

# Final Energy

```{r chart_finalEnergy, echo=FALSE, include = FALSE}

  plotData <- df_beforeFilter %>% 
    filter(region == "EU27", period %in% c(2020:2050), tgt2030 %in% tgt2030Scen, variable == "FE") %>%
    mutate(value = round(value,2))

  h <- hist %>% filter(model == "IEA", region == "EU27", variable == "FE", period >= 2005) %>% drop_na(value) %>% filter(value != 0) %>%
    mutate(value = round(value,2))

  g[["finalEnergy"]][["ggplot"]] <- ggplot(plotData,aes(x=period,y=value,group = scenario)) +
    geom_hline(yintercept=0, color = "black", linewidth=2, alpha = 0.5) +
    geom_vline(xintercept=2020, color = "black", linewidth=1.5, alpha = 0.5,linetype="dashed") +
    geom_line(size = 0.5, aes(linetype=tgt2030, color = scenario)) +
    geom_point(size = 3, aes(shape = lim, fill = scenario)) +
    geom_line(data=h,size=0.5, color = "#000000") +
    scale_linetype_manual(values = c("55"="twodash", "57"="solid", "59"="dashed"), guide = guide_legend(nrow =1, byrow = T, order = 1, override.aes = list(linewidth = 1.5))) + 
    scale_color_manual(values = scenarioColor, guide = guide_legend(nrow =4, byrow = T, order = 2, override.aes = list(linewidth = 1.5))) + 
    scale_fill_manual(values = scenarioFill, guide = guide_legend(nrow =2, byrow = T, override.aes = list(shape = 21, size = 8), order = 3)) +
    scale_shape_manual(values = c("default"=21,"limCCS"=22,"limH2"=23,"limVRE"=24), guide = guide_legend(nrow =2, byrow = T, order = 4, override.aes = list(size = 8))) +
    theme_doubleColumn() +
    theme(strip.text.y = element_blank()) +
    #guides(fill = guide_legend(nrow = 2, byrow = T)) +
    theme(legend.text=element_text(size=theme.single.text.size)) +
    theme(legend.key.height = unit(0.7, "cm"),
          legend.key.width = unit(2,"cm")) +
    theme(legend.position="right", legend.direction = "vertical", legend.box = "vertical") + 
    #theme(legend.key.width = unit(2,"cm")) +
    theme(legend.spacing.y = unit(0.5, "lines")) + 
    theme(legend.position = "none") +
    scale_x_continuous(breaks = seq(1990,2060,10))

  g[["finalEnergy"]][["plotly"]] <- ggplotly(g[["finalEnergy"]][["ggplot"]]) %>% layout(autosize = TRUE)

  plotly2js(
    g[["finalEnergy"]][["plotly"]],
    div.id = "finalEnergy", output.html = TRUE, output.file = "finalEnergy.js", 
    output.dir = jsOutputFolder, output.url = paste0(jsOutputFolder, "/finalEnergy.html"))

```

```{r, echo=FALSE, fig.width=8, fig.height=6}

  g[["Emissions"]][["plotly"]]

```

---

# Electricity share

```{r chart_electricityShare, echo=FALSE, include = FALSE}

  vars <- c("Total" = "Final Energy|Electricity Share",
            "Transport"="Final Energy|Transport|Electricity Share",
            "Industry"="Final Energy|Industry|Electricity Share",
            "Buildings"="Final Energy|Buildings|Electricity Share")

  invertVars <- setNames(names(vars),vars)

  plotData <- df_beforeFilter %>% 
    filter(region == "EU27", period %in% c(2020:2050), tgt2030 %in% tgt2030Scen, variable %in% unlist(vars)) %>%
    mutate(sector = factor(invertVars[as.character(variable)], c("Total", "Transport", "Industry", "Buildings"))) %>%
    mutate(value = round(value,4))

  h_Orig <- hist[hist$model == "IEA" & hist$region == "EU27" & hist$variable %in% c("FE","FE|Electricity","FE|Transport","FE|Transport|Electricity","FE|Industry", "FE|Industry|Electricity","FE|Buildings","FE|Buildings|Electricity"),] %>%
    drop_na(value) %>%
    mutate(value = round(value,4))

  h <- rbind(
    h_Orig %>% calc_addVariable("`Final Energy|Electricity Share`" = "`FE|Electricity` / `FE`", units = "%", only.new=T),
    h_Orig %>% calc_addVariable("`Final Energy|Transport|Electricity Share`" = "`FE|Transport|Electricity` / `FE|Transport`", units = "%", only.new=T),
    h_Orig %>% calc_addVariable("`Final Energy|Industry|Electricity Share`"  = "`FE|Industry|Electricity`  / `FE|Industry`", units = "%", only.new=T),
    h_Orig %>% calc_addVariable("`Final Energy|Buildings|Electricity Share`" = "`FE|Buildings|Electricity` / `FE|Buildings`", units = "%", only.new=T)) %>%
    mutate(sector = factor(invertVars[as.character(variable)], c("Total", "Transport", "Industry", "Buildings"))) %>%
    filter(period >= 2010)

  g[["Electricity Share"]][["ggplot"]] <- ggplot(plotData,aes(x=period,y=value,group = scenario)) +
    geom_hline(yintercept=0, color = "black", linewidth=2, alpha = 0.5) +
    geom_vline(xintercept=2020, color = "black", linewidth=1.5, alpha = 0.5,linetype="dashed") +
    geom_line(size = 0.5, aes(linetype=tgt2030, color = scenario)) +
    geom_point(size = 1.5, aes(shape = lim, fill = scenario)) +
    geom_line(data=h,size=0.5, color = "#000000") +
    #geom_point(size = 3, aes(shape = lim, fill = scenario)) +
    facet_wrap(~ sector) +
    #geom_line(data=h %>% filter(period >= 2005),linetype="dashed", linewidth = 1, aes(color=variable)) +
    # scale_colour_manual(values = c("Final Energy|Electricity Share"="black","Final Energy|Transport|Electricity Share"="#a1dd70","Final Energy|Industry|Electricity Share"="#5f6769","Final Energy|Buildings|Electricity Share"="#5edfff"), labels = setNames(names(vars), vars)) +
    # scale_fill_manual(values = c("Final Energy|Electricity Share"="black","Final Energy|Transport|Electricity Share"="#a1dd70","Final Energy|Industry|Electricity Share"="#5f6769","Final Energy|Buildings|Electricity Share"="#5edfff"), labels = setNames(names(vars), vars)) +
    scale_linetype_manual(values = c("55"="twodash", "57"="solid", "59"="dashed"), guide = guide_legend(nrow =1, byrow = T, order = 1, override.aes = list(linewidth = 1.5))) + 
    scale_color_manual(values = scenarioColor) + 
    scale_fill_manual(values = scenarioFill, guide = guide_legend(nrow =2, byrow = T, override.aes = list(shape = 21, size = 8), order = 3)) +
    scale_shape_manual(values = c("default"=21,"limCCS"=22,"limH2"=23,"limVRE"=24), guide = guide_legend(nrow =2, byrow = T, order = 4, override.aes = list(size = 8))) +
    theme_doubleColumn() +
    scale_y_continuous(breaks = seq(0, 1, 0.2),labels = scales::percent) +
    expand_limits(y=c(0,1)) +
    theme(legend.position = "none") +
    theme(axis.title.y=element_blank())

  g[["Electricity Share"]][["plotly"]] <- ggplotly(g[["Electricity Share"]][["ggplot"]]) %>% layout(autosize = TRUE)

  plotly2js(
    g[["Electricity Share"]][["ggplot"]],
    div.id = "electricityShare", output.html = TRUE, output.file = "electricityShare.js", 
    output.dir = jsOutputFolder, output.url = paste0(jsOutputFolder, "/electricityShare.html"))

```

```{r, echo=FALSE, fig.width=8, fig.height=6}

  g[["Electricity Share"]][["plotly"]]

```

---

#Hydrogen

```{r chart_hydrogen, echo=FALSE, include = FALSE}

  vars <- c(
    "Black H2 (coal)" = "SE|Hydrogen|Coal|w/o CC",
    "Gray H2 (gas)" = "SE|Hydrogen|Gas|w/o CC",
    "Blue H2 (fossil w/ CCS)" = c("SE|Hydrogen|Coal|w/ CC", "SE|Hydrogen|Gas|w/ CC"),
    "H2 imports" = "SE|Hydrogen|Net Imports",
    #  "Bio-based H2 (gasification)" = c("SE|Hydrogen|Biomass|w/ CC","SE|Hydrogen|Biomass|w/o CC"),
    #  "Green H2 (electrolysis)" = "SE|Hydrogen|Electricity",
    "Green and bio-based H2" = c("SE|Hydrogen|Electricity", "SE|Hydrogen|Biomass|w/ CC","SE|Hydrogen|Biomass|w/o CC")
    #"VRE Storage Electrolysis" = "SE|Hydrogen|Electricity|VRE Storage Electrolysis",
    #"Standard Electrolysis" = "SE|Hydrogen|Electricity|Standard Electrolysis",
  )

  invertVars <- setNames(names(vars),vars)

  plotData <- df_beforeFilter %>% 
    filter(region == "EU27", period %in% c(2020:2050), tgt2030 %in% tgt2030Scen, variable %in% unlist(vars)) %>%
    mutate(type = gsub("\\d$", "", as.character(invertVars[as.character(variable)]))) %>%
    group_by(model,scenario,region,unit,period,type,tgt2030,lim) %>%
    summarize(value=sum(value),.groups = "keep") %>%
    mutate(type=factor(type, levels=unique(gsub("\\d$", "", names(vars))))) %>%
    mutate(value = round(value,2))

  g[["hydrogen"]][["ggplot"]] <- ggplot(plotData,aes(x=period,y=value,group = scenario)) +
    geom_hline(yintercept=0, color = "black", linewidth=2, alpha = 0.5) +
    geom_vline(xintercept=2020, color = "black", linewidth=1.5, alpha = 0.5,linetype="dashed") +
    geom_line(size = 0.5, aes(linetype=tgt2030, color = scenario)) +
    geom_point(size = 1.5, aes(shape = lim, fill = scenario)) +
    facet_wrap(~ type, nrow=2) +
    scale_linetype_manual(values = c("55"="twodash", "57"="solid", "59"="dashed"), guide = guide_legend(nrow =1, byrow = T, order = 1, override.aes = list(linewidth = 1.5))) + 
    scale_color_manual(values = scenarioColor) + 
    scale_fill_manual(values = scenarioFill, guide = guide_legend(nrow =2, byrow = T, override.aes = list(shape = 21, size = 8), order = 3)) +
    scale_shape_manual(values = c("default"=21,"limCCS"=22,"limH2"=23,"limVRE"=24), guide = guide_legend(nrow =2, byrow = T, order = 4, override.aes = list(size = 8))) +
    theme_doubleColumn() +
    theme(legend.position = "none") +
    theme(axis.title.y=element_blank())

  g[["hydrogen"]][["plotly"]] <- ggplotly(g[["hydrogen"]][["ggplot"]]) %>% layout(autosize = TRUE)

  plotly2js(
    g[["hydrogen"]][["plotly"]],
    div.id = "hydrogen", output.html = TRUE, output.file = "hydrogen.js", 
    output.dir = jsOutputFolder, output.url = paste0(jsOutputFolder, "/hydrogen.html"))

```

```{r, echo=FALSE, fig.width=8, fig.height=6}

  g[["hydrogen"]][["plotly"]]

```

---

#BECCS

```{r chart_beccs, echo=FALSE, include = FALSE}

  vars <- c(
    "BECCS" = "Emi|CO2|CDR|BECCS"
  )

  plotData <- df_beforeFilter %>%
    filter(region == "EU27", period %in% c(2020:2050), tgt2030 %in% tgt2030Scen, variable %in% vars) %>%
    mutate(value = round(value,2))

  g[["BECCS"]][["ggplot"]] <- ggplot(plotData,aes(x=period,y=value,group = scenario)) +
    geom_hline(yintercept=0, color = "black", linewidth=2, alpha = 0.5) +
    geom_vline(xintercept=2020, color = "black", linewidth=1.5, alpha = 0.5,linetype="dashed") +
    geom_line(size = 0.5, aes(linetype=tgt2030, color = scenario)) +
    geom_point(size = 3, aes(shape = lim, fill = scenario)) +
    #scale_x_date(date_labels = "%Y") +
    scale_linetype_manual(values = c("55"="twodash", "57"="solid", "59"="dashed"), guide = guide_legend(nrow =1, byrow = T, order = 1, override.aes = list(linewidth = 1.5))) + 
    scale_color_manual(values = scenarioColor) + 
    #scale_color_manual(values = c("reference"="#ae0000", "EED 2018 Eff"="#db4646", "FitFor55 eff"="#0090ab", "RePowerEU eff"="#095786"), guide = guide_legend(nrow =4, byrow = T, order = 2, override.aes = list(linewidth = 1.5))) + 
    #scale_fill_manual(values = c("BioLim4"="#fedc97", "BioLim7.5"="#b5b682", "BioLim12"="#7c9885", "BioLim20"="#28666e"), guide = guide_legend(nrow =2, byrow = T, override.aes = list(shape = 21, size = 8), order = 3)) +
    scale_fill_manual(values = scenarioFill, guide = guide_legend(nrow =2, byrow = T, override.aes = list(shape = 21, size = 8), order = 3)) +
    scale_shape_manual(values = c("default"=21,"limCCS"=22,"limH2"=23,"limVRE"=24), guide = guide_legend(nrow =2, byrow = T, order = 4, override.aes = list(size = 8))) +
    theme_doubleColumn() +
    theme(strip.text.y = element_blank()) +
    #guides(fill = guide_legend(nrow = 2, byrow = T)) +
    theme(legend.text=element_text(size=theme.single.text.size)) +
    theme(legend.key.height = unit(0.7, "cm"),
          legend.key.width = unit(2,"cm")) +
    theme(legend.position="right", legend.direction = "vertical", legend.box = "vertical") + 
    #theme(legend.key.width = unit(2,"cm")) +
    theme(legend.spacing.y = unit(0.5, "lines")) +
    theme(legend.position = "none")

  g[["BECCS"]][["plotly"]] <- ggplotly(g[["BECCS"]][["ggplot"]]) %>% layout(autosize = TRUE)

  plotly2js(
    g[["BECCS"]][["plotly"]],
    div.id = "beccs", output.html = TRUE, output.file = "beccs.js", 
    output.dir = jsOutputFolder, output.url = paste0(jsOutputFolder, "/beccs.html"))

```

```{r, echo=FALSE, fig.width=8, fig.height=6}

  g[["BECCS"]][["plotly"]]

```

---

#DAC

```{r chart_dac, echo=FALSE, include = FALSE}

  vars <- c(
    "DAC" = "Emi|CO2|CDR|DACCS"
  )

  plotData <- df_beforeFilter %>%
    filter(region == "EU27", period %in% c(2020:2050), tgt2030 %in% tgt2030Scen, variable %in% vars) %>%
    mutate(value = round(value,2))

  g[["DAC"]][["ggplot"]] <- ggplot(plotData,aes(x=period,y=value,group = scenario)) +
    geom_hline(yintercept=0, color = "black", linewidth=2, alpha = 0.5) +
    geom_vline(xintercept=2020, color = "black", linewidth=1.5, alpha = 0.5,linetype="dashed") +
    geom_line(size = 0.5, aes(linetype=tgt2030, color = scenario)) +
    geom_point(size = 3, aes(shape = lim, fill = scenario)) +
    #scale_x_date(date_labels = "%Y") +
    scale_linetype_manual(values = c("55"="twodash", "57"="solid", "59"="dashed"), guide = guide_legend(nrow =1, byrow = T, order = 1, override.aes = list(linewidth = 1.5))) + 
    scale_color_manual(values = scenarioColor) + 
    #scale_color_manual(values = c("reference"="#ae0000", "EED 2018 Eff"="#db4646", "FitFor55 eff"="#0090ab", "RePowerEU eff"="#095786"), guide = guide_legend(nrow =4, byrow = T, order = 2, override.aes = list(linewidth = 1.5))) + 
    #scale_fill_manual(values = c("BioLim4"="#fedc97", "BioLim7.5"="#b5b682", "BioLim12"="#7c9885", "BioLim20"="#28666e"), guide = guide_legend(nrow =2, byrow = T, override.aes = list(shape = 21, size = 8), order = 3)) +
    scale_fill_manual(values = scenarioFill, guide = guide_legend(nrow =2, byrow = T, override.aes = list(shape = 21, size = 8), order = 3)) +
    scale_shape_manual(values = c("default"=21,"limCCS"=22,"limH2"=23,"limVRE"=24), guide = guide_legend(nrow =2, byrow = T, order = 4, override.aes = list(size = 8))) +
    theme_doubleColumn() +
    theme(strip.text.y = element_blank()) +
    #guides(fill = guide_legend(nrow = 2, byrow = T)) +
    theme(legend.text=element_text(size=theme.single.text.size)) +
    theme(legend.key.height = unit(0.7, "cm"),
          legend.key.width = unit(2,"cm")) +
    theme(legend.position="right", legend.direction = "vertical", legend.box = "vertical") + 
    #theme(legend.key.width = unit(2,"cm")) +
    theme(legend.spacing.y = unit(0.5, "lines")) +
    theme(legend.position = "none")

  g[["DAC"]][["plotly"]] <- ggplotly(g[["DAC"]][["ggplot"]]) %>% layout(autosize = TRUE)

  plotly2js(
    g[["DAC"]][["plotly"]],
    div.id = "dac", output.html = TRUE, output.file = "dac.js", 
    output.dir = jsOutputFolder, output.url = paste0(jsOutputFolder, "/dac.html"))

```

```{r, echo=FALSE, fig.width=8, fig.height=6}

  g[["DAC"]][["plotly"]]

```

---

#Emissions per sector

```{r chart_sectorEmi, echo=FALSE, include = FALSE}

  vars <- c(
    "Industrial Processes" = "Emi|GHG|Industrial Processes",
    "Industry" = "Emi|GHG|Gross|Energy|Demand|Industry",
    "Buildings" ="Emi|GHG|Energy|Demand|Buildings",
    "Transportation" ="Emi|GHG|Energy|Demand|Transport",
    "Bunkers" = "Emi|CO2|Energy|Demand|Transport|International Bunkers",
    "Heat"="Emi|CO2|Gross|Energy|Supply|Heat",
    "Other Energy Conversion" = "Emi|CO2|Gross|Other Energy Conversion",
    "Electricity"="Emi|GHG|Gross|Energy|Supply|Electricity"
  )

  invertVars <- setNames(names(vars),vars)

  plotData <- df_beforeFilter %>% 
    filter(region == "EU27", period %in% c(2020:2050), tgt2030 %in% tgt2030Scen, variable %in% vars) %>%
    mutate(sector = factor(invertVars[as.character(variable)], names(vars))) %>%
    mutate(value = round(value,2))

  g[["sectorEmi"]][["ggplot"]] <- ggplot(plotData,aes(x=period,y=value,group = scenario)) +
    geom_hline(yintercept=0, color = "black", linewidth=2, alpha = 0.5) +
    geom_vline(xintercept=2020, color = "black", linewidth=1.5, alpha = 0.5,linetype="dashed") +
    geom_line(size = 0.5, aes(linetype=tgt2030, color = scenario)) +
    geom_point(size = 1.5, aes(shape = lim, fill = scenario)) +
    #geom_point(size = 3, aes(shape = lim, fill = scenario)) +
    facet_wrap(~ sector, nrow=2) +
    #geom_line(data=h %>% filter(period >= 2005),linetype="dashed", linewidth = 1, aes(color=variable)) +
    # scale_colour_manual(values = c("Final Energy|Electricity Share"="black","Final Energy|Transport|Electricity Share"="#a1dd70","Final Energy|Industry|Electricity Share"="#5f6769","Final Energy|Buildings|Electricity Share"="#5edfff"), labels = setNames(names(vars), vars)) +
    # scale_fill_manual(values = c("Final Energy|Electricity Share"="black","Final Energy|Transport|Electricity Share"="#a1dd70","Final Energy|Industry|Electricity Share"="#5f6769","Final Energy|Buildings|Electricity Share"="#5edfff"), labels = setNames(names(vars), vars)) +
    scale_linetype_manual(values = c("55"="twodash", "57"="solid", "59"="dashed"), guide = guide_legend(nrow =1, byrow = T, order = 1, override.aes = list(linewidth = 1.5))) + 
    scale_color_manual(values = scenarioColor) + 
    scale_fill_manual(values = scenarioFill, guide = guide_legend(nrow =2, byrow = T, override.aes = list(shape = 21, size = 8), order = 3)) +
    scale_shape_manual(values = c("default"=21,"limCCS"=22,"limH2"=23,"limVRE"=24), guide = guide_legend(nrow =2, byrow = T, order = 4, override.aes = list(size = 8))) +
    theme_doubleColumn() +
    theme(legend.position = "none") +
    theme(axis.title.y=element_blank())

  g[["sectorEmi"]][["plotly"]] <- ggplotly(g[["sectorEmi"]][["ggplot"]]) %>% layout(autosize = TRUE)

  plotly2js(
    g[["sectorEmi"]][["plotly"]],
    div.id = "sectorEmi", output.html = TRUE, output.file = "sectorEmi.js", 
    output.dir = jsOutputFolder, output.url = paste0(jsOutputFolder, "/sectorEmi.html"))

```

```{r, echo=FALSE, fig.width=8, fig.height=6}

  g[["sectorEmi"]][["plotly"]]

```

---

# Primary Energy

```{r chart_primaryEnergy, echo=FALSE, include = FALSE}

  vars <- c("Oil"="PE|Oil",
            "Gas"="PE|Gas",
            "Coal"="PE|Coal",
            "Biomass" = "PE|Biomass")

  invertVars <- setNames(names(vars),vars)

  plotData <- df_beforeFilter %>% 
    filter(region == "EU27", period %in% c(2020:2050), tgt2030 %in% tgt2030Scen, variable %in% vars) %>%
    mutate(pe = factor(invertVars[as.character(variable)], names(vars))) %>%
    mutate(value = round(value,2))

  h <- hist %>% filter(model == "IEA", region == "EU27", variable %in% vars, period >= 2005, period <= 2020) %>% drop_na(value) %>% filter(value != 0) %>%
    mutate(pe = factor(invertVars[as.character(variable)], names(vars))) %>%
    mutate(value = round(value,2))

  g[["primaryEnergy"]][["ggplot"]] <- ggplot(plotData,aes(x=period,y=value,group = scenario)) +
    geom_hline(yintercept=0, color = "black", linewidth=2, alpha = 0.5) +
    geom_vline(xintercept=2020, color = "black", linewidth=1.5, alpha = 0.5,linetype="dashed") +
    geom_line(size = 0.5, aes(linetype=tgt2030, color = scenario)) +
    geom_point(size = 1.5, aes(shape = lim, fill = scenario)) +
    geom_line(data=h,size=0.5, color = "#000000") +
    #geom_point(size = 3, aes(shape = lim, fill = scenario)) +
    facet_wrap(~ pe, nrow=2) +
    #geom_line(data=h %>% filter(period >= 2005),linetype="dashed", linewidth = 1, aes(color=variable)) +
    # scale_colour_manual(values = c("Final Energy|Electricity Share"="black","Final Energy|Transport|Electricity Share"="#a1dd70","Final Energy|Industry|Electricity Share"="#5f6769","Final Energy|Buildings|Electricity Share"="#5edfff"), labels = setNames(names(vars), vars)) +
    # scale_fill_manual(values = c("Final Energy|Electricity Share"="black","Final Energy|Transport|Electricity Share"="#a1dd70","Final Energy|Industry|Electricity Share"="#5f6769","Final Energy|Buildings|Electricity Share"="#5edfff"), labels = setNames(names(vars), vars)) +
    scale_linetype_manual(values = c("55"="twodash", "57"="solid", "59"="dashed"), guide = guide_legend(nrow =1, byrow = T, order = 1, override.aes = list(linewidth = 1.5))) + 
    scale_color_manual(values = scenarioColor) + 
    scale_fill_manual(values = scenarioFill, guide = guide_legend(nrow =2, byrow = T, override.aes = list(shape = 21, size = 8), order = 3)) +
    scale_shape_manual(values = c("default"=21,"limCCS"=22,"limH2"=23,"limVRE"=24), guide = guide_legend(nrow =2, byrow = T, order = 4, override.aes = list(size = 8))) +
    theme_doubleColumn() +
    theme(legend.position = "none") +
    theme(axis.title.y=element_blank())

  g[["primaryEnergy"]][["plotly"]] <- ggplotly(g[["primaryEnergy"]][["ggplot"]]) %>% layout(autosize = TRUE)

  plotly2js(
    g[["primaryEnergy"]][["plotly"]],
    div.id = "primaryEnergy", output.html = TRUE, output.file = "primaryEnergy.js", 
    output.dir = jsOutputFolder, output.url = paste0(jsOutputFolder, "/primaryEnergy.html"))

```

```{r, echo=FALSE, fig.width=8, fig.height=6}

  g[["primaryEnergy"]][["plotly"]]

```

---

# Electricity Generation

```{r chart_generation, echo=FALSE, include = FALSE}

  vars <- c(
    #"Net Imports" = "SE|Electricity|Net Imports",
    "Coal" = "SE|Electricity|Coal",
    #"Coal w/ CC" = "SE|Electricity|Coal|w/ CC",
    #"Coal w/o CC" = "SE|Electricity|Coal|w/o CC",
    #"Oil" = "SE|Electricity|Oil",
    "Gas w/ CC" = "SE|Electricity|Gas|w/ CC",
    "Gas w/o CC" = "SE|Electricity|Gas|w/o CC",
    "Hydrogen" = "SE|Electricity|Hydrogen",
    "Nuclear" = "SE|Electricity|Nuclear",
    #"Geothermal" = "SE|Electricity|Geothermal",
    #"Hydro" = "SE|Electricity|Hydro",
    "Biomass" = "SE|Electricity|Biomass",
    #"Biomass w/ CC" = "SE|Electricity|Biomass|w/ CC",
    #"Biomass w/o CC" = "SE|Electricity|Biomass|w/o CC",
    "Solar" = "SE|Electricity|Solar",
    #"Solar CSP" = "SE|Electricity|Solar|CSP",
    #"Solar PV" = "SE|Electricity|Solar|PV",
    "Wind" = "SE|Electricity|Wind"
    #"Wind Onshore" = "SE|Electricity|Wind|Onshore",
    #"Wind Offshore" = "SE|Electricity|Wind|Offshore"
  )

  invertVars <- setNames(names(vars),vars)

  plotData <- df_beforeFilter %>% 
    filter(region == "EU27", period %in% c(2020:2050), tgt2030 %in% tgt2030Scen, variable %in% vars) %>%
    mutate(seel = factor(invertVars[as.character(variable)], names(vars))) %>%
    mutate(value = round(value * 1/0.0036))

  g[["generation"]][["ggplot"]] <- ggplot(plotData,aes(x=period,y=value,group = scenario)) +
    geom_hline(yintercept=0, color = "black", linewidth=2, alpha = 0.5) +
    geom_vline(xintercept=2020, color = "black", linewidth=1.5, alpha = 0.5,linetype="dashed") +
    geom_line(size = 0.5, aes(linetype=tgt2030, color = scenario)) +
    geom_point(size = 1.5, aes(shape = lim, fill = scenario)) +
    #geom_point(size = 3, aes(shape = lim, fill = scenario)) +
    facet_wrap(~ seel, nrow=2) +
    #geom_line(data=h %>% filter(period >= 2005),linetype="dashed", linewidth = 1, aes(color=variable)) +
    # scale_colour_manual(values = c("Final Energy|Electricity Share"="black","Final Energy|Transport|Electricity Share"="#a1dd70","Final Energy|Industry|Electricity Share"="#5f6769","Final Energy|Buildings|Electricity Share"="#5edfff"), labels = setNames(names(vars), vars)) +
    # scale_fill_manual(values = c("Final Energy|Electricity Share"="black","Final Energy|Transport|Electricity Share"="#a1dd70","Final Energy|Industry|Electricity Share"="#5f6769","Final Energy|Buildings|Electricity Share"="#5edfff"), labels = setNames(names(vars), vars)) +
    scale_linetype_manual(values = c("55"="twodash", "57"="solid", "59"="dashed"), guide = guide_legend(nrow =1, byrow = T, order = 1, override.aes = list(linewidth = 1.5))) + 
    scale_color_manual(values = scenarioColor) + 
    scale_fill_manual(values = scenarioFill, guide = guide_legend(nrow =2, byrow = T, override.aes = list(shape = 21, size = 8), order = 3)) +
    scale_shape_manual(values = c("default"=21,"limCCS"=22,"limH2"=23,"limVRE"=24), guide = guide_legend(nrow =2, byrow = T, order = 4, override.aes = list(size = 8))) +
    theme_doubleColumn() +
    theme(legend.position = "none") +
    theme(axis.title.y=element_blank())

  g[["generation"]][["plotly"]] <- ggplotly(g[["generation"]][["ggplot"]]) %>% layout(autosize = TRUE)

  plotly2js(
    ggplotly(ggplotly(g[["generation"]][["plotly"]])) %>% layout(autosize = TRUE),
    div.id = "generation", output.html = TRUE, output.file = "generation.js", 
    output.dir = jsOutputFolder, output.url = paste0(jsOutputFolder, "/generation.html"))

```

```{r, echo=FALSE, fig.width=8, fig.height=6}

  g[["generation"]][["plotly"]]

```

<!-- # # Final Energy -->
<!-- # vars <- c( -->
<!-- #   "Electricity" = "FE|Electricity", -->
<!-- #   "Heat" = "FE|Heat", -->
<!-- #   "Hydrogen" = "FE|Hydrogen", -->
<!-- #   "Gases" = "FE|Gases", -->
<!-- #   #"Gases" = "FE|Gases|Biomass", -->
<!-- #   #"Gases" = "FE|Gases|Hydrogen", -->
<!-- #   #"Gases" = "FE|Gases|Fossil", -->
<!-- #   "Liquids" = "FE|Liquids", -->
<!-- #   #"Liquids" = "FE|Liquids|Biomass", -->
<!-- #   #"Liquids" = "FE|Liquids|Hydrogen", -->
<!-- #   #"Liquids" = "FE|Liquids|Fossil", -->
<!-- #   "Solids" = "FE|Solids"	 -->
<!-- #   #"Solids" = "FE|Solids|Biomass",	 -->
<!-- #   #"Solids" = "FE|Solids|Fossil"	 -->
<!-- # ) -->
<!-- #  -->
<!-- # invertVars <- setNames(names(vars),vars) -->
<!-- #  -->
<!-- # plotData <- df_beforeFilter %>%  -->
<!-- #   filter(region == "EU27", period %in% c(2020:2060), tgt2030 %in% tgt2030Scen, variable %in% vars) %>% -->
<!-- #   mutate(fe = factor(invertVars[as.character(variable)], names(vars))) -->
<!-- #  -->
<!-- # g[["finalEnergy"]] <- ggplot(plotData,aes(x=period,y=value,group = scenario)) + -->
<!-- #     geom_hline(yintercept=0, color = "black", linewidth=2, alpha = 0.5) + -->
<!-- #     geom_vline(xintercept=2020, color = "black", linewidth=1.5, alpha = 0.5,linetype="dashed") + -->
<!-- #     geom_line(size = 0.5, aes(linetype=tgt2030, color = scenario)) + -->
<!-- #     facet_wrap(~ fe, nrow=2) + -->
<!-- #     scale_linetype_manual(values = c("55"="twodash", "57"="solid", "59"="dashed"), guide = guide_legend(nrow =1, byrow = T, order = 1, override.aes = list(linewidth = 1.5))) +  -->
<!-- #     scale_color_manual(values = scenarioColor) +  -->
<!-- #     scale_fill_manual(values = scenarioFill, guide = guide_legend(nrow =2, byrow = T, override.aes = list(shape = 21, size = 8), order = 3)) + -->
<!-- #     scale_shape_manual(values = c("default"=21,"limCCS"=22,"limH2"=23,"limVRE"=24), guide = guide_legend(nrow =2, byrow = T, order = 4, override.aes = list(size = 8))) + -->
<!-- #     theme_doubleColumn() + -->
<!-- #     theme(legend.position = "none") + -->
<!-- #     theme(axis.title.y=element_blank())  -->
<!-- #  -->
<!-- # plotly2js( -->
<!-- #   ggplotly(ggplotly(g[["finalEnergy"]])) %>% layout(autosize = TRUE), -->
<!-- #              div.id = "finalEnergy", output.html = TRUE, output.file = "finalEnergy.js",  -->
<!-- #              output.dir = jsOutputFolder, output.url = paste0(jsOutputFolder, "/finalEnergy.html")) -->